---
title: "Best Picture Nominees"
editor_options: 
  chunk_output_type: console
output:
  html_document:
    includes:
      in_header: ../header/header.html
      after_body: ../include_footer.html
output_file: "projects/best_picture_nominees.html"
---

```{r setup, include=FALSE}
# Set default chunk output
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  results = "asis"
)
```

```{r libraries}
library(tidyverse)
library(DT)
library(scales)
library(plotly)
```



```{r data}
best_pic_noms_final = read_csv(here::here("data", "best_picture_nominees.csv"))

```

Best Picture Nominees are cool too. Incredibly difficult to scrape though... IMDb sometimes implores JavaScript in their html, so I had to parse through even more noise than usual to get to my info. 

But now, I have all the Best Picture nominees since the first Academy Awards!


## Distributors

IMDb has a section called "Production companies" in their Details section of each movie homepage. I will define a Distributor as the first listed Production Company.

```{r distibutor_leaderboards}

only_winners = function(x) {
  no_noms = gsub(x = x, pattern = ".*(%)", replacement = "") 
  
  no_noms[no_noms != ""] %>%
    paste0(collapse = "\n")
}


distributors_plot_nominees = best_pic_noms_final %>%
  separate_rows(producers, sep = "\\|") %>%
  group_by(id) %>% 
  slice(1) %>% 
  group_by(distributor = producers) %>% 
  summarise(n = n(),
            all_titles = paste0(title, ifelse(nominees_or_winner == "winner", " *", "%"),
                               collapse = "\n"),
            n_winners = sum(nominees_or_winner == "winner")) %>% 
  ungroup() %>%
  mutate(check = strsplit(all_titles, split = "\\\n"),
         winners = check %>% lapply(only_winners) %>% unlist,
         all_titles = gsub(x = all_titles, pattern = "\\%", replacement = "")) %>% 
  arrange(desc(n)) %>%
  slice(1:15) %>% 
  ggplot(aes(x = n, y= reorder(distributor, -n), text = all_titles, fill = distributor)) +
  geom_col()+
  # labs(caption = "* = winners")+
  theme(axis.title.y=element_blank(),
        axis.title.x=element_blank(),
        plot.caption = element_text(colour = "gold"),
        legend.position = "none") +
  ggtitle("Number of Nominees per Distributor (Top 15)") +
  geom_text(aes(label = n), hjust = 0, position = position_dodge(width = 1)) +
  geom_col(aes(x = n_winners, y = reorder(distributor, -n)), fill = "gold")


distributors_plot_nominees %>% ggplotly(tooltip = c("x", "text", "fill"), height = 800) %>% 
  layout(annotations = 
           list(x = 1, y = -0.1, text = "* = winners", 
                showarrow = F, xref='paper', yref='paper', 
                xanchor='right', yanchor='auto', xshift=0, yshift=0,
                font=list(size=15, color="gold")))

```

For more detail, please look at this table:

```{r distributor_table}
best_pic_noms_final %>%
  separate_rows(producers, sep = "\\|") %>%
  group_by(id) %>% 
  slice(1) %>% 
  group_by(distributor = producers) %>% 
  summarise(n = n(), all_titles = paste(title, collapse = "<br>")) %>%
  ungroup() %>% 
  arrange(desc(n)) %>% 
  datatable(filter = "top", extensions = "Buttons",
            options = list(paging = FALSE, dom = "Bfrtip", buttons = c("copy", "csv", "excel"),
                           scrollY = "400px", scrollX = TRUE),
            escape = FALSE)
```

## Producers

```{r producer_leaderboards}

producers_plot_nominees = best_pic_noms_final %>%
  separate_rows(producers, sep = "\\|") %>%
  # group_by(id) %>% 
  # slice(1) %>% 
  group_by(producers) %>%
  summarise(n = n(),
            all_titles = paste0(title, ifelse(nominees_or_winner == "winner", " *", "%"),
                               collapse = "\n"),
            n_winners = sum(nominees_or_winner == "winner")) %>% 
  ungroup() %>%
  mutate(check = strsplit(all_titles, split = "\\\n"),
         winners = check %>% lapply(only_winners) %>% unlist,
         all_titles = gsub(x = all_titles, pattern = "\\%", replacement = "")) %>% 
  arrange(desc(n)) %>%
  slice(1:15) %>% 
  ggplot(aes(x = n, y= reorder(producers, -n), text = all_titles, fill = producers)) +
  geom_col()+
  # labs(caption = "* = winners")+
  theme(axis.title.y=element_blank(),
        axis.title.x=element_blank(),
        plot.caption = element_text(colour = "gold"),
        legend.position = "none") +
  ggtitle("Number of Nominees per producers (Top 15)") +
  geom_text(aes(label = n), hjust = 0, position = position_dodge(width = 1)) +
  geom_col(aes(x = n_winners, y = reorder(producers, -n)), fill = "gold")


producers_plot_nominees %>% ggplotly(tooltip = c("x", "text", "fill"), height = 800) %>% 
  layout(annotations = 
           list(x = 1, y = -0.1, text = "* = winners", 
                showarrow = F, xref='paper', yref='paper', 
                xanchor='right', yanchor='auto', xshift=0, yshift=0,
                font=list(size=15, color="gold")))

```

For more detail, please look at this table:

```{r producer_table}
best_pic_noms_final %>%
  separate_rows(producers, sep = "\\|") %>%
  # group_by(id) %>% 
  # slice(1) %>% 
  group_by(producers) %>% 
  summarise(n = n(), all_titles = paste(title, collapse = "<br>")) %>%
  ungroup() %>% 
  arrange(desc(n)) %>% 
  datatable(filter = "top", extensions = "Buttons",
            options = list(paging = FALSE, dom = "Bfrtip", buttons = c("copy", "csv", "excel"),
                           scrollY = "400px", scrollX = TRUE),
            escape = FALSE)
```



## All Nominees

```{r full_table}
best_pic_noms_final %>%
  select(year, title, nominees_or_winner, producers, imdb_url) %>% 
  datatable(filter = "top", extensions = "Buttons",
            options = list(paging = FALSE, dom = "Bfrtip", buttons = c("copy", "csv", "excel"),
                           scrollY = "400px", scrollX = TRUE))


```


