---
title: "IMDb Rating Plot by Season and Episode"
editor_options: 
  chunk_output_type: console
output:
  html_document:
    self_contained: false
    includes:
      in_header: ../header/header.html
      after_body: ../include_footer.html
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

I'm kinda obsessed with seeing a whole map of series and where they stride well and when they just absolutely fuck different episodes. Sometimes, critics are just a little harsh. Sometimes it's hard to completely crush an entire series, but we can explore some series that figure it out and quite while they're ahead.


```{r ci-outdir, include=FALSE}
# ensure output dir exists in clean CI environment
out_dir <- here::here("imdb_rating_plot_files")
if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)
```


```{r setup, include=FALSE}
# Set default chunk output
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  results = "asis"
)
```


```{r episode_plot}
library(tidyverse)
library(RColorBrewer)

# Function to create a plot of IMDb ratings by season and episode
create_episode_plot <- function(episode_data, title, export_png = TRUE) {
  # make sure we have imdb_rating + hover text
  episode_data <- episode_data %>%
    dplyr::mutate(
      imdb_rating = rating,
      hover_text = paste0(
        "S", stringr::str_pad(season, 2, pad = "0"),
        "E", stringr::str_pad(episode, 2, pad = "0"),
        " — ", title,
        "\nRating: ", sprintf("%.1f", imdb_rating),
        if (!"votes" %in% names(episode_data)) "" else
          paste0("\nVotes: ", scales::comma(votes))
      )
    )

  file_name <- title %>%
    tolower() %>%
    stringr::str_remove_all(":") %>%
    stringr::str_replace_all(" ", "_") %>%
    paste0(., ".png")

  episode_plot <- episode_data %>%
    ggplot2::ggplot(
      ggplot2::aes(
        x = season,
        y = episode,
        fill = imdb_rating,
        text = hover_text
      )
    ) +
    ggplot2::geom_tile(color = "white") +
    ggplot2::geom_text(
      ggplot2::aes(label = round(imdb_rating, 1)),
      color = "black"
    ) +
    ggplot2::scale_fill_gradientn(
      colors = RColorBrewer::brewer.pal(9, "RdYlGn"),
      values = scales::rescale(c(0, 4, 6, 8, 10)),
      limits = c(0, 10),
      name = "IMDb Rating"
    ) +
    ggplot2::scale_x_continuous(
      breaks = seq(min(episode_data$season), max(episode_data$season), 1),
      position = "top"
    ) +
    ggplot2::scale_y_reverse(
      breaks = seq(min(episode_data$episode), max(episode_data$episode), 1)
    ) +
    ggplot2::labs(
      title = "IMDb Ratings by Season and Episode",
      x = "Season",
      y = "Episode"
    ) +
    ggplot2::ggtitle(title) +
    ggplot2::theme_minimal()

  if (export_png) {
    ggplot2::ggsave(
      filename = file.path(out_dir, file_name),
      plot = episode_plot,
      width = 8,
      height = 6,
      dpi = 300
    )
  }

  episode_plot
}

interactive_plotly = function(gg_object) {
  gg_object %>%
    plotly::ggplotly(tooltip = "text") %>%
    plotly::layout(xaxis = list(side = "top"))
}
```

## Game of Thrones

```{r}
game_of_thrones = read_csv(here::here("data", "game_of_thrones_ep_ratings.csv"))

game_of_thrones %>% create_episode_plot("Game of Thrones") %>% interactive_plotly
```

## Breaking Bad

```{r}
breaking_bad = read_csv(here::here("data", "breaking_bad_ep_ratings.csv"))

breaking_bad %>% create_episode_plot("Breaking Bad") %>% interactive_plotly
```

## Avatar: The Last Airbender

```{r}
avatar = read_csv(here::here("data", "avatar_ep_ratings.csv"))

avatar %>% create_episode_plot("Avatar: The Last Airbender") %>% interactive_plotly
```

## Always Sunny in Philadelphia

```{r}
always_sunny = read_csv(here::here("data", "always_sunny_ep_ratings.csv"))

always_sunny %>% create_episode_plot("Always Sunny in Philadelphia") %>% interactive_plotly
```

## The Office (US)

```{r}
the_office = read_csv(here::here("data", "the_office_ep_ratings.csv"))

the_office %>% create_episode_plot("The Office (US)") %>% interactive_plotly
```

## HIMYM

```{r}
himym = read_csv(here::here("data", "himym_ep_ratings.csv"))

himym %>% create_episode_plot("How I Met Your Mother") %>% interactive_plotly
```

## Better Call Saul

```{r}
better_call_saul = read_csv(here::here("data", "better_call_saul_ep_ratings.csv"))

better_call_saul %>% create_episode_plot("Better Call Saul") %>% interactive_plotly
```

## ZeroZeroZero

```{r}
zerozerozero = read_csv(here::here("data", "zerozerozero_ep_ratings.csv"))

zerozerozero %>% create_episode_plot("ZeroZeroZero") %>% interactive_plotly
```

## House of the Dragon

```{r}
hotd = read_csv(here::here("data", "hotd_ep_ratings.csv"))

hotd %>% create_episode_plot("House of the Dragon") %>% interactive_plotly
```

## Bluey

```{r fig.height=9, fig.width=4}
bluey = read_csv(here::here("data", "bluey_ep_ratings.csv"))

bluey %>% create_episode_plot("Bluey") %>% interactive_plotly
```

## Westworld

```{r}
westworld = read_csv(here::here("data", "westworld_ep_ratings.csv"))

westworld %>% create_episode_plot("Westworld") %>% interactive_plotly
```

## Paris Hilton's My New BFF

```{r}
paris_hilton_bff = read_csv(here::here("data", "paris_hilton_bff_ep_ratings.csv"))

paris_hilton_bff %>% create_episode_plot("Paris Hilton's My New BFF") %>% interactive_plotly
```

## Stranger Things

```{r}
stranger_things = read_csv(here::here("data", "stranger_things_ep_ratings.csv"))

stranger_things %>% create_episode_plot("Stranger Things") %>% interactive_plotly
```

## Bojack Horseman

```{r}
bojack = read_csv(here::here("data", "bojack_ep_ratings.csv"))

bojack %>% create_episode_plot("Bojack Horseman") %>% interactive_plotly
```

## Velma

```{r}
velma = read_csv(here::here("data", "velma_ep_ratings.csv"))

velma %>% create_episode_plot("Velma") %>% interactive_plotly
```

## Suits

```{r}
suits = read_csv(here::here("data", "suits_ep_ratings.csv"))

suits %>% create_episode_plot("Suits") %>% interactive_plotly
```

## Vampire Diaries

```{r}
vampire_diaries = read_csv(here::here("data", "vampire_diaries_ep_ratings.csv"))

vampire_diaries %>% create_episode_plot("Vampire Diaries") %>% interactive_plotly
```

## Mad Men

```{r}
mad_men = read_csv(here::here("data", "mad_men_ep_ratings.csv"))

mad_men %>% create_episode_plot("Mad Men") %>% interactive_plotly
```

## Grey's Anatomy

```{r fig.height=9, fig.width=9}
greys_anatomy = read_csv(here::here("data", "greys_anatomy_ep_ratings.csv"))

greys_anatomy %>% create_episode_plot("Grey's Anatomy") %>% interactive_plotly
```

## Friends

```{r fig.height=7, fig.width=7}
friends = read_csv(here::here("data", "friends_ep_ratings.csv"))

friends %>% create_episode_plot("Friends") %>% interactive_plotly
```

## House MD

```{r fig.height=7, fig.width=7}
house = read_csv(here::here("data", "house_ep_ratings.csv"))

house %>% create_episode_plot("House MD") %>% interactive_plotly
```

## The Simpsons

```{r fig.width=15}
simpsons = read_csv(here::here("data", "simpsons_ep_ratings.csv"))

simpsons_gg = simpsons %>% create_episode_plot("The Simpsons")
simpsons_gg %>% interactive_plotly
# resave here for the correct width
ggsave(file.path(out_dir, "the_simpsons.png"), plot = simpsons_gg, width = 16, height = 6, dpi = 300)
```

## Pretty Little Liars

```{r}
pretty_little_liars = read_csv(here::here("data", "pretty_little_liars_ep_ratings.csv"))

pretty_little_liars %>% create_episode_plot("Pretty Little Liars") %>% interactive_plotly
```

## The Crown

```{r}
crown = read_csv(here::here("data", "crown_ep_ratings.csv"))

crown %>% create_episode_plot("The Crown") %>% interactive_plotly
```

## Love Island (UK)

```{r fig.height=11}
love_island = read_csv(here::here("data", "love_island_ep_ratings.csv"))

love_island %>% create_episode_plot("Love Island (UK)") %>% interactive_plotly
```

## South Park

```{r fig.width=15}
south_park = read_csv(here::here("data", "south_park_ep_ratings.csv"))

south_park_gg = south_park %>% create_episode_plot("South Park")
south_park_gg %>% interactive_plotly
# resave here for the correct width
ggsave(file.path(out_dir, "south_park.png"), plot = south_park_gg, width = 16, height = 6, dpi = 300)
```

## Family Guy

```{r fig.height=9, fig.width=15}
family_guy = read_csv(here::here("data", "family_guy_ep_ratings.csv"))

family_guy_gg = family_guy %>% create_episode_plot("Family Guy")
family_guy_gg %>% interactive_plotly
# resave here for the correct width
ggsave(file.path(out_dir, "family_guy.png"), plot = family_guy_gg, width = 16, height = 9, dpi = 300)
```

## Invincible

```{r}
invincible = read_csv(here::here("data", "invincible_ep_ratings.csv"))

invincible %>% create_episode_plot("Invincible") %>% interactive_plotly
```

## Clone Wars

```{r}
clone_wars = read_csv(here::here("data", "clone_wars_ep_ratings.csv"))

clone_wars %>% create_episode_plot("Clone Wars") %>% interactive_plotly
```

## Mr Robot

```{r}
mr_robot = read_csv(here::here("data", "mr_robot_ep_ratings.csv"))

mr_robot %>% create_episode_plot("Mr Robot") %>% interactive_plotly
```

## Sparticus

```{r}
sparticus = read_csv(here::here("data", "sparticus_ep_ratings.csv"))

sparticus %>% create_episode_plot("Sparticus") %>% interactive_plotly
```

## Attack on Titan

```{r fig.height=7}
attack_on_titan = read_csv(here::here("data", "attack_on_titan_ep_ratings.csv"))

attack_on_titan %>% create_episode_plot("Attack on Titan") %>% interactive_plotly
```

## Blue Eye Samurai

```{r}
blue_eye_samurai = read_csv(here::here("data", "blue_eye_samurai_ep_ratings.csv"))

blue_eye_samurai %>% create_episode_plot("Blue Eye Samurai") %>% interactive_plotly
```

## Six Feet Under

```{r}
six_feet_under = read_csv(here::here("data", "six_feet_under_ep_ratings.csv"))

six_feet_under %>% create_episode_plot("Six Feet Under") %>% interactive_plotly
```

## Hannibal

```{r}
hannibal = read_csv(here::here("data", "hannibal_ep_ratings.csv"))

hannibal %>% create_episode_plot("Hannibal") %>% interactive_plotly
```

## The Wire

```{r}
the_wire = read_csv(here::here("data", "the_wire_ep_ratings.csv"))

the_wire %>% create_episode_plot("The Wire") %>% interactive_plotly
```

## Person of Interest

```{r}
person_of_interest = read_csv(here::here("data", "person_of_interest_ep_ratings.csv"))

person_of_interest %>% create_episode_plot("Person of Interest") %>% interactive_plotly
```

## Chernobyl

```{r}
chernobyl = read_csv(here::here("data", "chernobyl_ep_ratings.csv"))

chernobyl %>% create_episode_plot("Chernobyl") %>% interactive_plotly
```

## Fleabag

```{r}
fleabag = read_csv(here::here("data", "fleabag_ep_ratings.csv"))

fleabag %>% create_episode_plot("Fleabag") %>% interactive_plotly
```

## Parks and Recreation

```{r}
parks_and_rec = read_csv(here::here("data", "parks_and_rec_ep_ratings.csv"))

parks_and_rec %>% create_episode_plot("Parks and Recreation") %>% interactive_plotly
```

## New Girl

```{r}
new_girl = read_csv(here::here("data", "new_girl_ep_ratings.csv"))

new_girl %>% create_episode_plot("New Girl") %>% interactive_plotly
```

## Brooklyn Nine-Nine

```{r}
brooklyn_nine_nine = read_csv(here::here("data", "brooklyn_nine_nine_ep_ratings.csv"))

brooklyn_nine_nine %>% create_episode_plot("Brooklyn Nine-Nine") %>% interactive_plotly
```

## Death Note

```{r fig.height=12}
death_note = read_csv("data/death_note_ep_ratings.csv")

death_note %>% create_episode_plot("Death Note") %>% interactive_plotly
```

## Andor

```{r}
andor = read_csv("data/andor_ep_ratings.csv")

andor %>% create_episode_plot("Andor") %>% interactive_plotly
```

## Modern Love

```{r}
modern_love = read_csv("data/modern_love_ep_ratings.csv")

modern_love %>% create_episode_plot("Modern Love") %>% interactive_plotly
```

## My Hero Academia

```{r}
my_hero_academia = read_csv("data/my_hero_academia_ep_ratings.csv")

my_hero_academia %>% create_episode_plot("My Hero Academia") %>% interactive_plotly
```

[Raw pngs found here](https://github.com/cavandonohoe/cavandonohoe.github.io/tree/main/imdb_rating_plot_files)
