---
title: "Airplane Boarding Simulation"
output:
  html_document:
    includes:
      in_header: header/header.html
      after_body: include_footer.html
lang: "en"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "asis")
```

How long does it really take to board a plane? It depends a lot on the order in which passengers walk down the aisle. This page simulates five popular boarding methods on a 30-row, single-aisle cabin (think 737 / A320) and compares how each one performs.

Pick a method below and hit **Start**. To race all five at once and see a ranked dashboard, scroll down to the **Comparison Mode** section.

```{r boarding-shell, echo=FALSE}
htmltools::HTML(readr::read_file("scripts/airplane_boarding/airplane_boarding.html"))
```

## How the methods work

- **Steffen Method** &mdash; window seats first, every other row, back-to-front, then middles, then aisles. The staggered geometry lets many people stow luggage simultaneously.
- **WILMA (Window-Middle-Aisle)** &mdash; all windows first (rear-to-front), then all middles, then all aisles. Eliminates seat-swap interference.
- **Back-to-Front** &mdash; classic zone-by-zone boarding from the rear forward. The natural intuition&hellip; that produces the worst aisle congestion.
- **Reverse Pyramid** &mdash; diagonal waves from the rear-window corner toward the front-aisle corner. Aisle seats board last.
- **Random** &mdash; same seat assignments, shuffled boarding order. Often beats Back-to-Front because passengers spread out naturally.

The simulation models walking speed, luggage-stowing time (2&ndash;5 seconds), and aisle blocking when an upstream passenger is in the way. Blocked passengers turn yellow.
