---
title: "AVCA Men's Volleyball First-Team All-Americans"
editor_options:
  chunk_output_type: console
output:
  html_document:
    includes:
      in_header: header/header.html
      after_body: include_footer.html
---

```{r pipe, include=FALSE}
`%>%` <- magrittr::`%>%`
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  results = "asis"
)
```

```{r data}
all_americans <- readr::read_csv(here::here("data", "avca_first_team_all_americans.csv")) %>%
  dplyr::mutate(school = stringr::str_replace_all(school, "\u2018|\u2019", "'"))

champions <- readr::read_csv(here::here("data", "avca_champions.csv"))
```

The [American Volleyball Coaches Association (AVCA)](https://www.avca.org/) selects All-America teams for National Collegiate (Division I-II) men's volleyball each year. This page tracks every **first-team** selection from `r min(all_americans$year)` to `r max(all_americans$year)` --- `r nrow(all_americans)` players across `r length(unique(all_americans$year))` seasons.

[Raw Data](data/avca_first_team_all_americans.csv)
[(Data Update Script)](https://github.com/cavandonohoe/cavandonohoe.github.io/blob/main/web_scraping/avca_all_americans.R)

## All First-Team Selections (`r min(all_americans$year)`--`r max(all_americans$year)`)

```{r all_table}
all_americans %>%
  dplyr::select(Year = year, Name = name, School = school,
                Position = position, Class = class) %>%
  dplyr::mutate(Position = factor(Position), Class = factor(Class)) %>%
  DT::datatable(filter = "top", extensions = "Buttons",
            options = list(paging = FALSE, dom = "Bfrtip", buttons = c("copy", "csv", "excel"),
                           scrollY = "400px", scrollX = TRUE))
```

## Multi-Year First-Team Selections

Players who made the first team more than once during this window.

```{r repeat_table}
repeats <- all_americans %>%
  dplyr::group_by(name) %>%
  dplyr::filter(dplyr::n() > 1) %>%
  dplyr::summarise(
    times = dplyr::n(),
    years = paste(year, collapse = ", "),
    schools = paste(unique(school), collapse = "; "),
    positions = paste(unique(position), collapse = "/"),
    .groups = "drop"
  ) %>%
  dplyr::arrange(dplyr::desc(times), name)

repeats %>%
  dplyr::select(Name = name, `Times Selected` = times, Years = years,
                School = schools, Position = positions) %>%
  DT::datatable(filter = "top", extensions = "Buttons",
            options = list(paging = FALSE, dom = "Bfrtip", buttons = c("copy", "csv", "excel"),
                           scrollY = "400px", scrollX = TRUE))
```

## Selections by School

```{r school_table}
all_americans %>%
  dplyr::count(school, name = "selections") %>%
  dplyr::arrange(dplyr::desc(selections)) %>%
  dplyr::rename(School = school, `First-Team Selections` = selections) %>%
  DT::datatable(filter = "top", extensions = "Buttons",
            options = list(paging = FALSE, dom = "Bfrtip", buttons = c("copy", "csv", "excel"),
                           scrollY = "400px", scrollX = TRUE))
```

## Did the National Champion Have a First-Team All-American Setter?

The setter runs the offense. How often does the national champion have their setter recognized as one of the best in the country?

```{r champ_setter}
setter_positions <- c("S", "S/H")

first_team_setters <- all_americans %>%
  dplyr::filter(position %in% setter_positions) %>%
  dplyr::select(year, setter = name, school, position)

champ_setter <- champions %>%
  dplyr::left_join(first_team_setters, by = c("year", "champion" = "school")) %>%
  dplyr::mutate(
    had_first_team_setter = dplyr::case_when(
      is.na(champion) ~ NA_character_,
      !is.na(setter) ~ paste0("\u2705 Yes (", setter, ")"),
      TRUE ~ "\u274C No"
    )
  )

all_setters_by_year <- first_team_setters %>%
  dplyr::group_by(year) %>%
  dplyr::summarise(
    all_first_team_setters = paste0(setter, " (", school, ")", collapse = ", "),
    .groups = "drop"
  )

champ_display <- champ_setter %>%
  dplyr::left_join(all_setters_by_year, by = "year") %>%
  dplyr::select(
    Year = year,
    Champion = champion_short,
    `Champion Had 1st-Team AA Setter?` = had_first_team_setter,
    `First-Team AA Setters That Year` = all_first_team_setters
  )

champ_display %>%
  DT::datatable(filter = "top", extensions = "Buttons",
            escape = FALSE,
            options = list(paging = FALSE, dom = "Bfrtip", buttons = c("copy", "csv", "excel"),
                           scrollY = "400px", scrollX = TRUE))
```

```{r champ_setter_summary, results='asis'}
valid_years <- champ_setter %>% dplyr::filter(!is.na(champion))
n_with <- sum(!is.na(valid_years$setter))
n_total <- nrow(valid_years)
cat(sprintf(
  "\n\n**Result:** In %d of %d championship seasons (%.0f%%), the national champion had a first-team All-American setter.\n\n",
  n_with, n_total, 100 * n_with / n_total
))
```

---

**Sources:** [AVCA All-America Teams by Year (1991--2024)](https://www.avca.org/wp-content/uploads/2024/05/AVCA-NCAA-National-Collegiate-MVB-All-America-Teams-Year-by-Year.xlsx) and [2025 AVCA All-Americans](https://www.avca.org/award/2025-avca-national-collegiate-mens-all-america-teams/).
