---
title: "AVCA Men's Volleyball First-Team All-Americans"
editor_options:
  chunk_output_type: console
output:
  html_document:
    includes:
      in_header: header/header.html
      after_body: include_footer.html
---

```{r pipe, include=FALSE}
`%>%` <- magrittr::`%>%`
```

The [American Volleyball Coaches Association (AVCA)](https://www.avca.org/) selects All-America teams for National Collegiate (Division I-II) men's volleyball each year. This page tracks every **first-team** selection from 2016 to 2025 --- 108 players across 10 seasons.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  results = "asis"
)
```

```{r data}
all_americans <- readr::read_csv(here::here("data", "avca_first_team_all_americans.csv"))
```

[Raw Data](data/avca_first_team_all_americans.csv)

## All First-Team Selections (2016--2025)

```{r all_table}
all_americans %>%
  dplyr::select(Year = year, Name = name, School = school,
                Position = position, Class = class) %>%
  dplyr::mutate(Position = factor(Position), Class = factor(Class)) %>%
  DT::datatable(filter = "top", extensions = "Buttons",
            options = list(paging = FALSE, dom = "Bfrtip", buttons = c("copy", "csv", "excel"),
                           scrollY = "400px", scrollX = TRUE))
```

## Multi-Year First-Team Selections

Players who made the first team more than once during this window.

```{r repeat_table}
repeats <- all_americans %>%
  dplyr::group_by(name) %>%
  dplyr::filter(dplyr::n() > 1) %>%
  dplyr::summarise(
    times = dplyr::n(),
    years = paste(year, collapse = ", "),
    schools = paste(unique(school), collapse = "; "),
    positions = paste(unique(position), collapse = "/"),
    .groups = "drop"
  ) %>%
  dplyr::arrange(dplyr::desc(times), name)

repeats %>%
  dplyr::select(Name = name, `Times Selected` = times, Years = years,
                School = schools, Position = positions) %>%
  DT::datatable(filter = "top", extensions = "Buttons",
            options = list(paging = FALSE, dom = "Bfrtip", buttons = c("copy", "csv", "excel"),
                           scrollY = "400px", scrollX = TRUE))
```

## Selections by School

```{r school_table}
all_americans %>%
  dplyr::count(school, name = "selections") %>%
  dplyr::arrange(dplyr::desc(selections)) %>%
  dplyr::rename(School = school, `First-Team Selections` = selections) %>%
  DT::datatable(filter = "top", extensions = "Buttons",
            options = list(paging = FALSE, dom = "Bfrtip", buttons = c("copy", "csv", "excel"),
                           scrollY = "400px", scrollX = TRUE))
```

---

**Sources:** [AVCA All-America Teams by Year (1991--2024)](https://www.avca.org/wp-content/uploads/2024/05/AVCA-NCAA-National-Collegiate-MVB-All-America-Teams-Year-by-Year.xlsx) and [2025 AVCA All-Americans](https://www.avca.org/award/2025-avca-national-collegiate-mens-all-america-teams/).
