---
title: "Lesson 2: Basic Operations in R"
author: "Cavan Donohoe"
description: "Variables, data types, vectors, comparisons, control flow, and writing your first function."
output:
  html_document:
    toc: true
    toc_float: true
    includes:
      in_header: header/header.html
      after_body: include_footer.html
lang: "en"
---

<link rel="stylesheet" href="/css/learn-r.css">

<div class="course-progress">
Lesson 2 of 7 · <a href="learn_r.html">Course overview</a>
</div>

```{r setup, include=FALSE}
options(scipen = 999)
knitr::opts_chunk$set(comment = "#>")
```

# Basic Operations in R

This lesson covers the building blocks: variables, data types, vectors, operators, control flow, and functions. None of it is exciting on its own, but every single thing in R is built out of these pieces. Get comfortable here and the rest of the course is mostly pattern-matching.

## Variables and assignment

A **variable** is a name that points to a value. Assign with `<-`:

```{r}
x <- 5
y <- 12
x + y
```

`=` works too, but `<-` is the convention in R for assignment. Stick with it.

You can reassign a variable as often as you like — R doesn't care:

```{r}
x <- 5
x <- x + 1
x
```

## Data types

R has a small set of basic types you'll use constantly.

```{r}
n  <- 3.14         # numeric (double)
i  <- 42L          # integer (the L makes it an integer, not a double)
s  <- "hello"      # character (string)
b  <- TRUE         # logical (TRUE / FALSE)
m  <- NA           # missing value
```

To check what type something is, use `class()` or one of the `is.*` family:

```{r}
class(n)
class(s)
class(b)
is.numeric(n)
is.character(s)
```

Missing values (`NA`) are everywhere in real data. R takes them seriously — most functions return `NA` if their input has any `NA`s, unless you tell them otherwise:

```{r}
mean(c(1, 2, NA))
mean(c(1, 2, NA), na.rm = TRUE)
```

<div class="callout callout-warning">
<div class="callout-title">⚠️ <code>NA</code> is contagious</div>
`NA == NA` is `NA`, not `TRUE`. To check for missingness use `is.na(x)`. This trips up beginners constantly.
</div>

## Arithmetic and math

The usual suspects:

```{r}
10 + 3
10 - 3
10 * 3
10 / 3
10 %% 3   # remainder (modulo)
10 %/% 3  # integer division
10 ^ 3    # exponentiation
```

Plus a stack of built-in math functions:

```{r}
sqrt(16)
log(100)        # natural log
log(100, base = 10)
exp(1)          # e
abs(-7)
round(3.14159, 2)
ceiling(2.1)
floor(2.9)
```

## Comparison and logical operators

Comparisons return `TRUE` or `FALSE`:

```{r}
5 > 3
5 == 5     # equality is double-equals; single-equals is assignment
5 != 6
5 >= 5
"a" < "b"  # alphabetical
```

Logical operators combine `TRUE`/`FALSE` values:

```{r}
TRUE & FALSE   # AND
TRUE | FALSE   # OR
!TRUE          # NOT
```

The single-character versions (`&`, `|`) work elementwise on vectors. The double-character versions (`&&`, `||`) only look at the first element and short-circuit — use them inside `if` statements, not for filtering data.

## Vectors

A **vector** is an ordered collection of values, all of the same type. It's the most fundamental data structure in R — even a single number is technically a vector of length 1.

Create a vector with `c()` ("combine"):

```{r}
ages <- c(23, 31, 27, 45, 19)
fruit <- c("apple", "banana", "cherry")
flags <- c(TRUE, FALSE, TRUE, TRUE)
```

A few handy ways to create vectors:

```{r}
1:5                  # integer sequence
seq(0, 1, by = 0.25)
rep("a", times = 3)
```

### Indexing vectors

Subsetting a vector by position uses `[ ]`. **R is 1-indexed** — the first element is at position 1, not 0.

```{r}
ages[1]            # first element
ages[c(1, 3)]      # first and third
ages[-1]           # everything except the first (negative indexing drops)
ages[ages > 25]    # elements that match a condition
```

That last one — subsetting by a logical vector — is one of the most-used patterns in R. The expression `ages > 25` returns a logical vector the same length as `ages`, and `ages[...]` keeps the elements where it's `TRUE`.

### Vectorized operations

R operations work on entire vectors at once. You almost never need a loop for elementwise math:

```{r}
ages
ages + 1            # add 1 to every element
ages * 2
ages > 30
sum(ages)
mean(ages)
length(ages)
```

This is the single biggest stylistic difference between R and most other languages, and getting comfortable with it is half the battle.

### Named vectors

Vectors can have names attached to each element:

```{r}
prices <- c(apple = 1.20, banana = 0.50, cherry = 3.00)
prices
prices["banana"]
```

## Lists

A **list** is like a vector, but each element can be a different type — even another list.

```{r}
person <- list(name = "Ada", age = 36, hobbies = c("math", "logic"))
person$name
person$hobbies
person[["age"]]
```

You'll meet lists most often as the return value of model-fitting functions like `lm()` (we'll get there in Lesson 5).

## Data frames

A **data frame** is a table — rows are observations, columns are variables, and each column is a vector of the same length. It's the workhorse data structure for most analysis.

```{r}
df <- data.frame(
  name = c("Alice", "Bob", "Charlie"),
  age = c(25, 30, 22),
  score = c(95, 88, 75)
)
df
```

Subset by column with `$` or `[[ ]]`, and by row/column with `[ , ]`:

```{r}
df$age
df[1, ]            # first row
df[, "name"]       # name column
df[df$age > 23, ]  # all rows where age > 23
```

We'll spend most of Lesson 3 working with data frames using a much nicer toolset (`dplyr`). For now, just know they exist and look like a spreadsheet.

## Control flow

### `if` / `else`

```{r}
score <- 78
if (score >= 90) {
  "A"
} else if (score >= 80) {
  "B"
} else if (score >= 70) {
  "C"
} else {
  "F"
}
```

### `for` loops

```{r}
for (n in 1:5) {
  print(n^2)
}
```

<div class="callout callout-tip">
<div class="callout-title">💡 You'll write fewer loops than you think</div>
In R, most things you'd loop over in another language can be done with vectorized operations or with `sapply`/`lapply`. Loops aren't wrong — they're just often unnecessary. We'll see this in Lesson 3.
</div>

### `while` loops

```{r}
n <- 1
while (n < 100) {
  n <- n * 2
}
n
```

## Functions

Writing your own functions is how you go from "running examples" to "actually using R for something." Syntax:

```{r}
celsius_to_fahrenheit <- function(c) {
  c * 9 / 5 + 32
}

celsius_to_fahrenheit(0)
celsius_to_fahrenheit(c(0, 20, 100))   # works on a vector for free
```

A few rules:

- The last expression in the body is what the function returns. You can also write `return(...)` explicitly, but it's not required.
- Arguments can have default values: `function(c, offset = 0) { ... }`.
- Functions are first-class — you can pass them around as values.

A slightly more useful one — given a numeric vector, compute summary stats:

```{r}
quick_summary <- function(x, na_rm = TRUE) {
  list(
    n = length(x),
    n_missing = sum(is.na(x)),
    mean = mean(x, na.rm = na_rm),
    sd = sd(x, na.rm = na_rm),
    min = min(x, na.rm = na_rm),
    max = max(x, na.rm = na_rm)
  )
}

quick_summary(c(2, 4, 6, 8, 10, NA))
```

## Putting it together

Let's combine everything in this lesson into a tiny analysis. Given a vector of test scores, classify each into a letter grade and report how many of each.

```{r}
scores <- c(72, 88, 95, 67, 81, 59, 91, 78, 84, 100)

grade <- function(score) {
  if (score >= 90) {
    "A"
  } else if (score >= 80) {
    "B"
  } else if (score >= 70) {
    "C"
  } else if (score >= 60) {
    "D"
  } else {
    "F"
  }
}

grades <- sapply(scores, grade)
grades

table(grades)
```

`sapply` applies a function to each element of a vector and returns the results. We need it here because `grade()` uses `if`, which only works on a single value at a time. (In Lesson 3 we'll see a much cleaner way to do this with `dplyr::case_when()`.)

<div class="exercise">
<div class="exercise-title">✏️ Exercise 2.1 — Vectorized math</div>

Given the temperatures vector below (in Fahrenheit), convert all of them to Celsius without writing a loop.

```{r eval=FALSE}
temps_f <- c(32, 50, 68, 86, 104)
```

<details>
<summary>Show solution</summary>

```{r}
temps_f <- c(32, 50, 68, 86, 104)
temps_c <- (temps_f - 32) * 5 / 9
temps_c
```

Or wrap it in a function so you can reuse it:

```{r}
fahrenheit_to_celsius <- function(f) (f - 32) * 5 / 9
fahrenheit_to_celsius(temps_f)
```

</details>
</div>

<div class="exercise">
<div class="exercise-title">✏️ Exercise 2.2 — Subsetting</div>

Using the built-in `mtcars` dataset:

1. Print the `mpg` column.
2. Print only the cars with `mpg > 25`.
3. What's the mean `mpg` of cars with 4 cylinders (`cyl == 4`)?

<details>
<summary>Show solution</summary>

```{r}
mtcars$mpg

mtcars[mtcars$mpg > 25, ]

mean(mtcars$mpg[mtcars$cyl == 4])
```

</details>
</div>

<div class="exercise">
<div class="exercise-title">✏️ Exercise 2.3 — Write a function</div>

Write a function `is_leap_year(year)` that returns `TRUE` if `year` is a leap year and `FALSE` otherwise. (A year is a leap year if it's divisible by 4, *except* century years are only leap years if also divisible by 400. So 2000 is, 1900 isn't.)

Test it on `c(2000, 1900, 2020, 2023, 2024)`.

<details>
<summary>Show solution</summary>

```{r}
is_leap_year <- function(year) {
  (year %% 4 == 0 & year %% 100 != 0) | (year %% 400 == 0)
}

is_leap_year(c(2000, 1900, 2020, 2023, 2024))
```

Note we used `&` and `|` (single character) so it works on a whole vector at once — exactly the vectorization point from earlier.

</details>
</div>

## What's next

You now have variables, types, vectors, control flow, and functions. That's enough to write real code. Lesson 3 introduces the **tidyverse** — a much nicer way to manipulate data frames than what you saw above.

<div class="lesson-nav">
  <a class="nav-prev" href="intro_to_r.html">
    <span class="lesson-nav-label">← Previous</span>
    <span class="lesson-nav-title">Lesson 1: Introduction to R</span>
  </a>
  <a class="nav-next" href="data_manipulation_r.html">
    <span class="lesson-nav-label">Next →</span>
    <span class="lesson-nav-title">Lesson 3: Data Manipulation</span>
  </a>
</div>
