---
title: "When Were Confederate Statues Created?"
editor_options: 
  chunk_output_type: console
output:
  html_document:
    includes:
      in_header: header/header.html
      after_body: include_footer.html
---

```{r pipe, include=FALSE}
`%>%` <- magrittr::`%>%`
```


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```

# When were Confederate Statues made?


R Project for web scraping Wikipedia looking for dates when confederate statues were erected


```{r prep, include=FALSE}
# Set default chunk output
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  tidy = FALSE, results = "asis"
)
```

This is so wild. Let this sink in.

This was motivated by my recent trip to South Carolina and seeing my first confederate statue. I am not sure why we glorify these people who fought for people to be enslaved.

Check these out btw:

* https://www.youtube.com/watch?v=dOkFXPblLpU
  - how the United Daughters of the Confederacy brainwashed the youth by pretending slavery wasn't the cause of the Civil War
* https://www.youtube.com/watch?v=WClgR6Q0aPE
  - the reason why each peek of confederate statues were created
* https://www.theguardian.com/us-news/2017/aug/16/confederate-monuments-civil-war-history-trump
  - how each peek and why these statues were created


```{r}
all_dates <- readr::read_csv(paste0(
  "https://raw.githubusercontent.com/cavandonohoe/",
  "confederate_statues/main/data/confederate_statue_dates.csv"
))

# find mode
max_statues <- all_dates %>%
  dplyr::group_by(value3) %>%
  
  dplyr::count() %>%
    
  dplyr::ungroup() %>%
    
  dplyr::arrange(dplyr::desc(n)) %>%
    
  dplyr::filter(n == max(n))

mode <- max_statues %>%
  
  dplyr::pull(value3)

with_max <- max_statues %>%
  
  dplyr::pull(n)
```

Also, we witness the most statues being built in a single year in `r mode` with `r with_max` statues.

These bins are of width = 5 years.

```{r graph1, fig.width=10, fig.height=7.5}
all_dates_clean <- all_dates %>%
  dplyr::group_by(value3) %>%
  
  dplyr::count() %>%
    
  dplyr::ungroup() %>%
    
  dplyr::mutate(date_bins = cut(value3, breaks = seq(min(value3), max(value3)+5, by = 5),
    right = FALSE),
         text = paste0(value3, ": ", n)) %>%
           
  dplyr::group_by(date_bins) %>%
    
  dplyr::summarise(n = sum(n),
            text = paste(text, collapse = "\n")) %>%
              
  dplyr::ungroup() %>%
    
  dplyr::mutate(date_min = as.numeric(gsub("\\[(\\d+),.*", "\\1", date_bins))) %>%
    
  dplyr::rename(date = date_min)

confederate_statue_hist <- all_dates_clean %>%
  
  ggplot2::ggplot(ggplot2::aes(x = date, text = text, y = n)) +
  ggplot2::geom_col(fill = "red") +
  ggplot2::scale_x_continuous(breaks = seq(min(all_dates_clean$date), max(all_dates_clean$date),
    by = 5)) +
  ggplot2::theme(axis.title.x = ggplot2::element_blank(),
        axis.text.x = ggplot2::element_text(size=12, angle = 90),
        axis.text.y = ggplot2::element_text(size=15),
        title = ggplot2::element_text(size=15)) +
  ggplot2::ylab("# of Statues") +
  ggplot2::ggtitle("Number of Confederate Statues Over Time")

confederate_statue_hist %>%
  plotly::ggplotly()
```

It's just something to think about...

Also web scraping is fun



sources:  
https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_Alabama
https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_Georgia
https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_Mississippi
https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_North_Carolina
https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_South_Carolina
https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials



To view web scraping methodology, please view: https://github.com/cavandonohoe/confederate_statues
