---
title: "Covid Cases: Then & Now"
date: '`r format(Sys.time(), "%Y-%m-%d %X")`'
editor_options: 
  chunk_output_type: console
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
    includes:
      in_header: header/header.html
---

```{r setup, include=FALSE}
# Set default chunk output
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  results = "asis"
)
```

```{r libraries}
library(tidyverse)
library(DT)
library(scales)
library(plotly)
library(zoo)
library(lubridate)
library(ggrepel)

```



```{r grab_data}
covid.cases = read_csv(paste0("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/",
                              "csse_covid_19_data/csse_covid_19_time_series/",
                              "time_series_covid19_confirmed_global.csv"))
```

## United States

```{r united_states_data}
covid.cases_v2 = covid.cases %>%
  gather(key="Date", value="confirmed", -c(`Country/Region`, `Province/State`, Lat, Long)) %>% 
  mutate(Date = as.Date(Date, format="%m/%d/%y"))


covid.cases_v3 = covid.cases_v2 %>% filter(grepl(x=`Country/Region`, pattern= "US")) %>%
  mutate(confirmed.previous.day = lag(confirmed)) %>% 
  mutate(New.Cases = confirmed - confirmed.previous.day)

covid.cases_v4 = covid.cases_v3 %>%
  mutate(Seven.Day.Average = rollmean(New.Cases, 7, na.pad = TRUE, align = "right"))


covid.cases_v5 = covid.cases_v4 %>% filter(!is.na(Seven.Day.Average)) %>% 
  mutate(Seven.Day.Average_lag = lag(Seven.Day.Average)) %>%
  mutate(Change = Seven.Day.Average - Seven.Day.Average_lag) %>% 
  mutate(Change = ifelse(Change >= 0, paste0("+", round(Change, 0)), 
                         paste0(round(Change, 0))))
```

Please note I only have data until `r max(covid.cases_v2$Date)`. The rolling average is the mean of that day's cases as well as the 6 days prior. 

```{r us_plots}
us_cases_plot = covid.cases_v5 %>% ggplot(aes(x=Date, y=Seven.Day.Average)) +
  geom_line() + ggtitle("Seven Day Average Across the US") + 
  scale_y_continuous(labels = scales::comma) +
  scale_x_date(date_labels = "%b %y", breaks = pretty_breaks(n=7))+
  geom_hline(yintercept = covid.cases_v5$Seven.Day.Average[covid.cases_v5$Date ==
                                                             max(covid.cases_v5$Date)],
             linetype=2, colour="red")+
  ylab("Seven Day Average") +
  theme(axis.title.x=element_blank())

us_cases_plot %>% ggplotly()

```

The red dashed line helps us visualize where our cases used to be. The following table provides the actual days and seven day average cases of those intersections.

```{r intersection_points_us}
covid.cases_v5 %>% 
  mutate(Intersection = Seven.Day.Average >= 
           covid.cases_v5$Seven.Day.Average[covid.cases_v5$Date == 
                                              max(covid.cases_v5$Date, na.rm = TRUE)]) %>% 
  mutate(Intersection_lead = lead(Intersection)) %>% 
  filter(Intersection_lead != Intersection | Date == max(Date)) %>% 
  select(-Lat, -Long) %>% 
  select(Date, seven_day_average = Seven.Day.Average) %>% 
  mutate(seven_day_average = round(seven_day_average)) %>% 
  datatable(filter = "top", extensions = "Buttons",
            options = list(paging = FALSE, dom = "Bfrtip", buttons = c("copy", "csv", "excel"),
                           scrollY = "400px", scrollX = TRUE))

```

## California

I like to check on my state as well.

```{r  grab_data_california}
covid.cases_state = read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv")

```

```{r clean_ca_data}
covid.cases_state_v2 = covid.cases_state %>% select(date, state, cases) %>% arrange(date) %>% 
  rename(Date = date, confirmed = cases, NAME = state) %>% 
  left_join(tibble(NAME = state.name, state = state.abb))
  

covid.cases_state_v3 = covid.cases_state_v2 %>% filter(state == "CA") %>%
  mutate(confirmed = replace_na(confirmed, 0)) %>% 
  mutate(confirmed.previous.day = lag(confirmed)) %>% 
  mutate(New.Cases = confirmed - confirmed.previous.day)

covid.cases_state_v4 = covid.cases_state_v3 %>%
  mutate(Seven.Day.Average = rollmean(New.Cases, 7, na.pad = TRUE, align = "right"))


covid.cases_state_v5 = covid.cases_state_v4 %>% filter(!is.na(Seven.Day.Average)) %>% 
  mutate(Seven.Day.Average_lag = lag(Seven.Day.Average)) %>%
  mutate(Change = Seven.Day.Average - Seven.Day.Average_lag) %>% 
  mutate(Change_format = ifelse(Change >= 0, paste0("+", round(Change, 0)), 
                         paste0(round(Change, 0))))
```


```{r ca_plot}
ca_plot = covid.cases_state_v5 %>% ggplot(aes(x=Date, y=Seven.Day.Average, label=Change_format)) + geom_line() + 
  ggtitle(paste("New", unique(covid.cases_state_v5$NAME), "Covid Cases")) + 
  scale_y_continuous(labels = scales::comma) +
  scale_x_date(breaks = pretty_breaks(7), date_labels = "%b %y")+
  geom_hline(yintercept = covid.cases_state_v5$Seven.Day.Average[covid.cases_state_v5$Date ==
                                                                   max(covid.cases_state_v5$Date)],
             linetype=2, colour="red") +
  ylab("Seven Day Average")

ca_plot %>% ggplotly()
```

Just like the US chart, the red dashed line can help us visualize the last times we were at today's COVID cases. Below is a table of the exact dates of intersection between today's cases vs previous cases. 

```{r ca_intersections}
covid.cases_state_v5 %>% 
  mutate(Intersection = Seven.Day.Average >= 
           covid.cases_state_v5$Seven.Day.Average[covid.cases_state_v5$Date == 
                                              max(covid.cases_state_v5$Date, na.rm = TRUE)]) %>% 
  mutate(Intersection_lead = lead(Intersection)) %>% 
  filter(Intersection_lead != Intersection | Date == max(Date)) %>% 
  select(Date, seven_day_average = Seven.Day.Average) %>% 
  mutate(seven_day_average = round(seven_day_average)) %>% 
  datatable(filter = "top", extensions = "Buttons",
            options = list(paging = FALSE, dom = "Bfrtip", buttons = c("copy", "csv", "excel"),
                           scrollY = "400px", scrollX = TRUE))

```

Also, the way the intersection table works is if today's 7 Day Average Cases = 10000, and we had 9,982 cases on August 2nd, 2021 and 10,190 cases on August 3rd, 2021, then August 2nd, 2021 would be listed in the table.

## State vs State

```{r grab_statewide_data}
# luckily we already have this in the script
# covid.cases_state = read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv")

# need to normalize everything to the same standard
state_pop = read_csv("http://www2.census.gov/programs-surveys/popest/datasets/2010-2019/national/totals/nst-est2019-alldata.csv")

```

```{r clean_statewide_data}
state_pop_v2 = state_pop %>% select(NAME, POPESTIMATE2019) %>%
  left_join(tibble(NAME = state.name, state = state.abb)) %>% 
  rename(population = POPESTIMATE2019)

covid.cases_state_v2 = covid.cases_state %>%
  select(date, state, cases, deaths) %>% arrange(date) %>% 
  rename(Date = date, confirmed = cases, NAME = state) %>% left_join(state_pop_v2) %>%
  select(Date, state, confirmed, population, deaths) %>%
  arrange(Date) %>% 
  mutate(confirmed = confirmed / population,
         death_per_cap = deaths / population) %>% 
  filter(state %in% state.abb)


covid.cases_state_v3 = covid.cases_state_v2 %>% group_by(state) %>%
  mutate(confirmed = replace_na(confirmed, 0),
         death_per_cap = replace_na(death_per_cap, 0)) %>% 
  mutate(confirmed.previous.day = lag(confirmed),
         death_per_cap_previous_day = lag(death_per_cap)) %>% 
  mutate(New.Cases = confirmed - confirmed.previous.day,
         new_deaths = death_per_cap - death_per_cap_previous_day)

covid.cases_state_v4 = covid.cases_state_v3 %>%
  mutate(Seven.Day.Average = rollmean(New.Cases, 7, na.pad = TRUE, align = "right"),
         seven_day_average_deaths = rollmean(new_deaths, 7, na.pad = TRUE, align = "right")) %>% 
  ungroup()


covid.cases_state_v5 = covid.cases_state_v4 %>%
  filter(!is.na(Seven.Day.Average))


```


```{r state_by_state_plot}
covid.cases_state_plot = covid.cases_state_v5 %>% ggplot(aes(x=Date, y=Seven.Day.Average, colour = state)) + geom_line() + 
  ggtitle("New Covid Cases per Capita") + 
  scale_y_continuous(labels = scales::comma) + scale_x_date(breaks = pretty_breaks(7), date_labels = "%b %y") +
  ylab("Seven Day Average per Capita")

covid.cases_state_plot %>% ggplotly()

```

Yes, this is noisy, but you have the option to double click any number of states on the right to isolate. 

## Vaccines

Now, I'm no public health expert, but it does seem like the higher vaccine rate a state has, the lower the COVID cases are. Let's just let the data talk.

```{r vaccine_data}
vaccines = read_csv("https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/us_state_vaccinations.csv")

state_tib = tibble(location = state.name, state = state.abb)

vaccines_current_rate = vaccines %>%
  mutate(location = ifelse(location == "New York State", "New York", location)) %>% 
  filter(date == max(date)) %>% 
  filter(location %in% state.name) %>% 
  select(location, people_fully_vaccinated_per_hundred) %>% 
  arrange(desc(people_fully_vaccinated_per_hundred))

covid_current = covid.cases_state_v5 %>% ungroup() %>% filter(Date == max(Date, na.rm = TRUE)) %>% 
  arrange(Seven.Day.Average) %>%
  mutate(new_cases_actual = New.Cases * population) %>% 
  select(state, Seven.Day.Average) %>% 
  left_join(state_tib)

covid_cases_w_vaccines = covid_current %>% full_join(vaccines_current_rate)

vaccines_historic_rate = vaccines %>%
  mutate(location = ifelse(location == "New York State", "New York", location)) %>% 
  filter(location %in% state.name) %>% 
  select(date, location, people_fully_vaccinated_per_hundred, people_fully_vaccinated) %>% 
  arrange(desc(people_fully_vaccinated_per_hundred))

```



```{r vaccines_by_state}

vaccines_historic_rate_plot = vaccines_historic_rate %>%
  ggplot(aes(x=date, y= people_fully_vaccinated_per_hundred, colour = location)) +
  geom_point() +
  scale_x_date(date_labels = "%h %d", breaks = scales::pretty_breaks(7)) +
  ylab("% fully vaxxed") +
  ggtitle("vax rate per state")

vaccines_historic_rate_plot %>% ggplotly()

```

Now if that is a little too noisy, I'll break it down between red and blue.

```{r vaccine_party_plot}
usa_pop = state_pop_v2 %>% filter(NAME == "United States") %>% pull(population)

vaccines_usa = vaccines %>%
  mutate(location = ifelse(location == "New York State", "New York", location)) %>% 
  filter(location %in% state.name) %>% 
  group_by(date) %>% 
  summarise(people_fully_vaccinated = sum(people_fully_vaccinated, na.rm = TRUE)) %>% 
  ungroup() %>% mutate(per_fully_vaxxed = people_fully_vaccinated / usa_pop)

# voting preference
voting_pref = read_csv("data/president_1976_2020.csv")

voting_pref2 = voting_pref %>% filter(year == 2020 & party_detailed %in% c("DEMOCRAT", "REPUBLICAN")) %>% 
  mutate(per_votes = candidatevotes / totalvotes) %>%
  group_by(state) %>% 
  filter(candidatevotes == max(candidatevotes)) %>%
  ungroup() %>% 
  select(state, state_po, party_detailed, per_votes) %>% 
  mutate(party_simp = ifelse(party_detailed == "REPUBLICAN", "R", "D"),
         party_color = ifelse(party_simp == "R", "red", "blue"),
         STATE = state,
         location = snakecase::to_any_case(string = tolower(STATE), case = "title"),
         per_votes_sign = ifelse(party_simp == "D", per_votes, -per_votes)) 
  
vaccine_party = vaccines_historic_rate %>% left_join(voting_pref2)


vaccine_party_plot = vaccine_party %>%
  ggplot(aes(x=date, y=people_fully_vaccinated_per_hundred, colour = per_votes_sign,
             label = location)) +
  geom_point() +
  scale_colour_gradient(low = ("red"),
                        high = ("blue")) +
  scale_x_date(date_labels = "%h %d", breaks = pretty_breaks(7)) +
  ylab("% fully vaxxed") +
  ggtitle("vax rate per state per party")

vaccine_party_plot %>% ggplotly()

```

## Deaths since July 2021

```{r deaths}
deaths_since_july = covid.cases_state %>%
  group_by(date) %>% summarise(total_deaths = sum(deaths)) %>% 
  ungroup() %>% 
  mutate(death_change = total_deaths - lag(total_deaths)) %>% 
  mutate(seven_day_average_death = rollmean(death_change, 7, na.pad = TRUE)) %>%
  filter(date >= as.Date("2021-07-01")) %>%
  ggplot(aes(x=date, y=seven_day_average_death)) +
  geom_path()+
  scale_y_continuous(breaks = scales::pretty_breaks(n = 20))+
  ggtitle("Covid Deaths per Day after 7/1/2021")+
  ylab("Seven Day Average Deaths")

deaths_since_july %>% ggplotly()

july2021_deaths = covid.cases_state %>%
  group_by(date) %>% summarise(total_deaths = sum(deaths)) %>% 
  ungroup() %>% 
  mutate(death_change = total_deaths - lag(total_deaths)) %>% 
  mutate(seven_day_average_death = rollmean(death_change, 7, na.pad = TRUE)) %>% 
  filter(date >= as.Date("2021-07-01")) %>% 
  filter(!is.na(seven_day_average_death))

perc_change =
  july2021_deaths$seven_day_average_death[july2021_deaths$date == max(july2021_deaths$date)] /
  min(july2021_deaths$seven_day_average_death,na.rm = TRUE)


```

This is `r round(perc_change, digits = 1)` times the amount we had at the lowest 7 day average death count in July.







