---
title: "Lesson 4: Data Visualization in R"
author: "Cavan Donohoe"
description: "The grammar of graphics with ggplot2: geoms, aesthetics, facets, themes, and saving plots."
output:
  html_document:
    toc: true
    toc_float: true
    includes:
      in_header: header/header.html
      after_body: include_footer.html
lang: "en"
---

<link rel="stylesheet" href="/css/learn-r.css">

<div class="course-progress">
Lesson 4 of 7 · <a href="learn_r.html">Course overview</a>
</div>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.width = 6.5,
  fig.height = 4.2,
  dpi = 100
)
options(scipen = 999)
```

# Data Visualization in R

Plotting in R is excellent. You've already seen `plot()` — it works, but the real toolkit is **ggplot2**. It's based on a "grammar of graphics" that, once it clicks, lets you describe almost any plot in a handful of lines.

This lesson covers the ggplot2 grammar, the most common geoms, faceting, themes, and how to save plots.

## Setup

```{r}
library(ggplot2)
library(dplyr)

cars <- mtcars |>
  tibble::rownames_to_column("model") |>
  tibble::as_tibble() |>
  mutate(
    cyl = factor(cyl),
    transmission = if_else(am == 1, "manual", "automatic")
  )
```

## The ggplot2 grammar

Every ggplot2 plot has three required pieces:

1. **Data** — the data frame you're plotting.
2. **Aesthetics** (`aes()`) — which columns map to which visual properties (x, y, color, size, shape).
3. **Geoms** — the geometric shapes to draw (points, bars, lines, boxes, ...).

The pattern:

```r
ggplot(data, aes(x = ..., y = ..., color = ...)) +
  geom_*()
```

You add layers with `+`. Yes, `+` (not the pipe). Inside a `ggplot()` call you `+`, between data steps you `|>`. It's a mild gotcha and you'll get it wrong a lot at first.

## A first plot

```{r}
ggplot(cars, aes(x = wt, y = mpg)) +
  geom_point()
```

That's the whole plot: data (`cars`), aesthetics (`wt` on x, `mpg` on y), one geom (`geom_point` for a scatterplot).

Now let's add color, size, and labels:

```{r}
ggplot(cars, aes(x = wt, y = mpg, color = cyl, size = hp)) +
  geom_point(alpha = 0.8) +
  labs(
    title = "Heavier, more powerful cars get worse mileage",
    subtitle = "Each point is a 1974 car model",
    x = "Weight (1000 lbs)",
    y = "Miles per gallon",
    color = "Cylinders",
    size = "Horsepower"
  )
```

Notice three things:

- `color = cyl` and `size = hp` are inside `aes()` because they map to data columns.
- `alpha = 0.8` is outside `aes()` because it's a fixed value (every point gets the same alpha), not a mapping.
- `labs()` controls every text label on the plot at once.

## The most useful geoms

### `geom_point()` — scatterplots

Shown above. Add `geom_smooth()` to overlay a regression line:

```{r}
ggplot(cars, aes(x = wt, y = mpg)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE)
```

`method = "lm"` is a straight line; the default `"loess"` is a smooth curve.

### `geom_line()` — connect points by some ordering

```{r}
year_data <- tibble::tibble(
  year = 2015:2024,
  sales = c(100, 108, 121, 119, 130, 145, 160, 158, 172, 185)
)

ggplot(year_data, aes(x = year, y = sales)) +
  geom_line(linewidth = 1.1, color = "steelblue") +
  geom_point(size = 2.5, color = "steelblue") +
  labs(title = "Sales over time", x = NULL, y = "Sales (units)")
```

### `geom_bar()` and `geom_col()` — bar charts

`geom_bar()` counts rows itself. `geom_col()` uses the y-value you give it.

```{r}
ggplot(cars, aes(x = cyl)) +
  geom_bar(fill = "steelblue") +
  labs(title = "How many of each cylinder count?", x = "Cylinders", y = "Count")
```

```{r}
cars |>
  group_by(cyl) |>
  summarise(mean_mpg = mean(mpg), .groups = "drop") |>
  ggplot(aes(x = cyl, y = mean_mpg, fill = cyl)) +
  geom_col() +
  geom_text(aes(label = round(mean_mpg, 1)), vjust = -0.4) +
  labs(title = "Average MPG by cylinder count", x = "Cylinders", y = "Mean MPG") +
  theme(legend.position = "none")
```

Notice we pipe a `dplyr` summary directly into `ggplot()`. This is the standard tidyverse workflow.

### `geom_histogram()` and `geom_density()` — distributions

```{r}
ggplot(cars, aes(x = mpg)) +
  geom_histogram(bins = 12, fill = "steelblue", color = "white") +
  labs(title = "Distribution of MPG across all cars", x = "MPG", y = "Count")
```

```{r}
ggplot(cars, aes(x = mpg, fill = transmission)) +
  geom_density(alpha = 0.5) +
  labs(title = "MPG by transmission type", x = "MPG", y = "Density")
```

### `geom_boxplot()` — distributions by group

```{r}
ggplot(cars, aes(x = cyl, y = mpg, fill = cyl)) +
  geom_boxplot() +
  labs(title = "MPG by cylinder count", x = "Cylinders", y = "MPG") +
  theme(legend.position = "none")
```

## Faceting — small multiples

`facet_wrap()` and `facet_grid()` let you split a plot into a grid of subplots, one per category. Often more informative than mapping a category to color.

```{r}
ggplot(cars, aes(x = wt, y = mpg)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ cyl, labeller = label_both) +
  labs(title = "Weight vs. MPG by cylinder count")
```

`facet_grid()` does a 2D grid of categories:

```{r}
ggplot(cars, aes(x = wt, y = mpg)) +
  geom_point() +
  facet_grid(transmission ~ cyl, labeller = label_both) +
  labs(title = "Weight vs. MPG by transmission and cylinder count")
```

## Scales — controlling axes and color

Scales control how data values map to visual properties. Common ones:

```{r}
ggplot(cars, aes(x = wt, y = mpg, color = hp)) +
  geom_point(size = 3) +
  scale_color_viridis_c(option = "plasma") +
  scale_x_continuous(limits = c(1, 6), breaks = 1:6) +
  scale_y_continuous(name = "Miles per gallon") +
  labs(title = "Weight vs. MPG, colored by horsepower")
```

A few worth knowing:

- `scale_*_continuous()` — for numeric axes; sets limits, breaks, labels.
- `scale_*_log10()` — log axis (useful when data spans orders of magnitude).
- `scale_color_viridis_*()` / `scale_fill_viridis_*()` — perceptually uniform, color-blind-friendly palettes. Use these.
- `scale_color_manual()` — assign specific colors to specific levels.

## Themes — controlling the non-data ink

Themes change how the plot *looks* — gridlines, fonts, backgrounds. ggplot2 ships with several:

```{r}
p <- ggplot(cars, aes(x = wt, y = mpg, color = cyl)) +
  geom_point(size = 3) +
  labs(title = "Weight vs. MPG", x = "Weight", y = "MPG")

p + theme_minimal()
```

```{r}
p + theme_classic()
```

You can also tweak individual theme elements:

```{r}
p +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

Setting a theme for the whole session:

```{r eval=FALSE}
theme_set(theme_minimal(base_size = 13))
```

## Saving plots

`ggsave()` saves the most recent plot — or one you pass to it — as a file:

```{r eval=FALSE}
my_plot <- ggplot(cars, aes(x = wt, y = mpg)) + geom_point()

ggsave("my_plot.png", my_plot, width = 6, height = 4, dpi = 300)
ggsave("my_plot.pdf", my_plot, width = 6, height = 4)
```

`ggsave()` infers the format from the file extension. Width and height are in inches by default.

## Putting it together

Let's build a single, polished figure that summarises something interesting from the data: how MPG depends on weight, broken out by cylinder count, with a fitted line for each group.

```{r fig.height=5}
ggplot(cars, aes(x = wt, y = mpg, color = cyl)) +
  geom_point(size = 3, alpha = 0.85) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.9) +
  scale_color_viridis_d(option = "viridis", end = 0.85) +
  labs(
    title = "Heavier cars get worse mileage — but cylinder count matters too",
    subtitle = "Each point is a 1974 car model from the mtcars dataset",
    x = "Weight (1000 lbs)",
    y = "Miles per gallon",
    color = "Cylinders",
    caption = "Data: Henderson & Velleman (1981). Building multiple regression models interactively."
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "grey40"),
    plot.caption = element_text(color = "grey50", hjust = 0),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

That's a publication-quality plot in about 20 lines.

<div class="exercise">
<div class="exercise-title">✏️ Exercise 4.1 — Your first ggplot</div>

Make a scatterplot of `hp` (x) versus `mpg` (y) from `cars`, with points colored by `cyl`. Add a title.

<details>
<summary>Show solution</summary>

```{r}
ggplot(cars, aes(x = hp, y = mpg, color = cyl)) +
  geom_point(size = 3) +
  labs(title = "Horsepower vs. MPG", x = "Horsepower", y = "MPG", color = "Cylinders")
```

</details>
</div>

<div class="exercise">
<div class="exercise-title">✏️ Exercise 4.2 — Bars from a summary</div>

For each transmission type (`automatic` / `manual`), compute the mean MPG. Plot it as a bar chart with the bar height equal to the mean.

<details>
<summary>Show solution</summary>

```{r}
cars |>
  group_by(transmission) |>
  summarise(mean_mpg = mean(mpg), .groups = "drop") |>
  ggplot(aes(x = transmission, y = mean_mpg, fill = transmission)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = round(mean_mpg, 1)), vjust = -0.4) +
  labs(title = "Mean MPG by transmission type", x = NULL, y = "Mean MPG") +
  theme_minimal() +
  theme(legend.position = "none")
```

</details>
</div>

<div class="exercise">
<div class="exercise-title">✏️ Exercise 4.3 — Faceting</div>

Make a scatterplot of `wt` vs. `mpg`, with one panel per cylinder count, and a linear fit line in each panel.

<details>
<summary>Show solution</summary>

```{r}
ggplot(cars, aes(x = wt, y = mpg)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ cyl, labeller = label_both) +
  theme_minimal()
```

</details>
</div>

## What's next

You can now turn a data frame into a clear, attractive plot. Combined with Lesson 3, you've got a real "explore the data" workflow. Lesson 5 brings in actual statistics — t-tests, correlation, and regression — so you can quantify what your plots are showing you.

<div class="lesson-nav">
  <a class="nav-prev" href="data_manipulation_r.html">
    <span class="lesson-nav-label">← Previous</span>
    <span class="lesson-nav-title">Lesson 3: Data Manipulation</span>
  </a>
  <a class="nav-next" href="statistics_with_r.html">
    <span class="lesson-nav-label">Next →</span>
    <span class="lesson-nav-title">Lesson 5: Statistics with R</span>
  </a>
</div>
