---
title: "Lesson 3: Data Manipulation in R"
author: "Cavan Donohoe"
description: "Read data, then reshape it with the tidyverse: filter, select, mutate, group_by, summarise, and the pipe."
output:
  html_document:
    toc: true
    toc_float: true
    includes:
      in_header: header/header.html
      after_body: include_footer.html
lang: "en"
---

<link rel="stylesheet" href="/css/learn-r.css">

<div class="course-progress">
Lesson 3 of 7 · <a href="learn_r.html">Course overview</a>
</div>

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "#>", message = FALSE, warning = FALSE)
options(scipen = 999)
```

# Data Manipulation in R

Most data analysis is the same six things, over and over: load data, filter rows, pick columns, derive new columns, group, summarise. The **tidyverse** is a family of R packages designed around these operations, and once you learn the vocabulary it becomes second-nature.

In this lesson we'll work with the built-in `mtcars` dataset (32 cars, fuel efficiency stats from a 1974 issue of *Motor Trend*). Tiny, but enough to show every concept.

## Setup

Install the tidyverse if you haven't already:

```{r eval=FALSE}
install.packages("tidyverse")
```

And load it:

```{r}
library(dplyr)
library(tidyr)
library(readr)
```

We're loading three packages: `dplyr` (the manipulation verbs), `tidyr` (reshaping), and `readr` (fast CSV reading). `library(tidyverse)` would load all of these and a few more in one shot.

<div class="callout callout-note">
<div class="callout-title">📝 <code>library()</code> vs <code>::</code></div>
You can also call any function with `package::function()` — e.g. `dplyr::filter(mtcars, mpg > 25)` — without ever calling `library()`. That's what production R code typically does, because it's explicit about where every function comes from. For learning, `library()` is fine and what most tutorials do.
</div>

## Reading data

For most analyses you'll start with a CSV. `readr::read_csv()` reads a CSV into a tibble (a tidyverse-flavored data frame):

```{r eval=FALSE}
my_data <- read_csv("path/to/file.csv")
```

`read_csv()` accepts a URL too, which is great for tutorials and reproducible examples:

```{r eval=FALSE}
penguins <- read_csv(
  "https://raw.githubusercontent.com/allisonhorst/palmerpenguins/main/inst/extdata/penguins.csv"
)
```

For Excel files use `readxl::read_excel()`. For databases use `DBI` plus a driver package. For R's native binary format use `readRDS()` and `saveRDS()`.

For this lesson we'll use the built-in `mtcars`. Let's convert it to a tibble and add a `model` column from the row names:

```{r}
cars <- mtcars |>
  tibble::rownames_to_column("model") |>
  tibble::as_tibble()
cars
```

A **tibble** prints nicer than a base data frame (only the first 10 rows by default, with column types in the header) and is otherwise mostly the same thing.

## The pipe

You just saw `|>`. That's R's **pipe operator**, and it's what makes tidyverse code readable. It takes the thing on its left and feeds it to the function on its right as the first argument:

```{r eval=FALSE}
mean(c(1, 2, 3))
# is equivalent to
c(1, 2, 3) |> mean()
```

By itself that's a wash. The win is when you chain operations:

```{r eval=FALSE}
# without the pipe — read inside-out:
arrange(filter(select(cars, model, mpg, cyl), cyl == 4), desc(mpg))

# with the pipe — read top-to-bottom:
cars |>
  select(model, mpg, cyl) |>
  filter(cyl == 4) |>
  arrange(desc(mpg))
```

The pipe version reads almost like English: "take cars, select these columns, filter to 4-cylinder cars, sort by mpg descending."

<div class="callout callout-tip">
<div class="callout-title">💡 The pipe shortcut</div>
RStudio has a built-in shortcut for the pipe: <kbd>Cmd</kbd>/<kbd>Ctrl</kbd> + <kbd>Shift</kbd> + <kbd>M</kbd>. Use it. You'll type the pipe thousands of times.

You may also see `%>%` in older code — that's the original pipe from the `magrittr` package. They're nearly equivalent; the new built-in `|>` is the modern default.
</div>

## The five core dplyr verbs

These are the verbs you'll use 95% of the time. Each takes a data frame and returns a data frame.

### `filter()` — keep rows

```{r}
cars |> filter(cyl == 6)
cars |> filter(mpg > 25, hp < 100)
cars |> filter(cyl == 4 | cyl == 6)
```

Multiple conditions separated by commas are combined with `&` (AND).

### `select()` — pick columns

```{r}
cars |> select(model, mpg, cyl)
cars |> select(-disp, -hp)              # drop columns
cars |> select(model, mpg:hp)           # range
cars |> select(starts_with("d"))        # by name pattern
```

### `mutate()` — create or change columns

```{r}
cars |>
  mutate(
    kpl = mpg * 0.425,                  # km per liter
    weight_lbs = wt * 1000,
    is_efficient = mpg > 25
  ) |>
  select(model, mpg, kpl, weight_lbs, is_efficient)
```

`mutate()` can use any R function — including ones you write yourself.

### `arrange()` — sort rows

```{r}
cars |> arrange(mpg)              # ascending
cars |> arrange(desc(mpg))        # descending
cars |> arrange(cyl, desc(mpg))   # multi-key
```

### `summarise()` + `group_by()` — collapse rows

`summarise()` reduces a data frame to a single row. By itself it's not that interesting:

```{r}
cars |> summarise(mean_mpg = mean(mpg), n = n())
```

But pair it with `group_by()` and you have a pivot table:

```{r}
cars |>
  group_by(cyl) |>
  summarise(
    n = n(),
    mean_mpg = mean(mpg),
    sd_mpg = sd(mpg),
    .groups = "drop"
  )
```

This is one of the most-used patterns in all of data analysis. "For each group, compute X."

You can group by multiple variables:

```{r}
cars |>
  group_by(cyl, am) |>     # am: 0 = automatic, 1 = manual
  summarise(mean_mpg = mean(mpg), n = n(), .groups = "drop")
```

## A recipe-style example

Suppose you want to know: **for cars with more than 100 horsepower, what's the average MPG by number of cylinders, sorted from worst to best?**

```{r}
cars |>
  filter(hp > 100) |>
  group_by(cyl) |>
  summarise(
    mean_mpg = mean(mpg),
    n = n(),
    .groups = "drop"
  ) |>
  arrange(mean_mpg)
```

That's the whole answer in 6 lines. No loops, no manual indexing, no temp variables. This is why people like the tidyverse.

## A few more useful verbs

### `count()` — quick frequency table

```{r}
cars |> count(cyl)
cars |> count(cyl, sort = TRUE)
cars |> count(cyl, am)
```

### `case_when()` — multi-branch `if`

Inside `mutate()`, `case_when()` is the right tool for "if-elseif-else" logic. Each `~` line is "condition ~ value":

```{r}
cars |>
  mutate(
    mpg_class = case_when(
      mpg < 15  ~ "thirsty",
      mpg < 22  ~ "average",
      mpg < 28  ~ "good",
      TRUE      ~ "great"
    )
  ) |>
  count(mpg_class)
```

The final `TRUE ~ ...` is the catch-all "else."

### `slice_*()` — top/bottom/random rows

```{r}
cars |> slice_max(mpg, n = 3)         # 3 highest mpg
cars |> slice_min(hp, n = 3)          # 3 lowest hp
cars |> slice_sample(n = 3)           # 3 random rows
```

### Joining tables

When you have two related tables, `*_join()` combines them on a shared key:

```{r}
classes <- tibble::tibble(
  cyl = c(4, 6, 8),
  class = c("compact", "midsize", "muscle")
)

cars |>
  select(model, cyl, mpg) |>
  left_join(classes, by = "cyl") |>
  head()
```

`left_join()` keeps every row from the left table and pulls in matching rows from the right. There's also `inner_join()`, `right_join()`, and `full_join()` for different behaviors when keys don't match on both sides.

## Reshaping with `tidyr`

Sometimes data comes in **wide** format (one column per category) when you need it **long** (one row per observation), or vice versa. `tidyr` has `pivot_longer()` and `pivot_wider()` for that.

```{r}
wide <- tibble::tibble(
  city = c("LA", "NYC", "Chicago"),
  jan = c(58, 32, 25),
  jul = c(83, 78, 76)
)
wide

long <- wide |>
  pivot_longer(cols = c(jan, jul), names_to = "month", values_to = "temp_f")
long

long |> pivot_wider(names_from = month, values_from = temp_f)
```

`pivot_longer()` is the one you'll reach for most — most analysis, plotting, and modeling functions expect long data.

## Putting it together

Here's a more realistic mini-analysis. We'll classify cars by efficiency, count how many fall into each class by transmission type, and report it as a clean wide table.

```{r}
cars |>
  mutate(
    transmission = if_else(am == 1, "manual", "automatic"),
    efficiency = case_when(
      mpg < 18 ~ "low",
      mpg < 24 ~ "medium",
      TRUE     ~ "high"
    )
  ) |>
  count(transmission, efficiency) |>
  pivot_wider(names_from = efficiency, values_from = n, values_fill = 0)
```

The whole pipeline is one connected thought, and you can comment out any line to see the intermediate result.

<div class="exercise">
<div class="exercise-title">✏️ Exercise 3.1 — Filter and sort</div>

Using `cars`, find the 5 cars with the worst MPG. Show only the `model`, `mpg`, and `hp` columns.

<details>
<summary>Show solution</summary>

```{r}
cars |>
  arrange(mpg) |>
  select(model, mpg, hp) |>
  head(5)
```

Or with `slice_min`:

```{r}
cars |>
  slice_min(mpg, n = 5) |>
  select(model, mpg, hp)
```

</details>
</div>

<div class="exercise">
<div class="exercise-title">✏️ Exercise 3.2 — Group and summarise</div>

For each unique combination of cylinders (`cyl`) and gears (`gear`), compute:

- the number of cars (`n`)
- the mean MPG
- the mean horsepower

Sort the result with the highest mean MPG first.

<details>
<summary>Show solution</summary>

```{r}
cars |>
  group_by(cyl, gear) |>
  summarise(
    n = n(),
    mean_mpg = mean(mpg),
    mean_hp = mean(hp),
    .groups = "drop"
  ) |>
  arrange(desc(mean_mpg))
```

</details>
</div>

<div class="exercise">
<div class="exercise-title">✏️ Exercise 3.3 — Build a feature</div>

Add two new columns to `cars`:

1. `power_to_weight` — horsepower per 1000 lbs (i.e. `hp / wt`).
2. `category` — `"sporty"` if `power_to_weight > 50`, otherwise `"normal"`.

Then count the cars in each category.

<details>
<summary>Show solution</summary>

```{r}
cars |>
  mutate(
    power_to_weight = hp / wt,
    category = if_else(power_to_weight > 50, "sporty", "normal")
  ) |>
  count(category)
```

</details>
</div>

## What's next

You now have the toolkit to load, clean, and summarise data — easily 60% of any real analysis. Next up: making the results visible. Lesson 4 is about plotting with `ggplot2`, which pairs cleanly with everything you just learned.

<div class="lesson-nav">
  <a class="nav-prev" href="basic_operations_r.html">
    <span class="lesson-nav-label">← Previous</span>
    <span class="lesson-nav-title">Lesson 2: Basic Operations</span>
  </a>
  <a class="nav-next" href="data_visualization_r.html">
    <span class="lesson-nav-label">Next →</span>
    <span class="lesson-nav-title">Lesson 4: Data Visualization</span>
  </a>
</div>
