---
title: "Director's Box Office Trajectory"
editor_options:
  chunk_output_type: console
output:
  html_document:
    includes:
      in_header: header/header.html
      after_body: include_footer.html
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
lang: "en"
---

```{r og-meta, echo=FALSE, results='asis'}
cat('
<meta property="og:title" content="Director Box Office Trajectory">
<meta property="og:description" content="Career box office trajectories for the 10 highest-grossing film directors of all time, in nominal and inflation-adjusted USD.">
<meta property="og:url" content="https://cavandonohoe.github.io/director_box_office_trajectory.html">
<meta property="og:type" content="article">
')
```

```{r pipe, include=FALSE}
`%>%` <- magrittr::`%>%`
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.width = 10, fig.height = 6, dpi = 110
)
```

```{r data}
films_path <- here::here("data", "director_filmographies.csv")
meta_path <- here::here("data", "director_filmographies_meta.csv")

data_available <- file.exists(films_path) && file.exists(meta_path)

if (data_available) {
  films <- readr::read_csv(films_path, show_col_types = FALSE)
  meta <- readr::read_csv(meta_path, show_col_types = FALSE)

  director_levels <- meta %>%
    dplyr::arrange(.data$director_rank) %>%
    dplyr::pull(.data$director_display_name)

  films <- films %>%
    dplyr::mutate(
      director_display_name = factor(
        .data$director_display_name,
        levels = director_levels
      )
    )
} else {
  films <- tibble::tibble()
  meta <- tibble::tibble()
}
```

```{r data-placeholder, eval = !data_available, results = 'asis'}
cat(
  '<div class="alert alert-info" style="padding:1em;border:1px solid #b8daff;',
  'background:#cce5ff;border-radius:4px;margin:1em 0;">',
  '<strong>Data not yet available.</strong> This page is populated by ',
  '<code>scripts/refresh_director_filmographies.R</code>, which runs ',
  'monthly via the <code>update_director_filmographies.yml</code> ',
  'GitHub Action. The action requires a <code>TMDB_API_KEY</code> ',
  'repo secret. Once it has run, this page will fill in.',
  '</div>\n'
)
```

```{r timestamp, eval = data_available, results='asis'}
cat(sprintf(
  '<span class="data-timestamp">Data last updated: %s</span>\n',
  format(as.Date(system(paste("git log -1 --format=%ai --",
    here::here("data", "director_filmographies.csv")), intern = TRUE)), "%B %d, %Y")
))
```

How does a director's box office change across their career? Here are the
career trajectories of the **10 highest-grossing film directors of all time**
(by cumulative worldwide box office, per
[Wikipedia](https://en.wikipedia.org/wiki/List_of_highest-grossing_film_directors)),
sourced from [TMDb](https://www.themoviedb.org/) and refreshed monthly.

All dollar amounts are **worldwide** box office. The "inflation-adjusted"
panels convert nominal USD to 2025-equivalent dollars using BLS CPI-U
All Urban Consumers.

## Leaderboard

```{r leaderboard, eval = data_available}
leaderboard <- meta %>%
  dplyr::transmute(
    Rank = .data$director_rank,
    Director = .data$director_display_name,
    Films = .data$n_films,
    `Active years` = paste0(.data$first_year, "-", .data$last_year),
    `Lifetime gross (nominal)` = scales::dollar(
      .data$lifetime_revenue,
      scale = 1e-9, suffix = "B", accuracy = 0.01
    ),
    `Lifetime gross (2025 USD)` = scales::dollar(
      .data$lifetime_revenue_2025,
      scale = 1e-9, suffix = "B", accuracy = 0.01
    ),
    `Avg per film (2025 USD)` = scales::dollar(
      .data$avg_revenue_per_film_2025,
      scale = 1e-6, suffix = "M", accuracy = 1
    ),
    `Avg TMDb vote` = round(.data$avg_vote, 2)
  )

DT::datatable(
  leaderboard,
  rownames = FALSE,
  options = list(
    pageLength = 10, dom = "tip",
    columnDefs = list(list(className = "dt-right", targets = c(2, 4, 5, 6, 7)))
  )
)
```

## Trajectory by career film number {.tabset .tabset-pills}

x-axis is **career film number** (their first directing credit on TMDb is 1,
their second is 2, etc.). Each panel uses its own y-axis to make the shape
of the trajectory visible regardless of scale.

### Inflation-adjusted gross (2025 USD)

```{r traj-real, eval = data_available}
plot_films <- films %>%
  dplyr::filter(.data$revenue_usable, !is.na(.data$revenue_2025))

ggplot2::ggplot(
  plot_films,
  ggplot2::aes(x = .data$career_film_number, y = .data$revenue_2025 / 1e6)
) +
  ggplot2::geom_line(color = "grey40", alpha = 0.6) +
  ggplot2::geom_point(
    ggplot2::aes(size = .data$vote_average, color = .data$release_year),
    alpha = 0.85
  ) +
  ggplot2::facet_wrap(
    ~ .data$director_display_name,
    scales = "free_y", ncol = 2
  ) +
  ggplot2::scale_color_viridis_c(option = "plasma", end = 0.9) +
  ggplot2::scale_size_continuous(range = c(1.5, 5), limits = c(4, 9)) +
  ggplot2::scale_y_continuous(
    labels = scales::label_dollar(scale = 1, suffix = "M")
  ) +
  ggplot2::labs(
    x = "Career film number",
    y = "Worldwide gross (2025 USD, millions)",
    color = "Release year",
    size = "TMDb vote",
    caption = "Data: TMDb. Inflation adjustment: BLS CPI-U."
  ) +
  ggplot2::theme_minimal(base_size = 11) +
  ggplot2::theme(
    panel.grid.minor = ggplot2::element_blank(),
    legend.position = "bottom"
  )
```

### Nominal gross

```{r traj-nominal, eval = data_available}
ggplot2::ggplot(
  plot_films,
  ggplot2::aes(x = .data$career_film_number, y = .data$revenue / 1e6)
) +
  ggplot2::geom_line(color = "grey40", alpha = 0.6) +
  ggplot2::geom_point(
    ggplot2::aes(size = .data$vote_average, color = .data$release_year),
    alpha = 0.85
  ) +
  ggplot2::facet_wrap(
    ~ .data$director_display_name,
    scales = "free_y", ncol = 2
  ) +
  ggplot2::scale_color_viridis_c(option = "plasma", end = 0.9) +
  ggplot2::scale_size_continuous(range = c(1.5, 5), limits = c(4, 9)) +
  ggplot2::scale_y_continuous(
    labels = scales::label_dollar(scale = 1, suffix = "M")
  ) +
  ggplot2::labs(
    x = "Career film number",
    y = "Worldwide gross (nominal USD, millions)",
    color = "Release year",
    size = "TMDb vote"
  ) +
  ggplot2::theme_minimal(base_size = 11) +
  ggplot2::theme(
    panel.grid.minor = ggplot2::element_blank(),
    legend.position = "bottom"
  )
```

### TMDb vote average

Showing only films with at least 100 votes so we don't trust thin samples
on early-career indie titles.

```{r traj-vote, eval = data_available}
vote_films <- films %>%
  dplyr::filter(!is.na(.data$vote_average), .data$vote_count >= 100)

ggplot2::ggplot(
  vote_films,
  ggplot2::aes(x = .data$career_film_number, y = .data$vote_average)
) +
  ggplot2::geom_line(color = "grey40", alpha = 0.6) +
  ggplot2::geom_point(
    ggplot2::aes(size = .data$vote_count, color = .data$release_year),
    alpha = 0.85
  ) +
  ggplot2::facet_wrap(
    ~ .data$director_display_name,
    scales = "free_y", ncol = 2
  ) +
  ggplot2::scale_color_viridis_c(option = "plasma", end = 0.9) +
  ggplot2::scale_size_continuous(range = c(1.5, 5)) +
  ggplot2::labs(
    x = "Career film number",
    y = "TMDb vote average (1-10)",
    color = "Release year",
    size = "Votes",
    caption = "Films with < 100 TMDb votes excluded."
  ) +
  ggplot2::theme_minimal(base_size = 11) +
  ggplot2::theme(
    panel.grid.minor = ggplot2::element_blank(),
    legend.position = "bottom"
  )
```

## Top 5 by director (2025 USD)

```{r top-films, eval = data_available}
top_films <- films %>%
  dplyr::filter(.data$revenue_usable) %>%
  dplyr::group_by(.data$director_display_name) %>%
  dplyr::slice_max(order_by = .data$revenue_2025, n = 5) %>%
  dplyr::arrange(.data$director_display_name, dplyr::desc(.data$revenue_2025)) %>%
  dplyr::ungroup() %>%
  dplyr::transmute(
    Director = .data$director_display_name,
    Title = .data$title,
    Year = .data$release_year,
    `Gross (nominal)` = scales::dollar(
      .data$revenue, scale = 1e-6, suffix = "M", accuracy = 1
    ),
    `Gross (2025 USD)` = scales::dollar(
      .data$revenue_2025, scale = 1e-6, suffix = "M", accuracy = 1
    ),
    `TMDb vote` = round(.data$vote_average, 1)
  )

DT::datatable(
  top_films,
  rownames = FALSE,
  filter = "top",
  options = list(pageLength = 10, dom = "ftip")
)
```

## All films

```{r all-films, eval = data_available}
all_films_table <- films %>%
  dplyr::transmute(
    Director = .data$director_display_name,
    `#` = .data$career_film_number,
    Title = .data$title,
    Year = .data$release_year,
    `Runtime (min)` = .data$runtime,
    `Budget (nominal)` = ifelse(
      .data$budget_usable,
      scales::dollar(.data$budget, scale = 1e-6, suffix = "M", accuracy = 1),
      "—"
    ),
    `Gross (nominal)` = ifelse(
      .data$revenue_usable,
      scales::dollar(.data$revenue, scale = 1e-6, suffix = "M", accuracy = 1),
      "—"
    ),
    `Gross (2025 USD)` = ifelse(
      .data$revenue_usable,
      scales::dollar(.data$revenue_2025, scale = 1e-6, suffix = "M", accuracy = 1),
      "—"
    ),
    `TMDb vote` = round(.data$vote_average, 1),
    Votes = .data$vote_count
  )

DT::datatable(
  all_films_table,
  rownames = FALSE,
  filter = "top",
  options = list(pageLength = 25, dom = "ftip")
)
```

## Notes

- Filmography data is pulled from
  [TMDb](https://www.themoviedb.org/), filtering each director's movie
  credits to those where `job == "Director"`.
- The Russo brothers (Anthony & Joe) are listed as a single
  "Russo Brothers" entry because they always co-direct. TMDb tracks
  them as separate people, so the filmography is pulled via Anthony
  Russo's credits to avoid double-counting.
- Revenue and budget come from TMDb's user-maintained data, which is
  reliable for major releases but can be missing or zero for older
  films and limited releases. Films without revenue are excluded from
  the gross trajectories but included in the vote-average panel and
  the all-films table.
- Inflation adjustment uses BLS CPI-U All Urban Consumers (annual
  averages) and treats the most recent year in the embedded table as
  the reference. Update once a year alongside the workflow.
