---
title: "IMDb Rating Plot by Season and Episode"
editor_options: 
  chunk_output_type: console
output:
  html_document:
    includes:
      in_header: header/header.html
      after_body: include_footer.html
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
lang: "en"
---

```{r pipe, include=FALSE}
`%>%` <- magrittr::`%>%`
```

I'm kinda obsessed with seeing a whole map of series and where they stride well and when they just absolutely fuck different episodes. Sometimes, critics are just a little harsh. Sometimes it's hard to completely crush an entire series, but we can explore some series that figure it out and quit while they're ahead.

Each show is collapsed by default. Click a title to expand its interactive heatmap, or use the summary table below to jump around.


```{r ci-outdir, include=FALSE}
# ensure output dir exists in clean CI environment
out_dir <- "imdb_rating_plot_files"
if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

# optional: make knitr put figures here too
knitr::opts_chunk$set(fig.path = paste0(out_dir, "/"))
```


```{r setup, include=FALSE}
# Set default chunk output
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  results = "asis"
)
```

```{css heatmap-styles, echo=FALSE}
/* Let a wide heatmap scroll sideways on narrow screens instead of being
   squished into an unreadable smear. .heatmap-inner carries a per-show
   min-width; when the viewport is narrower, .heatmap-scroll scrolls. */
.heatmap-scroll {
  overflow-x: auto;
  -webkit-overflow-scrolling: touch;
}
```


```{r episode_plot}
# Function to create a plot of IMDb ratings by season and episode
create_episode_plot <- function(episode_data, title, export_png = TRUE,
                                 png_width = 8, png_height = 6) {
  # make sure we have imdb_rating + hover text
  episode_data <- episode_data %>%
    dplyr::mutate(
      imdb_rating = rating,
      hover_text = paste0(
        "S", stringr::str_pad(season, 2, pad = "0"),
        "E", stringr::str_pad(episode, 2, pad = "0"),
        " — ", title,
        "\nRating: ", sprintf("%.1f", imdb_rating),
        if (!"votes" %in% names(episode_data)) "" else
          paste0("\nVotes: ", scales::comma(votes))
      )
    )

  file_name <- title %>%
    tolower() %>%
    stringr::str_remove_all(":") %>%
    stringr::str_replace_all(" ", "_") %>%
    paste0("imdb_rating_plot_files/", ., ".png")

  episode_plot <- episode_data %>%
    ggplot2::ggplot(
      ggplot2::aes(
        x = season,
        y = episode,
        fill = imdb_rating,
        text = hover_text
      )
    ) +
    ggplot2::geom_tile(color = "white") +
    ggplot2::geom_text(
      ggplot2::aes(label = round(imdb_rating, 1)),
      color = "black"
    ) +
    ggplot2::scale_fill_gradientn(
      colors = RColorBrewer::brewer.pal(9, "RdYlGn"),
      values = scales::rescale(c(0, 4, 6, 8, 10)),
      limits = c(0, 10),
      name = "IMDb Rating"
    ) +
    ggplot2::scale_x_continuous(
      breaks = seq(min(episode_data$season), max(episode_data$season), 1),
      position = "top"
    ) +
    ggplot2::scale_y_reverse(
      breaks = seq(min(episode_data$episode), max(episode_data$episode), 1)
    ) +
    ggplot2::labs(
      title = "IMDb Ratings by Season and Episode",
      x = "Season",
      y = "Episode"
    ) +
    ggplot2::ggtitle(title) +
    ggplot2::theme_minimal()

  if (export_png) {
    ggplot2::ggsave(
      filename = here::here(dirname(file_name), basename(file_name)),
      plot = episode_plot,
      width = png_width,
      height = png_height,
      dpi = 300
    )
  }

  episode_plot
}

interactive_plotly <- function(gg_object) {
  gg_object %>%
    plotly::ggplotly(tooltip = "text") %>%
    plotly::layout(xaxis = list(side = "top"))
}
```

```{r manifest}
# One row per show. Adding a show is a one-row edit here (no new chunk needed).
#   title      - heading shown on the page (also drives the PNG file name)
#   slug       - data/<slug>_ep_ratings.csv
#   fig_width  - knitr figure width (inches) for the on-page interactive plot
#   fig_height - knitr figure height (inches)
#   png_width  - width (inches) used for the exported PNG (defaults to fig_width)
#   png_height - height (inches) used for the exported PNG (defaults to fig_height)
# Leave dims as NA to use the document defaults.
shows <- tibble::tribble(
  ~title, ~slug, ~fig_width, ~fig_height, ~png_width, ~png_height,
  "Always Sunny in Philadelphia", "always_sunny", NA, NA, NA, NA,
  "Andor", "andor", NA, NA, NA, NA,
  "Attack on Titan", "attack_on_titan", NA, 7, NA, NA,
  "Avatar: The Last Airbender", "avatar", NA, NA, NA, NA,
  "Batman: Caped Crusader", "batman_caped_crusader", NA, NA, NA, NA,
  "Batman: The Animated Series", "batman_animated_series", 7, 14, NA, NA,
  "Better Call Saul", "better_call_saul", NA, NA, NA, NA,
  "Blue Eye Samurai", "blue_eye_samurai", NA, NA, NA, NA,
  "Bluey", "bluey", 4, 9, NA, NA,
  "Bojack Horseman", "bojack", NA, NA, NA, NA,
  "Breaking Bad", "breaking_bad", NA, NA, NA, NA,
  "Brooklyn Nine-Nine", "brooklyn_nine_nine", NA, NA, NA, NA,
  "Chernobyl", "chernobyl", NA, NA, NA, NA,
  "Clone Wars", "clone_wars", NA, NA, NA, NA,
  "Death Note", "death_note", NA, 12, NA, NA,
  "Family Guy", "family_guy", 15, 9, 16, 9,
  "Fleabag", "fleabag", NA, NA, NA, NA,
  "Friends", "friends", 7, 7, NA, NA,
  "Game of Thrones", "game_of_thrones", NA, NA, NA, NA,
  "Gravity Falls", "gravity_falls", NA, 8, NA, NA,
  "Grey's Anatomy", "greys_anatomy", 9, 9, NA, NA,
  "Hannibal", "hannibal", NA, NA, NA, NA,
  "House MD", "house", 7, 7, NA, NA,
  "House of the Dragon", "hotd", NA, NA, NA, NA,
  "How I Met Your Mother", "himym", NA, NA, NA, NA,
  "Invincible", "invincible", NA, NA, NA, NA,
  "Love Island (UK)", "love_island", NA, 11, NA, NA,
  "Mad Men", "mad_men", NA, NA, NA, NA,
  "Modern Love", "modern_love", NA, NA, NA, NA,
  "Mr Robot", "mr_robot", NA, NA, NA, NA,
  "My Hero Academia", "my_hero_academia", NA, NA, NA, NA,
  "New Girl", "new_girl", NA, NA, NA, NA,
  "Paris Hilton's My New BFF", "paris_hilton_bff", NA, NA, NA, NA,
  "Parks and Recreation", "parks_and_rec", NA, NA, NA, NA,
  "Person of Interest", "person_of_interest", NA, NA, NA, NA,
  "Pretty Little Liars", "pretty_little_liars", NA, NA, NA, NA,
  "Rick and Morty", "rick_and_morty", NA, NA, NA, NA,
  "Schitt's Creek", "schitts_creek", NA, NA, NA, NA,
  "Six Feet Under", "six_feet_under", NA, NA, NA, NA,
  "South Park", "south_park", 15, 6, 16, 6,
  "Sparticus", "sparticus", NA, NA, NA, NA,
  "Stranger Things", "stranger_things", NA, NA, NA, NA,
  "Suits", "suits", NA, NA, NA, NA,
  "Ted Lasso", "ted_lasso", NA, NA, NA, NA,
  "The Boys", "the_boys", NA, NA, NA, NA,
  # "The Crown", "crown", NA, NA, NA, NA,
  "The Office (US)", "the_office", NA, NA, NA, NA,
  "The Simpsons", "simpsons", 15, 6, 16, 6,
  "The Wire", "the_wire", NA, NA, NA, NA,
  "Vampire Diaries", "vampire_diaries", NA, NA, NA, NA,
  "Velma", "velma", NA, NA, NA, NA,
  "Westworld", "westworld", NA, NA, NA, NA,
  "X-Men '97", "x_men_97", NA, NA, NA, NA
  # "ZeroZeroZero", "zerozerozero", NA, NA, NA, NA
) %>%
  dplyr::arrange(tolower(title))

# helper: turn a title into the same anchor slug R Markdown auto-generates
# for a "## Title" heading, so the summary table can link to each section.
anchor_id <- function(x) {
  x %>%
    tolower() %>%
    stringr::str_replace_all("[^a-z0-9]+", "-") %>%
    stringr::str_replace_all("^-+|-+$", "")
}

read_show <- function(slug) {
  readr::read_csv(here::here("data", paste0(slug, "_ep_ratings.csv")))
}
```

```{r summary_table}
summary_tbl <- shows %>%
  dplyr::mutate(
    data = purrr::map(slug, read_show),
    Seasons = purrr::map_int(data, ~ dplyr::n_distinct(.x$season)),
    Episodes = purrr::map_int(data, nrow),
    `Avg rating` = purrr::map_dbl(data, ~ round(mean(.x$rating, na.rm = TRUE), 2)),
    `Best ep` = purrr::map_chr(data, function(d) {
      top <- d %>% dplyr::slice_max(rating, n = 1, with_ties = FALSE)
      sprintf("%.1f — %s (S%dE%d)", top$rating, top$title, top$season, top$episode)
    }),
    `Worst ep` = purrr::map_chr(data, function(d) {
      low <- d %>% dplyr::slice_min(rating, n = 1, with_ties = FALSE)
      sprintf("%.1f — %s (S%dE%d)", low$rating, low$title, low$season, low$episode)
    }),
    Show = sprintf("<a href='#%s'>%s</a>", anchor_id(title), title)
  ) %>%
  dplyr::select(Show, Seasons, Episodes, `Avg rating`, `Best ep`, `Worst ep`)

DT::datatable(
  summary_tbl,
  escape = FALSE,
  rownames = FALSE,
  filter = "top",
  options = list(
    pageLength = 15,
    autoWidth = TRUE,
    order = list(list(0, "asc"))
  ),
  caption = "Click a show name to jump to its heatmap. Sort or filter any column."
)
```

```{r plotly_dependency_primer, results='markup'}
# Guarantee the plotly + htmlwidgets JS/CSS dependencies are emitted into the
# page. The show heatmaps below are printed as raw HTML from a results='asis'
# loop, and dependency registration from inside such a loop is dropped by the
# render_site() build path (works with plain render(), fails on the site
# build). Printing one real, hidden plotly widget here routes its dependencies
# through knitr's normal htmlwidget path, so the library loads reliably; every
# raw-HTML widget below then reuses that already-loaded library.
htmltools::tagList(
  htmltools::tags$div(
    style = "display:none;",
    plotly::plot_ly(x = 1, y = 1, type = "scatter", mode = "markers")
  )
)
```

```{r render_shows}
# Emit one collapsible <details> section per show, in the manifest order
# (alphabetical). Each section keeps its own interactive plotly, so hovering
# still reveals the episode title.
render_one_show <- function(row) {
  data <- read_show(row$slug)

  png_w <- if (is.na(row$png_width)) {
    if (is.na(row$fig_width)) 8 else row$fig_width
  } else {
    row$png_width
  }
  png_h <- if (is.na(row$png_height)) {
    if (is.na(row$fig_height)) 6 else row$fig_height
  } else {
    row$png_height
  }

  gg <- create_episode_plot(
    data, row$title,
    png_width = png_w, png_height = png_h
  )
  widget <- interactive_plotly(gg)
  rendered <- htmltools::renderTags(widget)

  # Keep the heatmap legible on narrow screens: give the plot a per-season
  # minimum width and let the container scroll horizontally when the viewport
  # is too small. Wide shows (many seasons) then stay readable via a swipe
  # instead of being squished into a smear; shows with few seasons keep a
  # small min-width and still fit without scrolling.
  n_seasons <- dplyr::n_distinct(data$season)
  min_px <- max(320, n_seasons * 46 + 140)

  cat(sprintf("\n## %s\n\n", row$title))
  cat("<details>\n")
  cat(sprintf("<summary>Show %s heatmap</summary>\n\n", row$title))
  cat("\n\n```{=html}\n")
  cat("<div class=\"heatmap-scroll\">\n")
  cat(sprintf("<div class=\"heatmap-inner\" style=\"min-width:%dpx;\">\n", min_px))
  cat(rendered$html)
  cat("\n</div>\n</div>\n")
  cat("\n```\n\n")
  cat("\n</details>\n\n")
}

for (i in seq_len(nrow(shows))) {
  render_one_show(shows[i, ])
}
```

[Raw pngs found here](https://github.com/cavandonohoe/cavandonohoe.github.io/tree/gh-pages/imdb_rating_plot_files)

```{js resize-plots-on-expand, echo=FALSE}
// Plotly sizes each chart to its container when the page loads. A collapsed
// <details> has zero width at that point, so the heatmap renders blank until
// something forces a redraw. When a <details> is opened, resize any Plotly
// charts inside it so they fill the now-visible container.
(function () {
  function resizePlotsIn(container) {
    if (typeof Plotly === "undefined") return;
    container.querySelectorAll(".plotly.html-widget").forEach(function (el) {
      // only resize once it actually has a rendered plot
      if (el.querySelector(".plot-container")) {
        Plotly.Plots.resize(el);
      }
    });
  }

  document.addEventListener("DOMContentLoaded", function () {
    document.querySelectorAll("details").forEach(function (d) {
      d.addEventListener("toggle", function () {
        if (d.open) {
          // defer so layout has settled before measuring width
          window.requestAnimationFrame(function () {
            resizePlotsIn(d);
          });
        }
      });
    });
  });
})();
```
