---
title: "Which TV Shows Have Exceptional Episodes?"
description: "IMDb episodes rated 9.7 or higher across popular TV series, with vote thresholds and a Top 250 comparison."
lang: "en"
output:
  html_document:
    includes:
      in_header: header/header.html
      after_body: include_footer.html
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: true
      smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(dplyr)
source(here::here("scripts", "imdb_exceptional_metrics.R"))
source(here::here("scripts", "imdb_plotly_layout.R"))
result_dir <- here::here("analysis", "imdb-exceptional-episodes", "results")
series <- readr::read_csv(file.path(result_dir, "eligible_series.csv"), show_col_types = FALSE)
episodes <- readr::read_csv(file.path(result_dir, "exceptional_episodes.csv"), show_col_types = FALSE)
snapshot <- readr::read_csv(file.path(result_dir, "cohort_summary.csv"), show_col_types = FALSE)
validation <- readr::read_csv(file.path(result_dir, "validation_summary.csv"), show_col_types = FALSE)
counts <- lapply(c(1000L, 5000L, 10000L), function(n) imdb_exceptional_counts(series, episodes, n))
names(counts) <- c("1000", "5000", "10000")
primary <- counts[["5000"]]
sensitivity <- dplyr::bind_rows(lapply(names(counts), function(n) {
  d <- counts[[n]]
  tibble::tibble(
    episode_votes = as.integer(n), eligible_shows = nrow(d),
    qualifying_shows = sum(d$exceptional_episodes > 0),
    exceptional_episodes = sum(d$exceptional_episodes),
    percent_shows = 100 * mean(d$exceptional_episodes > 0)
  )
}))
expected <- snapshot |> dplyr::filter(cohort == "broader_eligible")
stopifnot(
  identical(sensitivity$episode_votes, as.integer(expected$episode_vote_threshold)),
  all(sensitivity$eligible_shows == expected$cohort_shows),
  all(sensitivity$qualifying_shows == expected$shows_with_exceptional_episodes),
  all(sensitivity$exceptional_episodes == expected$exceptional_episodes)
)
main <- sensitivity |> dplyr::filter(episode_votes == 5000)
outside <- primary |> dplyr::filter(!saved_cohort, exceptional_episodes > 0)
fmt <- function(x) format(x, big.mark = ",", scientific = FALSE, trim = TRUE)
```

<style>
/* Let a vertical swipe over either chart scroll the page on touch screens. */
.plotly.html-widget, .plotly .svg-container, .plotly .main-svg, .plotly .draglayer {
  touch-action: pan-y !important;
}
.plotly.html-widget { max-width: 100%; }
</style>

Some shows have an episode that everyone seems to agree is extraordinary. How rare is that?
I looked beyond the IMDb Top 250 at **all TV series and miniseries with at least 25,000 series votes**
in IMDb's official downloadable datasets. An exceptional episode earns **at least 9.7/10 with at least
5,000 episode votes**.

**Snapshot: 3 October 2026.** Ratings and vote counts can change.

```{r headline, results='asis'}
cat(sprintf(
  '<div class="row" aria-label="Primary results">
  <div class="col-sm-4"><div class="well"><strong style="font-size:2em">%s</strong><br>eligible shows</div></div>
  <div class="col-sm-4"><div class="well"><strong style="font-size:2em">%s</strong><br>shows with exceptional episodes (%s%%)</div></div>
  <div class="col-sm-4"><div class="well"><strong style="font-size:2em">%s</strong><br>exceptional episodes</div></div></div>',
  fmt(main$eligible_shows), fmt(main$qualifying_shows), sprintf("%.2f", main$percent_shows),
  fmt(main$exceptional_episodes)
))
```

**`r nrow(outside)` qualifying shows are outside the saved Top 250 cohort**, contributing
`r sum(outside$exceptional_episodes)` exceptional episodes. The saved list is a historical comparison:
I haven't verified today's Top 250 membership because IMDb's live chart blocked access.

## Which shows have the most?

This is an episode count, so longer shows have more chances to qualify. It is not a ranking of overall show quality.
The labels show each count and its series-vote total; the full searchable table follows.
On phones, labels sit above thicker bars that use the available chart width.

```{r leaderboard-chart}
leaders <- primary |> dplyr::filter(exceptional_episodes > 0) |> dplyr::slice_head(n = 20)
leaders$series_title <- factor(leaders$series_title, levels = rev(leaders$series_title))
chart <- plotly::plot_ly(
  leaders, height = 650, x = ~exceptional_episodes, y = ~series_title, type = "bar", orientation = "h",
  text = ~paste0(exceptional_episodes, " exceptional episodes<br>",
                 fmt(series_votes), " series votes"),
  hoverinfo = "skip", textposition = "outside", textfont = list(size = 11),
  cliponaxis = FALSE, marker = list(color = "#287d8e")
)
plotly::layout(chart, height = 650, dragmode = FALSE, margin = list(l = 190),
               xaxis = list(title = "Episodes rated ≥9.7 with ≥5,000 votes",
                            range = c(0, max(leaders$exceptional_episodes) * 1.65)),
               yaxis = list(title = "")) |>
  plotly::config(scrollZoom = FALSE, doubleClick = FALSE) |>
  imdb_plotly_layout(
    titles = leaders$series_title,
    labels = paste0(leaders$exceptional_episodes, " exceptional episodes<br>",
                    fmt(leaders$series_votes), " series votes"),
    left_margin = 190, axis_title = "Episodes rated ≥9.7 with ≥5,000 votes",
    compact_axis_title = "Exceptional episodes"
  )
```

## Explore the shows

All `r fmt(nrow(primary))` eligible shows are included, even those with zero exceptional episodes.
Search for a title or genre. On narrow screens, expand a show’s row to see all details and vote cutoffs.
“Saved Top 250” refers to historical membership, not today's chart.

**Exceptional share = 100 × exceptional episodes ÷ total episodes.** The table's percentage
uses the primary cutoff: rating ≥9.7 with ≥5,000 episode votes. “Total episodes” means every
episode linked to the show in IMDb's snapshot, including specials, unrated entries, and any
upcoming placeholders. It is not restricted to episodes that received enough votes.

```{r shows-table}
show_table <- primary |> dplyr::transmute(
  Show = series_title, Year = start_year, Type = title_type,
  Genres = genres, `Series rating` = series_rating, `Series votes` = series_votes,
  `Total episodes` = linked_episodes,
  `Episodes: ≥1,000 votes` = exceptional_1000,
  `Episodes: ≥5,000 votes` = exceptional_5000,
  `Episodes: ≥10,000 votes` = exceptional_10000,
  `Exceptional share (%)` = percent_exceptional,
  `Saved Top 250` = ifelse(saved_cohort, "Yes", "No")
)
DT::datatable(show_table, rownames = FALSE, extensions = c("Buttons", "Responsive"),
              caption = "Exceptional episodes have rating ≥9.7. All eligible shows remain in the denominator.",
              options = list(pageLength = 15, responsive = TRUE, order = list(list(8, "desc")),
                             columnDefs = list(
                               list(responsivePriority = 1, targets = 0),
                               list(responsivePriority = 2, targets = 10),
                               list(responsivePriority = 3, targets = 8),
                               list(responsivePriority = 4, targets = 6)
                             ),
                             dom = "Bfrtip", buttons = c("copy", "csv"))) |>
  DT::formatRound(c("Series votes", "Total episodes", "Episodes: ≥1,000 votes",
                    "Episodes: ≥5,000 votes", "Episodes: ≥10,000 votes"), digits = 0) |>
  DT::formatRound("Exceptional share (%)", digits = 2)
```

## Which shows have the highest exceptional share?

This view divides the primary exceptional-episode count by the show's total episode count.
Short shows can rank highly with only a few standout episodes. The labels show the numerator, denominator, and percentage.
Shows without an exceptional episode are omitted from this chart and remain in the full table above.

```{r percentage-chart}
share_leaders <- primary |>
  dplyr::filter(exceptional_episodes > 0, linked_episodes > 0) |>
  dplyr::arrange(dplyr::desc(percent_exceptional), dplyr::desc(exceptional_episodes), series_id) |>
  dplyr::slice_head(n = 20)
share_leaders$series_title <- factor(share_leaders$series_title, levels = rev(share_leaders$series_title))
plotly::plot_ly(
  share_leaders, height = 650, x = ~percent_exceptional, y = ~series_title, type = "bar", orientation = "h",
  text = ~paste0(exceptional_episodes, " / ", linked_episodes, " episodes<br>",
                 sprintf("%.2f", percent_exceptional), "%"),
  hoverinfo = "skip", textposition = "outside", textfont = list(size = 11),
  cliponaxis = FALSE, marker = list(color = "#6b5b95")
) |>
  plotly::layout(height = 650, dragmode = FALSE, margin = list(l = 210),
                 xaxis = list(title = "Share of all episodes rated ≥9.7 with ≥5,000 votes (%)",
                              range = c(0, max(share_leaders$percent_exceptional) * 1.55),
                              ticksuffix = "%"),
                 yaxis = list(title = "")) |>
  plotly::config(scrollZoom = FALSE, doubleClick = FALSE) |>
  imdb_plotly_layout(
    titles = share_leaders$series_title,
    labels = paste0(share_leaders$exceptional_episodes, " / ", share_leaders$linked_episodes,
                    " episodes<br>", sprintf("%.2f", share_leaders$percent_exceptional), "%"),
    left_margin = 210, axis_title = "Share of all episodes rated ≥9.7 with ≥5,000 votes (%)",
    compact_axis_title = "Exceptional share (%)", compact_row_height = 76
  )
```

## How sensitive is the result to episode votes?

The rating cutoff stays at 9.7 and the series-vote cutoff stays at 25,000. Only the minimum
number of episode votes changes.

```{r sensitivity-table}
knitr::kable(sensitivity |> dplyr::transmute(
  `Minimum episode votes` = fmt(episode_votes), `Shows with exceptional episodes` = qualifying_shows,
  `Share of eligible shows` = sprintf("%.2f%%", percent_shows),
  `Exceptional episodes` = exceptional_episodes
), align = c("r", "r", "r", "r"))
```

## Explore the episodes {.tabset}

Each tab applies its own minimum episode-vote cutoff. Every episode is from an eligible show and
has rating ≥9.7. Search for a show or episode title; use the IMDb link to open its episode page.

```{r episode-widget-function, include=FALSE}
episode_widget <- function(threshold) {
  selected <- episodes |>
    dplyr::filter(series_id %in% primary$series_id, rating >= 9.7, votes >= threshold) |>
    dplyr::arrange(dplyr::desc(votes)) |>
    dplyr::transmute(
      Show = series_title, Season = season, Episode = episode, Title = episode_title,
      Rating = rating, Votes = votes,
      IMDb = paste0('<a href="https://www.imdb.com/title/', episode_id, '/">Episode page</a>')
    )
  DT::datatable(selected, rownames = FALSE, escape = c(1:6),
                extensions = c("Buttons", "Responsive"),
                options = list(pageLength = 15, responsive = TRUE, order = list(list(5, "desc")),
                               columnDefs = list(
                                 list(responsivePriority = 1, targets = 0),
                                 list(responsivePriority = 2, targets = 4),
                                 list(responsivePriority = 3, targets = 3)
                               ),
                               dom = "Bfrtip", buttons = c("copy", "csv"))) |>
    DT::formatRound("Votes", digits = 0)
}
```

### 5,000 votes

```{r episodes-primary}
episode_widget(5000L)
```

### 1,000 votes

```{r episodes-1000}
episode_widget(1000L)
```

### 10,000 votes

```{r episodes-10000}
episode_widget(10000L)
```

## Comparison with the saved Top 250

The existing CSV contains 250 distinct show IDs. Of those, **218 meet the 25,000-series-vote cutoff**
in this snapshot. This comparison uses official episode ratings for those historical IDs.
I haven't rebuilt the chart by sorting bulk ratings: IMDb's Top 250 uses its own ranking rules.

```{r cohort-table}
cohort_table <- snapshot |>
  dplyr::filter(episode_vote_threshold == 5000) |>
  dplyr::mutate(Cohort = dplyr::recode(cohort,
    broader_eligible = "All eligible shows", saved_cohort_all = "Saved Top 250: all members",
    saved_cohort_eligible = "Saved Top 250: eligible members",
    outside_saved_cohort = "Eligible shows outside saved Top 250")) |>
  dplyr::transmute(Cohort, Shows = cohort_shows,
                   `Shows with exceptional episodes` = shows_with_exceptional_episodes,
                   Share = sprintf("%.2f%%", percent_shows), `Exceptional episodes` = exceptional_episodes)
knitr::kable(cohort_table)
```

The cohort rows overlap: eligible saved members and eligible shows outside the saved list together
make up the broader universe. The all-member saved cohort also includes 32 shows below the series-vote cutoff.

## Sources, validation, and limitations

The universe comes from [IMDb's official non-commercial datasets](https://data.imdb.com/non-commercial-datasets/):
`title.basics` identifies series and miniseries, `title.episode` links episodes to their parent shows,
and `title.ratings` provides weighted ratings and vote counts. Series votes are for the show itself;
they are not a sum of episode votes. There are no language, country, animation, or documentary exclusions.

Existing CSVs serve as validation checks. The stored Top 250 CSV has 15,519 rows; 15,405 match a unique
official series/season/episode key. Twenty matched rows change their primary exceptional classification
between snapshots. Because the stored CSV lacks episode IDs, matching is provisional and can be affected
by renumbering as well as changes in ratings and votes.

Specials and season-zero entries are included. The downloads lack exact episode air dates, so this analysis
does not apply an exact release-date filter. A rating record is required. Vote cutoffs reduce the influence
of very small samples but do not eliminate fandom effects or organized voting.

[Download all eligible shows](analysis/imdb-exceptional-episodes/results/eligible_series.csv) ·
[Download primary exceptional episodes](analysis/imdb-exceptional-episodes/results/primary_exceptional_episodes.csv) ·
[Download cohort sensitivity results](analysis/imdb-exceptional-episodes/results/cohort_summary.csv) ·
[Download CSV validation summary](analysis/imdb-exceptional-episodes/results/validation_summary.csv)

The [project source and refresh instructions](https://github.com/cavandonohoe/cavandonohoe.github.io/tree/75d095f24e258d90ef3f0d56b2e714df9bf0f6e5/analysis/imdb-exceptional-episodes)
include dataset hashes and source HTTP metadata for this dated snapshot. The page renders from checked-in
data; building the website does not download IMDb's full datasets.
