---
title: "Which TV Shows Have Exceptional Episodes?"
description: "IMDb episodes rated 9.7 or higher across popular TV series, with vote thresholds and a Top 250 comparison."
lang: "en"
output:
  html_document:
    includes:
      in_header: header/header.html
      after_body: include_footer.html
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: true
      smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(dplyr)
source(here::here("scripts", "imdb_exceptional_metrics.R"))
result_dir <- here::here("analysis", "imdb-exceptional-episodes", "results")
series <- readr::read_csv(file.path(result_dir, "eligible_series.csv"), show_col_types = FALSE)
episodes <- readr::read_csv(file.path(result_dir, "exceptional_episodes.csv"), show_col_types = FALSE)
snapshot <- readr::read_csv(file.path(result_dir, "cohort_summary.csv"), show_col_types = FALSE)
validation <- readr::read_csv(file.path(result_dir, "validation_summary.csv"), show_col_types = FALSE)
counts <- lapply(c(1000L, 5000L, 10000L), function(n) imdb_exceptional_counts(series, episodes, n))
names(counts) <- c("1000", "5000", "10000")
primary <- counts[["5000"]]
sensitivity <- dplyr::bind_rows(lapply(names(counts), function(n) {
  d <- counts[[n]]
  tibble::tibble(
    episode_votes = as.integer(n), eligible_shows = nrow(d),
    qualifying_shows = sum(d$exceptional_episodes > 0),
    exceptional_episodes = sum(d$exceptional_episodes),
    percent_shows = 100 * mean(d$exceptional_episodes > 0)
  )
}))
expected <- snapshot |> dplyr::filter(cohort == "broader_eligible")
stopifnot(
  identical(sensitivity$episode_votes, as.integer(expected$episode_vote_threshold)),
  all(sensitivity$eligible_shows == expected$cohort_shows),
  all(sensitivity$qualifying_shows == expected$shows_with_exceptional_episodes),
  all(sensitivity$exceptional_episodes == expected$exceptional_episodes)
)
main <- sensitivity |> dplyr::filter(episode_votes == 5000)
outside <- primary |> dplyr::filter(!saved_cohort, exceptional_episodes > 0)
fmt <- function(x) format(x, big.mark = ",", scientific = FALSE, trim = TRUE)
```

Some shows have an episode that everyone seems to agree is extraordinary. How rare is that?
I looked beyond the IMDb Top 250 at **all TV series and miniseries with at least 25,000 series votes**
in IMDb's official downloadable datasets. An exceptional episode earns **at least 9.7/10 with at least
5,000 episode votes**.

**Snapshot: 3 October 2026.** Ratings and vote counts can change.

```{r headline, results='asis'}
cat(sprintf(
  '<div class="row" aria-label="Primary results">
  <div class="col-sm-4"><div class="well"><strong style="font-size:2em">%s</strong><br>eligible shows</div></div>
  <div class="col-sm-4"><div class="well"><strong style="font-size:2em">%s</strong><br>shows with exceptional episodes (%s%%)</div></div>
  <div class="col-sm-4"><div class="well"><strong style="font-size:2em">%s</strong><br>exceptional episodes</div></div></div>',
  fmt(main$eligible_shows), fmt(main$qualifying_shows), sprintf("%.2f", main$percent_shows),
  fmt(main$exceptional_episodes)
))
```

**`r nrow(outside)` qualifying shows are outside the saved Top 250 cohort**, contributing
`r sum(outside$exceptional_episodes)` exceptional episodes. The saved list is a historical comparison:
I haven't verified today's Top 250 membership because IMDb's live chart blocked access.

## Which shows have the most?

This is an episode count, so longer shows have more chances to qualify. It is not a ranking of overall show quality.
Hover for details; the full searchable table follows.

```{r leaderboard-chart}
leaders <- primary |> dplyr::filter(exceptional_episodes > 0) |> dplyr::slice_head(n = 20)
leaders$series_title <- factor(leaders$series_title, levels = rev(leaders$series_title))
chart <- plotly::plot_ly(
  leaders, x = ~exceptional_episodes, y = ~series_title, type = "bar", orientation = "h",
  text = ~paste0(series_title, "<br>", exceptional_episodes, " exceptional episodes",
                 "<br>Series votes: ", fmt(series_votes)),
  hoverinfo = "text", textposition = "none", marker = list(color = "#287d8e")
)
plotly::layout(chart, height = 650, margin = list(l = 190),
               xaxis = list(title = "Episodes rated ≥9.7 with ≥5,000 votes", rangemode = "tozero"),
               yaxis = list(title = ""))
```

## Explore the shows

All `r fmt(nrow(primary))` eligible shows are included, even those with zero exceptional episodes.
Search for a title or genre. The three count columns let you compare episode-vote cutoffs.
“Saved Top 250” refers to historical membership, not today's chart.

```{r shows-table}
show_table <- primary |> dplyr::transmute(
  Show = series_title, Year = start_year, Type = title_type,
  Genres = genres, `Series rating` = series_rating, `Series votes` = series_votes,
  `Episodes: ≥1,000 votes` = exceptional_1000,
  `Episodes: ≥5,000 votes` = exceptional_5000,
  `Episodes: ≥10,000 votes` = exceptional_10000,
  `Saved Top 250` = ifelse(saved_cohort, "Yes", "No")
)
DT::datatable(show_table, rownames = FALSE, filter = "top", extensions = "Buttons",
              caption = "Exceptional episodes have rating ≥9.7. All eligible shows remain in the denominator.",
              options = list(pageLength = 15, scrollX = TRUE, order = list(list(7, "desc")),
                             dom = "Bfrtip", buttons = c("copy", "csv"))) |>
  DT::formatRound(c("Series votes", "Episodes: ≥1,000 votes", "Episodes: ≥5,000 votes",
                    "Episodes: ≥10,000 votes"), digits = 0)
```

## How sensitive is the result to episode votes?

The rating cutoff stays at 9.7 and the series-vote cutoff stays at 25,000. Only the minimum
number of episode votes changes.

```{r sensitivity-table}
knitr::kable(sensitivity |> dplyr::transmute(
  `Minimum episode votes` = fmt(episode_votes), `Shows with exceptional episodes` = qualifying_shows,
  `Share of eligible shows` = sprintf("%.2f%%", percent_shows),
  `Exceptional episodes` = exceptional_episodes
), align = c("r", "r", "r", "r"))
```

## Explore the episodes {.tabset}

Each tab applies its own minimum episode-vote cutoff. Every episode is from an eligible show and
has rating ≥9.7. Search for a show or episode title; use the IMDb link to open its episode page.

```{r episode-widget-function, include=FALSE}
episode_widget <- function(threshold) {
  selected <- episodes |>
    dplyr::filter(series_id %in% primary$series_id, rating >= 9.7, votes >= threshold) |>
    dplyr::arrange(dplyr::desc(votes)) |>
    dplyr::transmute(
      Show = series_title, Season = season, Episode = episode, Title = episode_title,
      Rating = rating, Votes = votes,
      IMDb = paste0('<a href="https://www.imdb.com/title/', episode_id, '/">Episode page</a>')
    )
  DT::datatable(selected, rownames = FALSE, escape = c(1:6), extensions = "Buttons",
                options = list(pageLength = 15, scrollX = TRUE, order = list(list(5, "desc")),
                               dom = "Bfrtip", buttons = c("copy", "csv"))) |>
    DT::formatRound("Votes", digits = 0)
}
```

### 5,000 votes

```{r episodes-primary}
episode_widget(5000L)
```

### 1,000 votes

```{r episodes-1000}
episode_widget(1000L)
```

### 10,000 votes

```{r episodes-10000}
episode_widget(10000L)
```

## Comparison with the saved Top 250

The existing CSV contains 250 distinct show IDs. Of those, **218 meet the 25,000-series-vote cutoff**
in this snapshot. This comparison uses official episode ratings for those historical IDs.
I haven't rebuilt the chart by sorting bulk ratings: IMDb's Top 250 uses its own ranking rules.

```{r cohort-table}
cohort_table <- snapshot |>
  dplyr::filter(episode_vote_threshold == 5000) |>
  dplyr::mutate(Cohort = dplyr::recode(cohort,
    broader_eligible = "All eligible shows", saved_cohort_all = "Saved Top 250: all members",
    saved_cohort_eligible = "Saved Top 250: eligible members",
    outside_saved_cohort = "Eligible shows outside saved Top 250")) |>
  dplyr::transmute(Cohort, Shows = cohort_shows,
                   `Shows with exceptional episodes` = shows_with_exceptional_episodes,
                   Share = sprintf("%.2f%%", percent_shows), `Exceptional episodes` = exceptional_episodes)
knitr::kable(cohort_table)
```

The cohort rows overlap: eligible saved members and eligible shows outside the saved list together
make up the broader universe. The all-member saved cohort also includes 32 shows below the series-vote cutoff.

## Sources, validation, and limitations

The universe comes from [IMDb's official non-commercial datasets](https://data.imdb.com/non-commercial-datasets/):
`title.basics` identifies series and miniseries, `title.episode` links episodes to their parent shows,
and `title.ratings` provides weighted ratings and vote counts. Series votes are for the show itself;
they are not a sum of episode votes. There are no language, country, animation, or documentary exclusions.

Existing CSVs serve as validation checks. The stored Top 250 CSV has 15,519 rows; 15,405 match a unique
official series/season/episode key. Twenty matched rows change their primary exceptional classification
between snapshots. Because the stored CSV lacks episode IDs, matching is provisional and can be affected
by renumbering as well as changes in ratings and votes.

Specials and season-zero entries are included. The downloads lack exact episode air dates, so this analysis
does not apply an exact release-date filter. A rating record is required. Vote cutoffs reduce the influence
of very small samples but do not eliminate fandom effects or organized voting.

[Download all eligible shows](analysis/imdb-exceptional-episodes/results/eligible_series.csv) ·
[Download primary exceptional episodes](analysis/imdb-exceptional-episodes/results/primary_exceptional_episodes.csv) ·
[Download cohort sensitivity results](analysis/imdb-exceptional-episodes/results/cohort_summary.csv) ·
[Download CSV validation summary](analysis/imdb-exceptional-episodes/results/validation_summary.csv)

The [project source and refresh instructions](https://github.com/cavandonohoe/cavandonohoe.github.io/tree/75d095f24e258d90ef3f0d56b2e714df9bf0f6e5/analysis/imdb-exceptional-episodes)
include dataset hashes and source HTTP metadata for this dated snapshot. The page renders from checked-in
data; building the website does not download IMDb's full datasets.
