---
title: "LA Housing Prices Over Time: Animated Heat Map"
editor_options:
  chunk_output_type: console
output:
  html_document:
    includes:
      in_header: header/header.html
      after_body: include_footer.html
lang: "en"
---

```{r og-meta, echo=FALSE, results='asis'}
cat('
<meta property="og:title" content="LA Housing Prices Over Time: Animated Heat Map">
<meta property="og:description" content="Animated choropleth of typical home values across Los Angeles County ZIP codes from 2000 to today (Zillow ZHVI), plus 130+ years of US home prices from Shiller and Case-Shiller LA Metro from FRED.">
<meta property="og:url" content="https://cavandonohoe.github.io/la_housing_heatmap.html">
<meta property="og:type" content="article">
')
```

```{r pipe, include=FALSE}
`%>%` <- magrittr::`%>%`
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE
)
```

```{r data}
zhvi_history <- readr::read_csv(
  here::here("data", "la_zhvi_zip_history.csv"),
  show_col_types = FALSE,
  col_types = readr::cols(zip = readr::col_character())
)

geojson <- jsonlite::read_json(
  here::here("data", "la_county_zips.geojson"),
  simplifyVector = FALSE
)
```

```{r prep}
year_min <- min(zhvi_history$year)
year_max <- max(zhvi_history$year)
n_zips <- dplyr::n_distinct(zhvi_history$zip)

latest_snapshot <- zhvi_history %>%
  dplyr::filter(year == year_max) %>%
  dplyr::pull(date) %>%
  unique()

zhvi_history <- zhvi_history %>%
  dplyr::mutate(
    year_label = ifelse(
      year == year_max,
      paste0(year, " (", format(latest_snapshot, "%b"), ")"),
      paste0(year, " (Dec)")
    ),
    zhvi_fmt = scales::dollar(zhvi, accuracy = 1e3),
    log_zhvi = log10(zhvi)
  ) %>%
  dplyr::arrange(year, zip)
```

How have home values rippled across Los Angeles since the dot-com era? This map animates **`r scales::comma(n_zips)` LA County ZIP codes** from `r year_min` through `r year_max` using Zillow's typical home value (ZHVI). Press **Play**, drag the slider, or hover any ZIP to see its year-by-year arc and the neighborhood it covers.

For context, this page also includes two **long-run companion charts** that pre-date Zillow's ZIP-level coverage: Robert Shiller's US home-price series back to 1890, and the Case-Shiller LA Metro index back to 1987.

<span class="data-timestamp">Zillow ZHVI snapshot: `r format(latest_snapshot, "%B %Y")`. Years before `r year_max` are December values; `r year_max` is the most recent month available.</span>

### Data Sources

- [Zillow ZHVI (ZIP-level, all homes, smoothed)](https://www.zillow.com/research/data/) --- the typical home value for each ZIP, defined as the smoothed seasonally adjusted mean of the middle third of the market. ZIP-level coverage starts in 2000.
- [OpenDataDE State ZIP Code GeoJSON](https://github.com/OpenDataDE/State-zip-code-GeoJSON) --- ZIP code polygons converted from the U.S. Census Bureau's TIGER/Line ZCTA shapefiles.
- [LA Times Mapping L.A. neighborhood boundaries](https://maps.latimes.com/about/) --- 272 named LA County neighborhoods (Hollywood, Venice, Sherman Oaks, etc.). Spatial-joined to ZIP centroids to give every ZIP a recognizable name. (CC-BY-SA, mirrored via [`blackmad/neighborhoods`](https://github.com/blackmad/neighborhoods).)
- [Robert Shiller's "Online Data" (Fig 3-1)](http://www.econ.yale.edu/~shiller/data.htm) --- US home prices since **1890**, real and nominal, from the dataset behind *Irrational Exuberance*.
- [S&P Cotality Case-Shiller CA-Los Angeles Home Price Index](https://fred.stlouisfed.org/series/LXXRSA) (`LXXRSA`) via FRED --- monthly LA-metro repeat-sales index since **January 1987**.

To refresh the data, run `Rscript scripts/la_heatmap_data.R`.

## The 130-year view

Before zooming into the map: how does the period we're animating fit into the long arc of US housing? There is **no LA-neighborhood-level data going back a century** --- that resolution only exists from the late 1990s on. But we can stitch in two long-run series that reach much further back.

```{r long-run-data}
shiller <- readr::read_csv(
  here::here("data", "shiller_us_home_prices.csv"),
  show_col_types = FALSE
)

cs_la <- readr::read_csv(
  here::here("data", "case_shiller_la.csv"),
  show_col_types = FALSE
)

shiller_year_min <- min(shiller$year)
shiller_year_max <- max(shiller$year)
cs_la_year_min <- as.integer(format(min(cs_la$date), "%Y"))
cs_la_year_max <- as.integer(format(max(cs_la$date), "%Y"))
```

### US real home prices since `r shiller_year_min` (Shiller)

This is the series Robert Shiller assembled for *Irrational Exuberance* --- a US-wide home-price index in real (inflation-adjusted) terms, normalized so that **1890 = 100**. The flat century from 1890 to the late 1990s is famous: until the dot-com era, US homes did not get more expensive in real terms, on average. Then the 2000s and 2020s happened.

```{r shiller-chart, fig.height=5}
shiller_plot_df <- shiller %>%
  dplyr::mutate(
    real_indexed = real_hpi / real_hpi[year == shiller_year_min] * 100
  )

p_shiller <- shiller_plot_df %>%
  ggplot2::ggplot(ggplot2::aes(
    x = year, y = real_indexed,
    text = paste0(year, "<br>Real HPI (1890=100): ",
                  format(round(real_indexed, 1), nsmall = 1))
  )) +
  ggplot2::geom_hline(yintercept = 100, linetype = "dashed",
                      color = "#999999") +
  ggplot2::geom_line(ggplot2::aes(group = 1), color = "#C0392B",
                     size = 0.9) +
  ggplot2::scale_x_continuous(breaks = seq(1890, 2030, 20)) +
  ggplot2::scale_y_continuous(
    name = "Real Home Price Index (1890 = 100)"
  ) +
  ggplot2::labs(x = NULL) +
  ggplot2::theme_minimal(base_size = 12)

plotly::ggplotly(p_shiller, tooltip = "text") %>%
  plotly::config(displaylogo = FALSE)
```

For the entire 20th century, US home prices *barely budged* in inflation-adjusted terms. Two structural breaks dominate everything else: **the 2000s housing bubble** (peak in 2006, ~85% above the long-run baseline) and **the post-COVID surge** (an even higher peak around 2022). The dashed horizontal line marks the 1890 level --- everything above it is "more expensive than 1890" in real terms.

### Case-Shiller LA Metro since `r cs_la_year_min`

For something LA-specific that pre-dates Zillow's 2000 start, the canonical series is the **Case-Shiller LA Metro** repeat-sales index. It begins in January 1987 (the index is normalized to January 2000 = 100).

```{r cs-la-chart, fig.height=5}
p_cs <- cs_la %>%
  ggplot2::ggplot(ggplot2::aes(
    x = date, y = index,
    text = paste0(format(date, "%b %Y"),
                  "<br>Index (Jan 2000 = 100): ",
                  format(round(index, 1), nsmall = 1))
  )) +
  ggplot2::geom_line(ggplot2::aes(group = 1), color = "#C0392B",
                     size = 0.9) +
  ggplot2::scale_y_continuous(
    name = "Case-Shiller LA Index (Jan 2000 = 100)"
  ) +
  ggplot2::labs(x = NULL) +
  ggplot2::theme_minimal(base_size = 12)

plotly::ggplotly(p_cs, tooltip = "text") %>%
  plotly::config(displaylogo = FALSE)
```

LA tells the same story as the nation, only louder: a late-80s peak (around 1990), a long slide through the early 90s, the 2000s bubble (peak 2006, sharp 2008 crash), and an aggressive post-COVID climb that has just started to retrace.

### Why no neighborhood-level history?

ZIP-, tract-, and neighborhood-level price indexes require a large enough volume of repeat home sales (or appraisal-model inputs) to be statistically meaningful in each tiny geography. That data really only exists from the **late 1990s onward**, when MLS records and computational hedonic/repeat-sales models like Zillow's ZHVI made it feasible to estimate a typical home value for every ZIP, every month. Pre-1996, the best you can do for LA is a single metro-wide line. Pre-1987, only the national series exists.

The animated map below, then, is the *finest* spatial resolution for which a long time series is possible.

## Heat map by year

```{r heatmap, results='asis'}
zhvi_for_js <- zhvi_history %>%
  dplyr::transmute(
    zip,
    year,
    zhvi = round(zhvi)
  )

# Build {zip: [v_year0, v_year1, ...]} where year index aligns with YEARS.
years_vec <- sort(unique(zhvi_history$year))
values_by_zip <- zhvi_for_js %>%
  dplyr::group_by(zip) %>%
  dplyr::summarise(
    by_year = list({
      v <- rep(NA_real_, length(years_vec))
      idx <- match(year, years_vec)
      v[idx] <- zhvi
      v
    }),
    .groups = "drop"
  ) %>%
  tibble::deframe()

year_labels <- vapply(years_vec, function(y) {
  if (y == year_max) {
    paste0(y, " (", format(latest_snapshot, "%b"), ")")
  } else {
    paste0(y, " (Dec)")
  }
}, character(1))

zlim <- range(zhvi_history$zhvi, na.rm = TRUE)
log_min <- log10(zlim[1])
log_max <- log10(zlim[2])

color_breaks <- c(2e5, 5e5, 1e6, 2e6, 5e6, 1e7)
color_breaks <- color_breaks[color_breaks >= zlim[1] & color_breaks <= zlim[2]]

geojson_str <- jsonlite::toJSON(geojson, auto_unbox = TRUE, digits = NA)
values_str <- jsonlite::toJSON(values_by_zip, na = "null")
years_str <- jsonlite::toJSON(years_vec)
year_labels_str <- jsonlite::toJSON(year_labels)
breaks_str <- jsonlite::toJSON(color_breaks)
break_labels_str <- jsonlite::toJSON(
  scales::dollar(color_breaks, scale_cut = scales::cut_short_scale(),
                 accuracy = 1)
)

widget_id <- "la-heatmap"

knitr::asis_output(sprintf('
<link rel="stylesheet"
href="https://unpkg.com/leaflet@1.9.4/dist/leaflet.css"
integrity="sha256-p4NxAoJBhIIN+hmNHrzRCf9tD/miZyoHS5obTRR9BMY="
crossorigin="" />
<script
src="https://unpkg.com/leaflet@1.9.4/dist/leaflet.js"
integrity="sha256-20nQCchB9co0qIjJZRGuk2/Z9VM+kNiyxNV1lvTlZBo="
crossorigin=""></script>
<div class="la-heatmap-wrap">
<div id="%s" class="la-heatmap-map"></div>
<div class="la-heatmap-controls">
<button type="button" id="%s-play" class="la-heatmap-play"
aria-label="Play animation">&#9654; Play</button>
<span id="%s-year-label" class="la-heatmap-year-label"></span>
<input type="range" id="%s-slider" class="la-heatmap-slider"
min="0" max="%d" value="%d" step="1"
aria-label="Year selector" />
<div class="la-heatmap-legend" id="%s-legend"></div>
</div>
</div>
<style>
.la-heatmap-wrap {
  position: relative;
  margin: 1rem 0 2rem 0;
}
.la-heatmap-map {
  height: 600px;
  border-radius: 6px;
  box-shadow: 0 1px 4px var(--card-shadow, rgba(0,0,0,0.1));
}
.la-heatmap-controls {
  display: flex;
  align-items: center;
  gap: 12px;
  padding: 10px 14px;
  margin-top: 8px;
  background: var(--card-bg, #f8f9fa);
  border: 1px solid var(--card-border, #dee2e6);
  border-radius: 6px;
  flex-wrap: wrap;
}
.la-heatmap-play {
  background: var(--accordion-bg, #337ab7);
  color: white;
  border: none;
  border-radius: 4px;
  padding: 6px 14px;
  font-size: 14px;
  cursor: pointer;
  font-weight: 500;
}
.la-heatmap-play:hover { background: var(--accordion-hover, #286090); }
.la-heatmap-year-label {
  font-weight: 600;
  font-size: 16px;
  min-width: 110px;
  color: var(--text-color, #333);
}
.la-heatmap-slider { flex: 1; min-width: 200px; }
.la-heatmap-legend {
  display: flex;
  align-items: center;
  gap: 6px;
  font-size: 12px;
  color: var(--muted-text, #777);
  flex-wrap: wrap;
}
.la-heatmap-legend .swatch {
  display: inline-block;
  width: 14px;
  height: 14px;
  border-radius: 2px;
  margin-right: 2px;
  vertical-align: middle;
  border: 1px solid rgba(0,0,0,0.1);
}
.la-heatmap-tooltip {
  font-family: -apple-system, BlinkMacSystemFont, "Segoe UI", sans-serif;
  font-size: 13px;
  line-height: 1.4;
}
</style>

<script>
(function() {
  var ID = "%s";
  var GEOJSON = %s;
  var VALUES = %s;
  var YEARS = %s;
  var YEAR_LABELS = %s;
  var COLOR_BREAKS = %s;
  var BREAK_LABELS = %s;
  var LOG_MIN = %f;
  var LOG_MAX = %f;

  // Sequential heat scale: pale yellow -> orange -> red -> deep red.
  // Designed so the most expensive ZIPs read as "hot" red.
  var HEAT_SCALE = [
    [0.00, [255, 255, 204]],
    [0.20, [255, 237, 160]],
    [0.40, [254, 178,  76]],
    [0.60, [253, 141,  60]],
    [0.75, [252,  78,  42]],
    [0.88, [227,  26,  28]],
    [1.00, [128,   0,  38]]
  ];

  function lerp(a, b, t) { return a + (b - a) * t; }

  function heatColor(t) {
    if (t <= 0) t = 0; if (t >= 1) t = 1;
    for (var i = 1; i < HEAT_SCALE.length; i++) {
      if (t <= HEAT_SCALE[i][0]) {
        var t0 = HEAT_SCALE[i-1][0], t1 = HEAT_SCALE[i][0];
        var c0 = HEAT_SCALE[i-1][1], c1 = HEAT_SCALE[i][1];
        var u = (t - t0) / (t1 - t0);
        return "rgb(" +
          Math.round(lerp(c0[0], c1[0], u)) + "," +
          Math.round(lerp(c0[1], c1[1], u)) + "," +
          Math.round(lerp(c0[2], c1[2], u)) + ")";
      }
    }
    return "rgb(128,0,38)";
  }

  function colorFor(zhvi) {
    if (zhvi == null || isNaN(zhvi)) return "#cccccc";
    var t = (Math.log10(zhvi) - LOG_MIN) / (LOG_MAX - LOG_MIN);
    return heatColor(t);
  }

  function fmtUSD(v) {
    if (v == null || isNaN(v)) return "—";
    if (v >= 1e6) return "$" + (v / 1e6).toFixed(2) + "M";
    if (v >= 1e3) return "$" + Math.round(v / 1e3) + "K";
    return "$" + Math.round(v);
  }

  var map = L.map(ID, {
    scrollWheelZoom: false,
    zoomControl: true
  }).setView([34.05, -118.30], 9);

  L.tileLayer(
    "https://cartodb-basemaps-{s}.global.ssl.fastly.net/light_all/{z}/{x}/{y}.png",
    {
      attribution:
        "Map &copy; <a href=\\"https://carto.com/\\">CARTO</a> | " +
        "&copy; <a href=\\"https://www.openstreetmap.org/copyright\\">OSM</a> | " +
        "Data &copy; <a href=\\"https://www.zillow.com/research/data/\\">Zillow ZHVI</a>",
      subdomains: "abcd",
      maxZoom: 18
    }
  ).addTo(map);

  var currentYearIdx = YEARS.length - 1;

  function styleFor(feature) {
    var zip = feature.properties.zip;
    var arr = VALUES[zip];
    var v = arr ? arr[currentYearIdx] : null;
    return {
      fill: true,
      fillColor: colorFor(v),
      weight: 0.4,
      color: "#ffffff",
      fillOpacity: 0.85
    };
  }

  function buildTooltip(feature) {
    var zip = feature.properties.zip;
    var name = feature.properties.name || "";
    var arr = VALUES[zip];
    var year = YEARS[currentYearIdx];
    var v = arr ? arr[currentYearIdx] : null;
    var firstYear = null, firstVal = null;
    if (arr) {
      for (var i = 0; i < arr.length; i++) {
        if (arr[i] != null) {
          firstYear = YEARS[i];
          firstVal = arr[i];
          break;
        }
      }
    }
    var growth = "";
    if (firstVal && v) {
      var pct = ((v / firstVal) - 1) * 100;
      growth = "<br>Since " + firstYear + ": " +
        (pct >= 0 ? "+" : "") + pct.toFixed(0) + "%%";
    }
    var header = name ?
      "<b>" + name + "</b><br>ZIP " + zip :
      "<b>ZIP " + zip + "</b>";
    return "<div class=\\"la-heatmap-tooltip\\">" +
      header + "<br>" +
      "Year: " + year + "<br>" +
      "Typical home value: " + fmtUSD(v) +
      growth +
      "</div>";
  }

  var layer = L.geoJSON(GEOJSON, {
    style: styleFor,
    onEachFeature: function(feature, lyr) {
      lyr.bindTooltip(buildTooltip(feature), { sticky: true });
      lyr.on("mouseover", function() {
        lyr.setStyle({ weight: 2, color: "#222" });
        lyr.bringToFront();
      });
      lyr.on("mouseout", function() {
        layer.resetStyle(lyr);
      });
    }
  }).addTo(map);

  try {
    map.fitBounds(layer.getBounds(), { padding: [10, 10] });
  } catch (e) { /* ignore */ }

  var slider = document.getElementById(ID + "-slider");
  var label  = document.getElementById(ID + "-year-label");
  var btn    = document.getElementById(ID + "-play");
  var legend = document.getElementById(ID + "-legend");

  function updateLabel() {
    label.textContent = YEAR_LABELS[currentYearIdx];
  }

  function refreshAll() {
    layer.eachLayer(function(lyr) {
      lyr.setStyle(styleFor(lyr.feature));
      lyr.setTooltipContent(buildTooltip(lyr.feature));
    });
  }

  slider.addEventListener("input", function() {
    currentYearIdx = parseInt(slider.value, 10);
    updateLabel();
    refreshAll();
  });

  // Build legend swatches.
  legend.innerHTML = "";
  var legendLabel = document.createElement("span");
  legendLabel.textContent = "Typical home value:";
  legendLabel.style.marginRight = "4px";
  legend.appendChild(legendLabel);
  for (var i = 0; i < COLOR_BREAKS.length; i++) {
    var sw = document.createElement("span");
    sw.className = "swatch";
    sw.style.background = colorFor(COLOR_BREAKS[i]);
    legend.appendChild(sw);
    var t = document.createElement("span");
    t.textContent = BREAK_LABELS[i];
    t.style.marginRight = "8px";
    legend.appendChild(t);
  }

  // Play / pause animation.
  var playing = false;
  var timer = null;
  function step() {
    currentYearIdx = (currentYearIdx + 1) %% YEARS.length;
    slider.value = currentYearIdx;
    updateLabel();
    refreshAll();
  }
  function setPlaying(on) {
    playing = on;
    if (playing) {
      btn.innerHTML = "&#10074;&#10074; Pause";
      timer = setInterval(step, 700);
    } else {
      btn.innerHTML = "&#9654; Play";
      if (timer) clearInterval(timer);
      timer = null;
    }
  }
  btn.addEventListener("click", function() { setPlaying(!playing); });

  updateLabel();

  // Re-fit map when navigated back via Plotly tab visibility, etc.
  setTimeout(function() {
    map.invalidateSize();
    try { map.fitBounds(layer.getBounds(), { padding: [10, 10] }); }
    catch (e) {}
  }, 200);
})();
</script>
', widget_id, widget_id, widget_id, widget_id,
   length(years_vec) - 1L, length(years_vec) - 1L, widget_id,
   widget_id, geojson_str, values_str, years_str, year_labels_str,
   breaks_str, break_labels_str, log_min, log_max))
```

## Year-over-year change

The map above shows levels. The chart below shows growth: each line is one ZIP code, indexed to its `r year_min` value. Hover to find a specific area.

```{r yoy, fig.height=6, fig.width=10}
indexed <- zhvi_history %>%
  dplyr::group_by(zip) %>%
  dplyr::filter(any(year == year_min)) %>%
  dplyr::mutate(
    base = zhvi[year == year_min][1],
    indexed_value = zhvi / base
  ) %>%
  dplyr::ungroup()

city_lookup <- zhvi_history %>%
  dplyr::distinct(zip, city, name)

p_lines <- indexed %>%
  dplyr::left_join(city_lookup, by = c("zip", "city", "name")) %>%
  ggplot2::ggplot(ggplot2::aes(
    x = date, y = indexed_value, group = zip,
    text = paste0(
      "<b>", name, "</b><br>",
      "ZIP: ", zip, "<br>",
      "Year: ", year, "<br>",
      scales::dollar(zhvi, accuracy = 1e3),
      " (", scales::percent(indexed_value - 1, accuracy = 1),
      " vs ", year_min, ")"
    )
  )) +
  ggplot2::geom_line(alpha = 0.18, color = "#2E86C1") +
  ggplot2::scale_y_continuous(
    labels = scales::label_percent(scale = 100, accuracy = 1),
    trans = "log10"
  ) +
  ggplot2::labs(
    x = NULL,
    y = paste0("Indexed to ", year_min, " (log scale)")
  ) +
  ggplot2::theme_minimal(base_size = 12)

plotly::ggplotly(p_lines, tooltip = "text") %>%
  plotly::config(displaylogo = FALSE)
```

## ZIP-level table

<details class="table-accordion"><summary>Show table</summary>

```{r table}
table_data <- zhvi_history %>%
  dplyr::group_by(zip, name, city) %>%
  dplyr::summarise(
    `First year` = min(year),
    `First value` = scales::dollar(zhvi[year == min(year)][1], accuracy = 1e3),
    `Latest value` = scales::dollar(zhvi[year == max(year)][1], accuracy = 1e3),
    `Total growth` = scales::percent(
      (zhvi[year == max(year)][1] / zhvi[year == min(year)][1]) - 1,
      accuracy = 1
    ),
    .groups = "drop"
  ) %>%
  dplyr::rename(ZIP = zip, Neighborhood = name, City = city) %>%
  dplyr::arrange(dplyr::desc(`Latest value`))

DT::datatable(
  table_data,
  filter = "top", extensions = "Buttons",
  options = list(
    pageLength = 50, dom = "Bfrtip",
    buttons = c("copy", "csv", "excel"),
    scrollY = "400px", scrollX = TRUE
  )
)
```

</details>

---

**Methodology notes.** ZHVI is a smoothed, seasonally adjusted estimate of typical home value for the middle 33% to 67% of the market for a given region. It is not a transaction-based median sale price; it includes the full housing stock (condo/SFR), so values reflect the *typical* home rather than what is currently changing hands. Comparisons across years on a single ZIP are robust; comparisons across ZIPs reflect both market dynamics and the underlying composition of housing.

For median sale prices and price-per-square-foot at the neighborhood level, see [LA Neighborhoods: Price per Square Foot](la_real_estate.html).
