---
title: "LA Neighborhoods: Price per Square Foot"
editor_options:
  chunk_output_type: console
output:
  html_document:
    includes:
      in_header: header/header.html
      after_body: include_footer.html
---

```{r pipe, include=FALSE}
`%>%` <- magrittr::`%>%`
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.width = 10, fig.height = 8
)
```

```{r data}
la_neighborhoods <- tibble::tribble(
  ~neighborhood,           ~tier,             ~median_price, ~price_per_sqft, ~typical_sqft, ~median_rent,
  "Bel Air",               "Ultra-Luxury",     4021000,       1318,            3050,          NA,
  "Beverly Crest",         "Ultra-Luxury",     2695000,       1310,            2060,          NA,
  "Beverly Hills",         "Ultra-Luxury",     3581000,       1250,            2865,          NA,
  "Holmby Hills",          "Ultra-Luxury",     7000000,       1200,            5835,          NA,
  "Malibu (Point Dume)",   "Ultra-Luxury",     5600000,       1100,            5090,          NA,
  "Hollywood Hills West",  "Luxury",           2300000,       1209,            1905,          NA,
  "Pacific Palisades",     "Luxury",           3200000,       1050,            3050,          12225,
  "Manhattan Beach",       "Luxury",           3000000,       1150,            2610,          3500,
  "Brentwood",             "Luxury",           2648000,       1040,            2545,          4995,
  "Santa Monica",          "Luxury",           2500000,        950,            2630,          4330,
  "Los Feliz",             "Luxury",           2275000,        867,            2625,          3500,
  "Venice",                "Upper-Mid",        1851000,       1213,            1525,          3200,
  "Cheviot Hills",         "Upper-Mid",        2200000,       1145,            1920,          NA,
  "Silver Lake",           "Upper-Mid",        1675000,        848,            1975,          3500,
  "West Hollywood",        "Upper-Mid",        1400000,        800,            1750,          2700,
  "Hancock Park",          "Upper-Mid",        1900000,        780,            2435,          2500,
  "Studio City",           "Upper-Mid",        1567000,        720,            2175,          4750,
  "Sherman Oaks",          "Upper-Mid",        1450000,        680,            2130,          3600,
  "Culver City",           "Upper-Mid",        1303000,        750,            1735,          2600,
  "Pasadena",              "Mid-Range",        1200000,        680,            1765,          2200,
  "Echo Park",             "Mid-Range",        1100000,        720,            1530,          2400,
  "Atwater Village",       "Mid-Range",        1150000,        700,            1645,          2200,
  "Highland Park",         "Mid-Range",        1199000,        936,            1280,          2100,
  "Eagle Rock",            "Mid-Range",        1050000,        700,            1500,          2000,
  "Encino",                "Mid-Range",        1300000,        580,            2240,          5495,
  "Westchester",           "Mid-Range",        1200000,        650,            1845,          2300,
  "North Hollywood",       "Affordable",        850000,        550,            1545,          2997,
  "Boyle Heights",         "Affordable",        742000,        471,            1575,          1500,
  "El Sereno",             "Affordable",        750000,        520,            1440,          1600,
  "Koreatown",             "Affordable",        680000,        580,            1170,          2200,
  "Eastside LA",           "Affordable",        749000,        624,            1200,          1700,
  "South Los Angeles",     "Affordable",        682000,        456,            1495,          1400,
  "Pico-Union",            "Affordable",        650000,        453,            1435,          1500,
  "Historic South Central","Affordable",        550000,        425,            1295,          1300,
  "South Park (DTLA)",     "Affordable",        500000,        419,            1195,          1400
) %>%
  dplyr::mutate(
    tier = factor(tier, levels = c(
      "Ultra-Luxury", "Luxury", "Upper-Mid", "Mid-Range", "Affordable"
    )),
    sqft_per_million = round(1e6 / price_per_sqft)
  )

tier_colors <- c(
  "Ultra-Luxury" = "#8B0000",
  "Luxury"       = "#D4760A",
  "Upper-Mid"    = "#2E86C1",
  "Mid-Range"    = "#27AE60",
  "Affordable"   = "#7D3C98"
)
```

How much space does your dollar buy across Los Angeles? This page compares 35 neighborhoods by **price per square foot** --- the ratio of median home price to typical home size. Neighborhoods are grouped into five tiers based on median home price:

- **Ultra-Luxury** (>$3M): Bel Air, Beverly Hills, Holmby Hills, Malibu
- **Luxury** ($2M--$3M): Brentwood, Santa Monica, Pacific Palisades
- **Upper-Mid** ($1.3M--$2M): Venice, Silver Lake, Sherman Oaks
- **Mid-Range** ($900K--$1.3M): Highland Park, Pasadena, Eagle Rock
- **Affordable** (<$900K): Boyle Heights, North Hollywood, Koreatown

Data cross-referenced from [Zillow ZHVI](https://www.zillow.com/research/data/) (Feb 2026), [Redfin](https://www.redfin.com/city/11203/CA/Los-Angeles/housing-market) median sale prices (Feb 2026), and [American Home Field](https://americanhomefield.com/) $/sqft study.

## Price per Square Foot by Neighborhood {.tabset}

### Bar Chart

```{r bar_chart, fig.height=10}
la_neighborhoods %>%
  dplyr::mutate(neighborhood = forcats::fct_reorder(neighborhood, price_per_sqft)) %>%
  ggplot2::ggplot(ggplot2::aes(x = neighborhood, y = price_per_sqft, fill = tier)) +
  ggplot2::geom_col(width = 0.7) +
  ggplot2::geom_text(
    ggplot2::aes(label = scales::dollar(price_per_sqft, accuracy = 1)),
    hjust = -0.1, size = 3, color = "grey30"
  ) +
  ggplot2::coord_flip(ylim = c(0, max(la_neighborhoods$price_per_sqft) * 1.15)) +
  ggplot2::scale_fill_manual(values = tier_colors) +
  ggplot2::scale_y_continuous(labels = scales::label_dollar()) +
  ggplot2::labs(
    x = NULL, y = "Median Price per Sq Ft", fill = "Price Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    legend.position = "top",
    panel.grid.major.y = ggplot2::element_blank()
  )
```

### Interactive Table

```{r table}
la_neighborhoods %>%
  dplyr::arrange(dplyr::desc(price_per_sqft)) %>%
  dplyr::transmute(
    Neighborhood = neighborhood,
    Tier = tier,
    `Median Price` = scales::dollar(median_price),
    `$/sqft` = scales::dollar(price_per_sqft),
    `Typical sqft` = scales::comma(typical_sqft),
    `$1M Buys (sqft)` = scales::comma(sqft_per_million),
    `Median Rent` = dplyr::if_else(
      is.na(median_rent), "---", scales::dollar(median_rent)
    )
  ) %>%
  DT::datatable(
    filter = "top", extensions = "Buttons",
    options = list(
      paging = FALSE, dom = "Bfrtip",
      buttons = c("copy", "csv", "excel"),
      scrollY = "500px", scrollX = TRUE
    )
  )
```

## What Does $1 Million Buy You? {.tabset}

### Bar Chart

```{r million_chart, fig.height=10}
la_neighborhoods %>%
  dplyr::mutate(neighborhood = forcats::fct_reorder(neighborhood, sqft_per_million)) %>%
  ggplot2::ggplot(ggplot2::aes(x = neighborhood, y = sqft_per_million, fill = tier)) +
  ggplot2::geom_col(width = 0.7) +
  ggplot2::geom_text(
    ggplot2::aes(label = paste0(scales::comma(sqft_per_million), " sqft")),
    hjust = -0.1, size = 3, color = "grey30"
  ) +
  ggplot2::coord_flip(ylim = c(0, max(la_neighborhoods$sqft_per_million) * 1.15)) +
  ggplot2::scale_fill_manual(values = tier_colors) +
  ggplot2::scale_y_continuous(labels = scales::label_comma(suffix = " sqft")) +
  ggplot2::labs(
    x = NULL, y = "Square Feet per $1M", fill = "Price Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    legend.position = "top",
    panel.grid.major.y = ggplot2::element_blank()
  )
```

### Scatter: Price vs Size

```{r scatter, fig.height=7}
la_neighborhoods %>%
  ggplot2::ggplot(ggplot2::aes(x = typical_sqft, y = median_price, color = tier)) +
  ggplot2::geom_point(size = 4, alpha = 0.85) +
  ggrepel::geom_text_repel(
    ggplot2::aes(label = neighborhood),
    size = 3, max.overlaps = 25, segment.color = "grey70"
  ) +
  ggplot2::scale_x_continuous(labels = scales::label_comma(suffix = " sqft")) +
  ggplot2::scale_y_continuous(labels = scales::label_dollar(scale_cut = scales::cut_short_scale())) +
  ggplot2::scale_color_manual(values = tier_colors) +
  ggplot2::labs(
    x = "Typical Home Size (sq ft)", y = "Median Sale Price", color = "Price Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(legend.position = "top")
```

## Price-to-Rent Ratio

The price-to-rent ratio divides median home price by annual rent. Under 15 favors buying; 15--20 is a toss-up; over 20 favors renting. Every LA neighborhood lands above 20.

```{r rent_ratio, fig.height=8}
la_neighborhoods %>%
  dplyr::filter(!is.na(median_rent)) %>%
  dplyr::mutate(
    annual_rent = median_rent * 12,
    price_to_rent = median_price / annual_rent,
    neighborhood = forcats::fct_reorder(neighborhood, price_to_rent)
  ) %>%
  ggplot2::ggplot(ggplot2::aes(x = neighborhood, y = price_to_rent, fill = tier)) +
  ggplot2::geom_col(width = 0.7) +
  ggplot2::geom_text(
    ggplot2::aes(label = round(price_to_rent, 1)),
    hjust = -0.1, size = 3, color = "grey30"
  ) +
  ggplot2::geom_hline(yintercept = 15, linetype = "dashed", color = "#27AE60", linewidth = 0.6) +
  ggplot2::geom_hline(yintercept = 20, linetype = "dashed", color = "#E74C3C", linewidth = 0.6) +
  ggplot2::coord_flip(ylim = c(0, 80)) +
  ggplot2::scale_fill_manual(values = tier_colors) +
  ggplot2::annotate("text", y = 15, x = 0.5, label = "Buy favors (<15)",
                    hjust = 1.1, size = 3, color = "#27AE60") +
  ggplot2::annotate("text", y = 20, x = 0.5, label = "Rent favors (>20)",
                    hjust = -0.1, size = 3, color = "#E74C3C") +
  ggplot2::labs(
    x = NULL,
    y = "Price-to-Rent Ratio (median price / annual rent)",
    fill = "Price Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    legend.position = "top",
    panel.grid.major.y = ggplot2::element_blank()
  )
```

---

**Sources:** [Zillow Home Value Index](https://www.zillow.com/research/data/) (Feb 2026), [Redfin Housing Market](https://www.redfin.com/city/11203/CA/Los-Angeles/housing-market) (Feb 2026), [American Home Field](https://americanhomefield.com/) $/sqft study, [Redfin Rental Market](https://www.redfin.com/city/11203/CA/Los-Angeles/rental-market) (Aug 2025). LA citywide median: $1.01M, $633/sqft per Redfin.
