---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r ci-skip-gs4_readme, include = FALSE}
# This file only documents Google Sheets usage, doesn't actually call APIs
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# Cavan's Website

<!-- badges: start -->
![](https://github.com/cavandonohoe/cavandonohoe.github.io/actions/workflows/pages-rmarkdown.yml/badge.svg)
<!-- badges: end -->

The goal of my website is to display some fun side projects I have been working on. Movie data intrigues me, so I scraped IMDb to use some of their clean data. I also have my résumé here since it's easier to share a link than an attachment.

Check out my website at [https://cavandonohoe.github.io](https://cavandonohoe.github.io)

Highlights:

* Data projects and visualizations (IMDb, box office, ratings)
* Personal pages (blog, reunion, schedule)
* CV and learning resources

Movies & data:

* [IMDb TV Series](https://cavandonohoe.github.io/imdb_top_250_tv_series.html)
* [Best Picture vs Top 1000 Box Office](https://cavandonohoe.github.io/best_pic_vs_top_1000.html)
* [Top 250 Rated Movies with Rotten Tomato Scores](https://cavandonohoe.github.io/top_250_imdb_with_rt.html)
* [IMDb Series Rating Plots](https://cavandonohoe.github.io/imdb_rating_plot.html)
* [Best Picture Nominees](https://cavandonohoe.github.io/best_picture_nominees.html)
* [Director Box Office Trajectory](https://cavandonohoe.github.io/director_box_office_trajectory.html)
* [Rent vs. Buy Calculator](https://cavandonohoe.shinyapps.io/rent-vs-buy/) ([source](https://github.com/cavandonohoe/rent-vs-buy))
* [Confederate Statue Analysis](https://github.com/cavandonohoe/confederate_statues)
* [Sierpinski Triangle](https://cavandonohoe.github.io/sierpinski_triangle.html)
* [Airplane Boarding Simulation](https://cavandonohoe.github.io/airplane_boarding.html)

Personal:

* [Kelly Schedule](https://cavandonohoe.github.io/firefighter_schedule.html)
* [TOHS Class of 2012 Reunion](https://cavandonohoe.github.io/tohs_reunion.html)
* [Blog](https://cavandonohoe.github.io/my_travels.html)

Learning:

* [Learn R (Course on Learning R)](https://cavandonohoe.github.io/learn_r.html)

CV:

* [CV](https://cavandonohoe.github.io/cv.html)

## Automation

The repo runs the following on a schedule via GitHub Actions. Anything
manual is a click away on the Actions tab.

### Site build & previews

* `pages-branch-previews.yml` -- main deploy + per-PR preview deploy
  and cleanup. This is what publishes the site (pushes to the
  `gh-pages` branch).
* `pages-rmarkdown.yml` -- daily safety-net build with DESCRIPTION /
  `library()` validation and gs4 guard checks. Does **not** deploy
  (deploy is owned by `pages-branch-previews.yml` to avoid Pages
  deployment races).

### Data refreshes (commits to `main`)

* `update_imdb_ratings.yml` -- weekly IMDb episode ratings refresh.
* `update_top1000_box_office.yml` -- monthly Best Picture + Top 1000
  box office refresh.
* `update_top250_with_rt.yml` -- monthly Top 250 + Rotten Tomatoes
  refresh.
* `update_director_filmographies.yml` -- monthly TMDb director
  filmography refresh for the Director Box Office Trajectory page.
* `update_sp500.yml` -- weekly VOO price refresh.
* `update_us_rentals.yml` -- monthly Zillow ZHVI/ZORI refresh.
* `update_mymaps_export.yml` -- weekly Google My Maps export.

### Data refreshes (open PR)

* `update_best_picture_winners.yml` -- annual Oscar refresh.
* `rebuild_cv.yml` -- weekly CV PDF/DOCX rebuild from Google Sheet,
  opens a PR if anything changed.
* `refresh-sitemap.yml` -- weekly `sitemap.xml` regeneration.

### Site health

* `link-check.yml` -- weekly + on-PR broken-link check (lychee).
* `pa11y.yml` -- weekly accessibility audit (pa11y-ci).
* `lighthouse.yml` -- two modes: on-PR audit of the deployed preview URL
  with a score comment, plus a post-build audit of the production
  artifact after each successful main build.
* `repo-size.yml` -- weekly + on-PR repo-size report.

### Code quality

* `lint.yml` -- on-PR lintr scan of changed R/Rmd files.
* `typos.yml` -- on-PR spell check.
* `pr-smoke.yml` -- on-PR fast render of only the changed Rmds.

### Repo hygiene

* `.github/dependabot.yml` -- weekly bump of pinned GitHub Actions.
* `labeler.yml` -- auto-label PRs by changed paths.

## Activity over time

The chart below summarizes commit activity on this repo: monthly
commit count and lines changed (insertions + deletions), split by
my own commits vs. automated commits from the scheduled data
refresh workflows above. Regenerated each time `README.Rmd` is
knit.

```{r commit-activity, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 14, fig.height = 8, fig.path = "images/readme/", dev = "png", dpi = 110, fig.alt = "Two bar charts showing monthly commits and monthly lines changed (insertions + deletions) for this repository, split by Cavan's own commits vs. scheduled-bot commits."}
suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(tidyr)
  library(lubridate)
  library(scales)
  library(patchwork)
})

raw <- system(
  "git log --no-merges --pretty=format:'COMMIT|%H|%at|%an' --shortstat",
  intern = TRUE
)

commits <- list()
current <- NULL
for (line in raw) {
  if (startsWith(line, "COMMIT|")) {
    if (!is.null(current)) commits[[length(commits) + 1L]] <- current
    parts <- strsplit(line, "|", fixed = TRUE)[[1L]]
    current <- list(
      sha = parts[2L],
      ts = as.numeric(parts[3L]),
      author = parts[4L],
      ins = 0L,
      del = 0L
    )
  } else if (nchar(line) > 0 && !is.null(current)) {
    ins_m <- regmatches(line, regexpr("([0-9]+) insertion", line))
    del_m <- regmatches(line, regexpr("([0-9]+) deletion", line))
    if (length(ins_m)) current$ins <- as.integer(sub(" insertion.*", "", ins_m))
    if (length(del_m)) current$del <- as.integer(sub(" deletion.*", "", del_m))
  }
}
if (!is.null(current)) commits[[length(commits) + 1L]] <- current

df <- dplyr::bind_rows(commits) %>%
  dplyr::mutate(
    date = as.POSIXct(ts, origin = "1970-01-01", tz = "UTC"),
    month = lubridate::floor_date(as.Date(date), "month"),
    # All non-bot commits are mine; STEPHANIE HUSSEY is just me on her laptop.
    is_bot = grepl("\\[bot\\]|github-actions", author, ignore.case = TRUE),
    author_type = factor(
      ifelse(is_bot, "Automated (bots)", "Cavan"),
      levels = c("Cavan", "Automated (bots)")
    ),
    lines_changed = ins + del
  )

month_seq <- seq(min(df$month), max(df$month), by = "month")
monthly <- tidyr::expand_grid(
  month = month_seq,
  author_type = levels(df$author_type)
) %>%
  dplyr::left_join(
    df %>%
      dplyr::group_by(month, author_type) %>%
      dplyr::summarise(
        commits = dplyr::n(),
        lines_changed = sum(lines_changed),
        .groups = "drop"
      ),
    by = c("month", "author_type")
  ) %>%
  tidyr::replace_na(list(commits = 0L, lines_changed = 0L)) %>%
  dplyr::mutate(
    author_type = factor(author_type, levels = c("Cavan", "Automated (bots)"))
  )

palette <- c("Cavan" = "#2c7fb8", "Automated (bots)" = "#f0a202")

theme_readme <- ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    plot.title = ggplot2::element_text(face = "bold"),
    plot.title.position = "plot",
    panel.grid.minor = ggplot2::element_blank(),
    legend.position = "top",
    axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust = 1)
  )

p_commits <- ggplot2::ggplot(
  monthly,
  ggplot2::aes(x = month, y = commits, fill = author_type)
) +
  ggplot2::geom_col(width = 25) +
  ggplot2::scale_fill_manual(values = palette, name = NULL) +
  ggplot2::scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %Y",
    expand = ggplot2::expansion(mult = c(0.01, 0.01))
  ) +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::labs(
    title = "Commits per month",
    x = NULL,
    y = "Commits"
  ) +
  theme_readme

p_lines <- ggplot2::ggplot(
  monthly %>% dplyr::filter(lines_changed > 0),
  ggplot2::aes(x = month, y = lines_changed, fill = author_type)
) +
  ggplot2::geom_col(
    position = ggplot2::position_dodge2(width = 25, preserve = "single"),
    width = 25
  ) +
  ggplot2::scale_fill_manual(values = palette, name = NULL) +
  ggplot2::scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %Y",
    limits = range(month_seq),
    expand = ggplot2::expansion(mult = c(0.01, 0.01))
  ) +
  ggplot2::scale_y_continuous(
    trans = "log10",
    labels = scales::label_number(scale_cut = scales::cut_short_scale())
  ) +
  ggplot2::labs(
    title = "Lines changed per month (insertions + deletions, log scale)",
    x = NULL,
    y = "Lines changed"
  ) +
  theme_readme

(p_commits / p_lines) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "top")
```

```{r commit-activity-cumulative, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 14, fig.height = 7, fig.path = "images/readme/", dev = "png", dpi = 110, fig.alt = "Two cumulative line charts showing total commits and total lines changed (insertions + deletions) over time, split by Cavan's own commits vs. scheduled-bot commits."}
cumulative <- monthly %>%
  dplyr::arrange(author_type, month) %>%
  dplyr::group_by(author_type) %>%
  dplyr::mutate(
    cum_commits = cumsum(commits),
    cum_lines = cumsum(lines_changed)
  ) %>%
  dplyr::ungroup()

p_cum_commits <- ggplot2::ggplot(
  cumulative,
  ggplot2::aes(x = month, y = cum_commits, colour = author_type)
) +
  ggplot2::geom_line(linewidth = 0.9) +
  ggplot2::geom_point(size = 0.8) +
  ggplot2::scale_colour_manual(values = palette, name = NULL) +
  ggplot2::scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %Y",
    expand = ggplot2::expansion(mult = c(0.01, 0.01))
  ) +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::labs(
    title = "Cumulative commits",
    x = NULL,
    y = "Cumulative commits"
  ) +
  theme_readme

p_cum_lines <- ggplot2::ggplot(
  cumulative,
  ggplot2::aes(x = month, y = cum_lines, colour = author_type)
) +
  ggplot2::geom_line(linewidth = 0.9) +
  ggplot2::geom_point(size = 0.8) +
  ggplot2::scale_colour_manual(values = palette, name = NULL) +
  ggplot2::scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %Y",
    expand = ggplot2::expansion(mult = c(0.01, 0.01))
  ) +
  ggplot2::scale_y_continuous(
    labels = scales::label_number(scale_cut = scales::cut_short_scale())
  ) +
  ggplot2::labs(
    title = "Cumulative lines changed (insertions + deletions)",
    x = NULL,
    y = "Cumulative lines changed"
  ) +
  theme_readme

(p_cum_commits / p_cum_lines) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "top")
```

```{r commit-activity-summary, echo = FALSE, message = FALSE, results = "asis"}
total_commits <- nrow(df)
total_lines <- sum(df$lines_changed)
total_cavan <- sum(df$author_type == "Cavan")
total_bots <- sum(df$author_type == "Automated (bots)")
first_commit <- min(df$date)
cat(sprintf(
  paste0(
    "*%s commits since %s &mdash; %s by me, %s by scheduled bots ",
    "&mdash; touching %s lines in total.*\n"
  ),
  format(total_commits, big.mark = ","),
  format(first_commit, "%b %Y"),
  format(total_cavan, big.mark = ","),
  format(total_bots, big.mark = ","),
  format(total_lines, big.mark = ",")
))
```

## Setup and Local Development

Local development, prerequisites, build steps, lint/test commands, and PR conventions all live in [CONTRIBUTING.md](CONTRIBUTING.md).

### Project Structure

- `*.Rmd` files: Individual pages of the website
- `_site.yml`: Site configuration and navigation
- `_common.R`: Common R code (Google auth setup for CI)
- `header/header.html`: HTML header with meta tags and favicons
- `include_footer.html`: Footer with contact links
- `css/footer.css`: Footer styling
- `.github/workflows/`: GitHub Actions workflows for CI/CD
- `data/`: Data files used by various pages
- `scripts/`: Utility scripts (e.g., ensuring gs4 guards)

### Deployment

The site is automatically deployed to GitHub Pages via GitHub Actions when changes are pushed to the `main` branch. See the Automation section above for the full list of workflows.
