---
title: "Lesson 6: Reproducible Reports with R Markdown"
author: "Cavan Donohoe"
description: "Write code and prose together, knit to HTML or PDF, and share something a colleague can rerun."
output:
  html_document:
    toc: true
    toc_float: true
    includes:
      in_header: header/header.html
      after_body: include_footer.html
lang: "en"
---

<link rel="stylesheet" href="/css/learn-r.css">

<div class="course-progress">
Lesson 6 of 7 · <a href="learn_r.html">Course overview</a>
</div>

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "#>", message = FALSE, warning = FALSE)
```

# Reproducible Reports with R Markdown

R Markdown is how R programmers write reports. You combine narrative prose, R code, and the *output* of that code (tables, plots, model summaries) into a single document. When you "knit" it, R runs all the code from scratch and produces an HTML, PDF, or Word file.

Every lesson in this course was written in R Markdown. So is the page you're reading.

The big win: **the report is the source of truth**. There's no copy-pasting numbers from R into Word, no screenshots of plots that you forgot to update. You change the data or the code, you re-knit, and everything updates.

## The anatomy of an Rmd file

An R Markdown file (`.Rmd`) has three kinds of content:

1. A **YAML header** at the top, between `---` lines, that controls metadata (title, author, output format).
2. **Markdown prose** — same as on GitHub, headings with `#`, lists with `-`, links with `[text](url)`.
3. **Code chunks** — R code in fenced blocks, marked with `` ```{r} ``.

A minimal example:

````
---
title: "My First Report"
author: "Your Name"
output: html_document
---

## Setup

```{r}`r ''`
library(dplyr)
```

The mtcars dataset has `r '\x60r nrow(mtcars)\x60'` rows.

```{r}`r ''`
mtcars |> count(cyl)
```
````

When you knit that, R runs the code chunks, captures their output, splices in the inline values (the <code>&#96;r nrow(mtcars)&#96;</code> syntax inside the prose), and renders the whole thing into HTML.

## Creating an Rmd in RStudio

In RStudio, **File → New File → R Markdown…** opens a dialog. Pick a title, an output format (HTML is the default), and click OK. RStudio creates a template you can knit immediately by clicking the **Knit** button in the toolbar (or pressing <kbd>Cmd</kbd>/<kbd>Ctrl</kbd> + <kbd>Shift</kbd> + <kbd>K</kbd>).

The first time you knit, it will install whatever packages are missing. Be patient.

## Code chunks

A code chunk looks like this in your `.Rmd` source:

````
```{r}`r ''`
library(dplyr)
mtcars |> count(cyl)
```
````

The `{r}` says "this is R" — there are also chunks for Python, SQL, and a few other languages, but you'll mostly use R.

Inside the curly braces you can pass **chunk options** that control how the chunk behaves:

````
```{r my-chunk-name, echo=FALSE, fig.width=8}`r ''`
plot(mtcars$wt, mtcars$mpg)
```
````

The most useful chunk options:

| Option | What it does |
|--------|--------------|
| `echo = FALSE` | Hide the code, show the output (good for figures in a report) |
| `eval = FALSE` | Show the code but don't run it (good for "here's how you'd do it" examples) |
| `include = FALSE` | Run the code, but show nothing (good for setup chunks) |
| `message = FALSE` | Suppress package startup messages |
| `warning = FALSE` | Suppress warnings |
| `fig.width`, `fig.height` | Set the size of any plots produced |
| `cache = TRUE` | Cache the chunk's results to disk; only re-run if the code changes |

You can set defaults for the whole document in a setup chunk near the top:

````
```{r setup, include=FALSE}`r ''`
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 4
)
```
````

## Inline R

The <code>&#96;r nrow(mtcars)&#96;</code> syntax inside the prose runs an R expression and splices the result into the text. It's how you keep numbers in your prose in sync with the data:

```
The dataset has `r '\x60r nrow(mtcars)\x60'` rows and
`r '\x60r ncol(mtcars)\x60'` columns. The mean MPG is
`r '\x60r round(mean(mtcars$mpg), 1)\x60'`.
```

When this knits, it produces:

> The dataset has `r nrow(mtcars)` rows and `r ncol(mtcars)` columns. The mean MPG is `r round(mean(mtcars$mpg), 1)`.

If the data changes, the numbers change. No manual edits.

## Output formats

The `output:` line in the YAML header picks the format. Common ones:

```yaml
output: html_document
```

```yaml
output: pdf_document    # requires LaTeX (TinyTeX is easiest: tinytex::install_tinytex())
```

```yaml
output: word_document
```

You can also customize the HTML output:

```yaml
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: cosmo
    df_print: paged
```

- `toc: true` — generate a table of contents from your headings.
- `toc_float: true` — make it sticky on the side as you scroll.
- `code_folding: show` — readers can hide code chunks with one click.
- `theme: cosmo` — Bootstrap theme. Try `"flatly"`, `"journal"`, `"cosmo"`, etc.
- `df_print: paged` — pretty pagination for data frames.

R Markdown can also produce slides (`output: ioslides_presentation` or `revealjs::revealjs_presentation`), websites, books, and dashboards. The package ecosystem here is huge.

## A tiny worked example

Here's a complete, working `.Rmd` file end to end. Save this as `cars_report.Rmd` and knit it:

````
---
title: "Cars Report"
author: "You"
date: "`r '\x60r Sys.Date()\x60'`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}`r ''`
knitr::opts_chunk$set(message = FALSE, warning = FALSE, fig.width = 6, fig.height = 4)
library(dplyr)
library(ggplot2)
```

## Overview

This report uses the built-in `mtcars` dataset, which has
`r '\x60r nrow(mtcars)\x60'` rows.

## Distribution of MPG

```{r}`r ''`
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(bins = 10, fill = "steelblue", color = "white") +
  theme_minimal()
```

## Average MPG by cylinder count

```{r}`r ''`
mtcars |>
  group_by(cyl) |>
  summarise(mean_mpg = mean(mpg), n = n(), .groups = "drop")
```

## Conclusion

Heavier, more-cylindered cars get worse mileage. Shocking.
````

Knit it once. Then change something — say, swap `mean` for `median`. Knit again. The whole report updates without you touching the prose.

## Tips for real-world reports

- **Keep one Rmd per analysis.** Don't try to put your whole life in one document.
- **Put setup at the top.** All `library()` calls and `opts_chunk$set()` go in a single chunk near the top, often named `setup`.
- **Name your chunks.** `{r my-thing}` instead of `{r}`. Makes errors much easier to find ("error in chunk *my-thing*").
- **Cache slow chunks.** If something takes 30 seconds to run, set `cache = TRUE` so you don't re-run it on every knit.
- **Don't fight the layout.** If you find yourself wanting fine control over fonts and margins, you might want a different tool (`quarto`, LaTeX, Word). R Markdown is amazing for "data-driven document," less so for "perfect typography."
- **Quarto is the next-generation R Markdown.** Same idea, supports more languages, slightly better defaults. If you're starting fresh in 2025+, [Quarto](https://quarto.org/) is worth knowing about. The mental model is identical.

<div class="callout callout-note">
<div class="callout-title">📝 Quarto vs. R Markdown</div>
Quarto (`.qmd` files) is essentially R Markdown's successor, built by the same team. It supports R, Python, Julia, and Observable in the same document, and has nicer defaults for things like cross-references. Everything you learn here transfers directly. For a personal site or one-off analysis, R Markdown is still totally fine.
</div>

<div class="exercise">
<div class="exercise-title">✏️ Exercise 6.1 — Your first knitted document</div>

Create a new R Markdown file in RStudio (File → New File → R Markdown). Replace the template content with:

- A title and author in the YAML header.
- A heading "About the data."
- A short paragraph that uses inline R to report the number of rows in `mtcars`.
- A code chunk that prints `summary(mtcars)`.
- A code chunk that plots `mtcars$wt` vs. `mtcars$mpg`.

Knit it to HTML. You should get a report with a working table of contents.

<details>
<summary>Show solution</summary>

````
---
title: "Cars Quick Look"
author: "Your Name"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}`r ''`
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## About the data

The `mtcars` dataset has `r '\x60r nrow(mtcars)\x60'` rows and
`r '\x60r ncol(mtcars)\x60'` columns.

## Summary

```{r}`r ''`
summary(mtcars)
```

## Weight vs. MPG

```{r}`r ''`
plot(mtcars$wt, mtcars$mpg,
  xlab = "Weight (1000 lbs)", ylab = "MPG",
  main = "Heavier cars get worse mileage")
```
````

</details>
</div>

<div class="exercise">
<div class="exercise-title">✏️ Exercise 6.2 — Hide the code</div>

Take the report from Exercise 6.1 and modify it so that the **code is hidden** in the knitted HTML — only the output (the summary table and the plot) shows. Add a single line so a reader can still toggle code on if they want.

<details>
<summary>Show solution</summary>

Two changes:

1. Add `echo=FALSE` to the global chunk options in the setup chunk.
2. Add `code_folding: show` (or `hide`) to the YAML to put a toggle in the output.

```yaml
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
```

```r
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

</details>
</div>

## What's next

You now know how to share an analysis as a real document. The last lesson is the capstone: a full, end-to-end project that puts every previous lesson together — load data, manipulate it, plot it, model it, and write it up.

<div class="lesson-nav">
  <a class="nav-prev" href="statistics_with_r.html">
    <span class="lesson-nav-label">← Previous</span>
    <span class="lesson-nav-title">Lesson 5: Statistics with R</span>
  </a>
  <a class="nav-next" href="capstone_r.html">
    <span class="lesson-nav-label">Next →</span>
    <span class="lesson-nav-title">Lesson 7: Capstone Project</span>
  </a>
</div>
