---
title: "SDTM with R: Clinical Trial Data Standards"
author: "Cavan Donohoe"
description: "What SDTM is, why clinical trials use it, and how to build tidy SDTM domains (DM, AE, VS, LB) from raw data with reusable R functions."
output:
  html_document:
    toc: true
    toc_float: true
    includes:
      in_header: header/header.html
      after_body: include_footer.html
lang: "en"
---

<link rel="stylesheet" href="/css/learn-r.css">

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.width = 6.5,
  fig.height = 4.2,
  dpi = 100
)
options(scipen = 999, digits = 4)
```

# SDTM with R

Every drug and diagnostic that gets submitted to the FDA (and most other
regulators) arrives with its clinical trial data in a very specific shape.
That shape is **SDTM** ã the Study Data Tabulation Model, a CDISC standard.
If you work in clinical data science, biostatistics, or statistical
programming, SDTM is the lingua franca. This page is a from-scratch
introduction: what it is, why it exists, and how to build SDTM-style
datasets in R using a handful of small, reusable functions.

You don't need any clinical background. If you've done a little tidyverse
(see the [Learn R course](learn_r.html)), you can follow everything here.

## What is SDTM?

Imagine every pharma company invented its own column names, table layouts,
and coding conventions for their trial data. A reviewer at the FDA would
have to relearn the data from scratch for every single submission. SDTM
exists to stop that.

**SDTM standardizes the *content and structure* of the raw, collected trial
data** so that any reviewer, any validation tool, and any downstream program
knows exactly where to look. It defines:

- A fixed set of **domains** ã one per type of data collected. Each has a
  two-letter code.
- Standard **variable names** within each domain (again, mostly two-to-eight
  letter codes).
- Rules for how rows, subjects, visits, and timing are represented.

<div class="callout callout-tip">
<div class="callout-title">ÞÀ The mental model</div>
SDTM is "one tidy table per kind of thing that happened in the trial."
Demographics is one table. Adverse events is another. Vital signs is
another. Every table shares a common backbone (subject ID, study ID) so they
can be joined back together.
</div>

### The domains you'll see most often

| Code | Domain | One row per... |
|------|--------|----------------|
| `DM` | Demographics | subject |
| `AE` | Adverse Events | adverse event a subject had |
| `VS` | Vital Signs | vital-sign measurement |
| `LB` | Laboratory | lab test result |
| `CM` | Concomitant Meds | medication a subject took |
| `EX` | Exposure | dose of study drug administered |

There are dozens more, but if you understand these you understand the model.

### The naming conventions

SDTM variable names follow a pattern. Once you see it, the columns stop
looking like alphabet soup:

- **`--SEQ`** ã a sequence number, unique within a subject and domain
  (`AESEQ`, `VSSEQ`).
- **`--TESTCD` / `--TEST`** ã the short code and long name of what was
  measured (`VSTESTCD = "SYSBP"`, `VSTEST = "Systolic Blood Pressure"`).
- **`--ORRES` / `--ORRESU`** ã the original result as collected, and its unit
  (`VSORRES = "120"`, `VSORRESU = "mmHg"`).
- **`--STRESC` / `--STRESN`** ã the result standardized to a character and a
  numeric form.
- **`--DTC`** ã a date/time in ISO 8601 text (`AESTDTC = "2023-04-12"`).
- **`STUDYID` / `USUBJID`** ã the study ID and the *unique* subject ID, the
  keys that tie every domain together.

The leading `--` is a placeholder for the domain code. So the "test code"
variable is `VSTESTCD` in Vital Signs and `LBTESTCD` in Labs.

## Some raw data to start from

In real life the raw data comes out of an electronic data capture (EDC)
system and looks nothing like SDTM. Our job is to reshape it. Let's fabricate
a tiny, messy "raw" export so we have something concrete to transform.

```{r}
library(dplyr)
library(tidyr)

set.seed(42)

raw_demographics <- tibble::tribble(
  ~subject, ~site, ~birth_date,  ~sex,     ~race,    ~arm,
  "001",    "01",  "1968-03-11", "Male",   "White",  "Active",
  "002",    "01",  "1975-11-30", "Female", "Asian",  "Placebo",
  "003",    "02",  "1954-07-02", "Female", "White",  "Active",
  "004",    "02",  "1981-01-19", "Male",   "Black",  "Placebo",
  "005",    "03",  "1990-09-25", "Male",   "White",  "Active"
)

raw_demographics
```

A raw vital-signs export, in "long" form, one measurement per row:

```{r}
raw_vitals <- tibble::tribble(
  ~subject, ~visit,      ~measure, ~value, ~units, ~date,
  "001",    "Screening", "SYSBP",  128,    "mmHg", "2023-01-05",
  "001",    "Screening", "DIABP",  82,     "mmHg", "2023-01-05",
  "001",    "Week 4",    "SYSBP",  122,    "mmHg", "2023-02-02",
  "001",    "Week 4",    "DIABP",  79,     "mmHg", "2023-02-02",
  "002",    "Screening", "SYSBP",  141,    "mmHg", "2023-01-06",
  "002",    "Screening", "DIABP",  90,     "mmHg", "2023-01-06",
  "003",    "Screening", "SYSBP",  118,    "mmHg", "2023-01-09",
  "003",    "Week 4",    "SYSBP",  115,    "mmHg", "2023-02-07"
)

raw_vitals
```

And a raw adverse-event log:

```{r}
raw_ae <- tibble::tribble(
  ~subject, ~event,        ~severity,  ~start,       ~end,
  "001",    "Headache",    "Mild",     "2023-01-20", "2023-01-22",
  "001",    "Nausea",      "Moderate", "2023-02-05", NA,
  "003",    "Fatigue",     "Mild",     "2023-01-15", "2023-01-30",
  "004",    "Dizziness",   "Severe",   "2023-01-25", "2023-01-26"
)

raw_ae
```

## Building the DM (Demographics) domain

DM is the anchor domain: exactly one row per subject. Everything else joins
back to it. Building it means renaming to SDTM variables, constructing the
`USUBJID` key, and deriving `AGE` from the birth date.

Here's a reusable function:

```{r}
make_dm <- function(raw, studyid) {
  raw |>
    mutate(
      STUDYID = studyid,
      DOMAIN = "DM",
      SITEID = site,
      SUBJID = subject,
      USUBJID = paste(studyid, site, subject, sep = "-"),
      RFSTDTC = birth_date,                       # reference start (placeholder)
      BRTHDTC = birth_date,
      AGE = as.integer(
        floor(as.numeric(Sys.Date() - as.Date(birth_date)) / 365.25)
      ),
      AGEU = "YEARS",
      SEX = dplyr::recode(sex, "Male" = "M", "Female" = "F"),
      RACE = toupper(race),
      ARM = arm,
      ACTARM = arm
    ) |>
    select(
      STUDYID, DOMAIN, USUBJID, SUBJID, SITEID,
      BRTHDTC, AGE, AGEU, SEX, RACE, ARM, ACTARM
    )
}

dm <- make_dm(raw_demographics, studyid = "DEMO01")
dm
```

A few SDTM conventions on display here:

- **`USUBJID` must be unique across the whole study**, so we build it from
  study + site + subject. Two sites could both have a subject "001"; the
  `USUBJID` keeps them distinct.
- **`SEX` is coded** to controlled terminology (`M`/`F`), not free text.
- **`AGE` is derived**, not collected. SDTM is full of derived variables like
  this.

<div class="callout callout-warning">
<div class="callout-title">ã ÿ¡ Controlled terminology</div>
Real SDTM uses CDISC **controlled terminology** ã fixed, published code lists
for values like `SEX`, `RACE`, and units. You can't put "Male" in `SEX`; it
has to be `M`. Here we're doing a simplified `recode()`, but in a real study
you'd validate every coded column against the official code list.
</div>

## Building the VS (Vital Signs) domain

Vital signs are a **findings** domain: one row per measurement, using the
`--TESTCD` / `--ORRES` pattern. Our raw data is already long, which is the
right shape ã we mostly need to rename and add the standard result columns.

```{r}
vs_test_labels <- c(
  SYSBP = "Systolic Blood Pressure",
  DIABP = "Diastolic Blood Pressure",
  PULSE = "Pulse Rate",
  TEMP = "Temperature",
  RESP = "Respiratory Rate"
)

make_vs <- function(raw, dm, studyid) {
  subject_key <- dm |>
    select(subject = SUBJID, USUBJID)
  raw |>
    left_join(subject_key, by = "subject") |>
    mutate(
      STUDYID = studyid,
      DOMAIN = "VS",
      VSTESTCD = measure,
      VSTEST = unname(vs_test_labels[measure]),
      VSORRES = as.character(value),
      VSORRESU = units,
      VSSTRESC = as.character(value),
      VSSTRESN = as.numeric(value),
      VSSTRESU = units,
      VISIT = visit,
      VSDTC = date
    ) |>
    arrange(USUBJID, VSDTC, VSTESTCD) |>
    group_by(USUBJID) |>
    mutate(VSSEQ = row_number()) |>
    ungroup() |>
    select(
      STUDYID, DOMAIN, USUBJID, VSSEQ, VSTESTCD, VSTEST,
      VSORRES, VSORRESU, VSSTRESC, VSSTRESN, VSSTRESU, VISIT, VSDTC
    )
}

vs <- make_vs(raw_vitals, dm, studyid = "DEMO01")
vs
```

The key moves:

- We **joined to `dm`** to attach the proper `USUBJID` ã findings domains
  never re-derive the subject key, they inherit it.
- **`VSSEQ`** is generated per subject with `group_by()` + `row_number()`,
  giving each row a unique sequence number within the subject.
- We split the result into original (`VSORRES`, text) and standardized
  (`VSSTRESN`, numeric) forms. Downstream analysis uses the numeric one.

## Building the AE (Adverse Events) domain

Adverse events are an **events** domain: one row per event, with start/end
dates and a severity. AE also carries a couple of the most-scrutinized
variables in all of clinical data ã seriousness and severity.

```{r}
make_ae <- function(raw, dm, studyid) {
  subject_key <- dm |>
    select(subject = SUBJID, USUBJID)
  raw |>
    left_join(subject_key, by = "subject") |>
    mutate(
      STUDYID = studyid,
      DOMAIN = "AE",
      AETERM = event,
      AESEV = toupper(severity),
      AESTDTC = start,
      AEENDTC = end,
      AEONGO = if_else(is.na(end), "Y", "N")   # ongoing at data cutoff?
    ) |>
    arrange(USUBJID, AESTDTC) |>
    group_by(USUBJID) |>
    mutate(AESEQ = row_number()) |>
    ungroup() |>
    select(
      STUDYID, DOMAIN, USUBJID, AESEQ,
      AETERM, AESEV, AESTDTC, AEENDTC, AEONGO
    )
}

ae <- make_ae(raw_ae, dm, studyid = "DEMO01")
ae
```

Note how a missing end date becomes a derived "ongoing" flag rather than
just an `NA` ã deriving meaningful flags from raw gaps is a huge part of the
real work.

## Why the common structure pays off

Because every domain shares `STUDYID` and `USUBJID`, answering
cross-domain questions is just a join. For example: "for each subject, how
many adverse events did they have, and what arm were they in?"

```{r}
dm_arm <- dm |>
  select(USUBJID, ARM)

dm_arm |>
  left_join(
    ae |> count(USUBJID, name = "n_ae"),
    by = "USUBJID"
  ) |>
  mutate(n_ae = tidyr::replace_na(n_ae, 0)) |>
  arrange(desc(n_ae))
```

Or a quick clinical summary ã mean systolic blood pressure at screening by
treatment arm:

```{r}
vs |>
  filter(VSTESTCD == "SYSBP", VISIT == "Screening") |>
  left_join(dm_arm, by = "USUBJID") |>
  group_by(ARM) |>
  summarise(
    n = n(),
    mean_sysbp = mean(VSSTRESN),
    .groups = "drop"
  )
```

This is the whole point of SDTM: once the data is in this shape, *any*
analysis is a predictable combination of filters, joins, and summaries.

## A tiny validation function

Regulators run automated conformance checks (Pinnacle 21 is the famous one)
against submitted SDTM. You can write lightweight versions of those checks
yourself. Here's one that flags the most common structural problems:

```{r}
check_domain <- function(df, domain, key_vars = c("STUDYID", "USUBJID")) {
  issues <- character(0)

  missing_keys <- setdiff(key_vars, names(df))
  if (length(missing_keys) > 0) {
    issues <- c(issues, paste("Missing key variable(s):",
                              paste(missing_keys, collapse = ", ")))
  }

  if ("DOMAIN" %in% names(df) && any(df$DOMAIN != domain)) {
    issues <- c(issues, "DOMAIN value does not match expected domain")
  }

  seq_var <- paste0(domain, "SEQ")
  if (seq_var %in% names(df)) {
    dup <- df |>
      count(USUBJID, .data[[seq_var]]) |>
      filter(n > 1)
    if (nrow(dup) > 0) {
      issues <- c(issues, paste0(seq_var, " is not unique within USUBJID"))
    }
  }

  if (length(issues) == 0) {
    cat("PASS:", domain, "looks structurally valid\n")
  } else {
    cat("ISSUES in", domain, ":\n")
    cat(paste0("  - ", issues, collapse = "\n"), "\n")
  }
  invisible(issues)
}

check_domain(dm, "DM")
check_domain(vs, "VS")
check_domain(ae, "AE")
```

Real conformance rules number in the hundreds, but the idea is the same:
programmatic, repeatable checks that either pass or point you at the exact
problem.

<div class="callout callout-tip">
<div class="callout-title">ÞÀ The pharmaverse</div>
Everything here is hand-rolled to show the mechanics, but you rarely start
from zero in practice. The open-source **pharmaverse** ã packages like
<a href="https://pharmaverse.github.io/admiral/">admiral</a> (for building
the analysis-ready ADaM datasets that come <em>after</em> SDTM),
<a href="https://pharmaverse.github.io/sdtm.oak/">sdtm.oak</a> (for building
SDTM itself), and <code>pharmaverseadam</code> / <code>pharmaversesdtm</code>
(realistic example data) ã gives you tested, regulator-aware tooling. This
page is the "under the hood" version so those packages make sense when you
reach for them.
</div>

## SDTM vs. ADaM (the next step)

One more thing worth knowing: SDTM is not the end of the pipeline. It's the
*tabulation* layer ã a faithful, standardized record of what was collected.
For the actual statistical analysis, that data gets transformed again into
**ADaM** (Analysis Data Model) datasets, which are built to be
one-row-per-analysis-record and carry all the derived flags and baselines a
statistician needs.

The typical flow:

```
Raw EDC data  ->  SDTM (standardized tabulation)  ->  ADaM (analysis-ready)  ->  Tables, Figures, Listings
```

Each arrow is a set of documented, validated R (or SAS) programs. If you can
build the SDTM step, the ADaM step is the same skills applied one layer up.

## Recap

- **SDTM** is a CDISC standard that puts clinical trial data into a fixed set
  of domains (`DM`, `AE`, `VS`, `LB`, ãÎ) with standardized variable names.
- Every domain shares `STUDYID` and `USUBJID`, so cross-domain analysis is
  just a join.
- Building a domain in R is: **rename to SDTM variables, derive the keys and
  flags, add sequence numbers, and validate**.
- Findings domains (`VS`, `LB`) use the `--TESTCD` / `--ORRES` / `--STRESN`
  pattern; events domains (`AE`) use start/end dates; `DM` is one row per
  subject.
- In production you'd lean on the **pharmaverse**, but the transformations
  are ordinary tidyverse under the hood.

If you found this useful or spotted something to fix, email me at
**cavandonohoe@gmail.com** or open an issue on
[GitHub](https://github.com/cavandonohoe/cavandonohoe.github.io).
