---
title: "SDTM with R: Clinical Trial Data Standards"
author: "Cavan Donohoe"
description: "What SDTM is, why clinical trials use it, how to build tidy SDTM domains (DM, AE, VS, LB) from raw data with reusable R functions, and the hard edge cases that trip everyone up."
output:
  html_document:
    toc: true
    toc_float: true
    includes:
      in_header: header/header.html
      after_body: include_footer.html
lang: "en"
---

<link rel="stylesheet" href="/css/learn-r.css">

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.width = 6.5,
  fig.height = 4.2,
  dpi = 100
)
options(scipen = 999, digits = 4)
```

# SDTM with R

Every drug and diagnostic that gets submitted to the FDA (and most other
regulators) arrives with its clinical trial data in a very specific shape.
That shape is **SDTM** — the Study Data Tabulation Model, a CDISC standard.
If you work in clinical data science, biostatistics, or statistical
programming, SDTM is the lingua franca. This page is a from-scratch
introduction: what it is, why it exists, and how to build SDTM-style
datasets in R using a handful of small, reusable functions.

You don't need any clinical background. If you've done a little tidyverse
(see the [Learn R course](learn_r.html)), you can follow everything here.

## What is SDTM?

Imagine every pharma company invented its own column names, table layouts,
and coding conventions for their trial data. A reviewer at the FDA would
have to relearn the data from scratch for every single submission. SDTM
exists to stop that.

**SDTM standardizes the *content and structure* of the raw, collected trial
data** so that any reviewer, any validation tool, and any downstream program
knows exactly where to look. It defines:

- A fixed set of **domains** — one per type of data collected. Each has a
  two-letter code.
- Standard **variable names** within each domain (again, mostly two-to-eight
  letter codes).
- Rules for how rows, subjects, visits, and timing are represented.

<div class="callout callout-tip">
<div class="callout-title">💡 The mental model</div>
SDTM is "one tidy table per kind of thing that happened in the trial."
Demographics is one table. Adverse events is another. Vital signs is
another. Every table shares a common backbone (subject ID, study ID) so they
can be joined back together.
</div>

### The domains you'll see most often

| Code | Domain | One row per... |
|------|--------|----------------|
| `DM` | Demographics | subject |
| `AE` | Adverse Events | adverse event a subject had |
| `VS` | Vital Signs | vital-sign measurement |
| `LB` | Laboratory | lab test result |
| `CM` | Concomitant Meds | medication a subject took |
| `EX` | Exposure | dose of study drug administered |

There are dozens more, but if you understand these you understand the model.

### The naming conventions

SDTM variable names follow a pattern. Once you see it, the columns stop
looking like alphabet soup:

- **`--SEQ`** — a sequence number, unique within a subject and domain
  (`AESEQ`, `VSSEQ`).
- **`--TESTCD` / `--TEST`** — the short code and long name of what was
  measured (`VSTESTCD = "SYSBP"`, `VSTEST = "Systolic Blood Pressure"`).
- **`--ORRES` / `--ORRESU`** — the original result as collected, and its unit
  (`VSORRES = "120"`, `VSORRESU = "mmHg"`).
- **`--STRESC` / `--STRESN`** — the result standardized to a character and a
  numeric form.
- **`--DTC`** — a date/time in ISO 8601 text (`AESTDTC = "2023-04-12"`).
- **`STUDYID` / `USUBJID`** — the study ID and the *unique* subject ID, the
  keys that tie every domain together.

The leading `--` is a placeholder for the domain code. So the "test code"
variable is `VSTESTCD` in Vital Signs and `LBTESTCD` in Labs.

## Some raw data to start from

In real life the raw data comes out of an electronic data capture (EDC)
system and looks nothing like SDTM. Our job is to reshape it. Let's fabricate
a tiny, messy "raw" export so we have something concrete to transform.

```{r}
library(dplyr)
library(tidyr)

set.seed(42)

raw_demographics <- tibble::tribble(
  ~subject, ~site, ~birth_date,  ~sex,     ~race,    ~arm,
  "001",    "01",  "1968-03-11", "Male",   "White",  "Active",
  "002",    "01",  "1975-11-30", "Female", "Asian",  "Placebo",
  "003",    "02",  "1954-07-02", "Female", "White",  "Active",
  "004",    "02",  "1981-01-19", "Male",   "Black",  "Placebo",
  "005",    "03",  "1990-09-25", "Male",   "White",  "Active"
)

raw_demographics
```

A raw vital-signs export, in "long" form, one measurement per row:

```{r}
raw_vitals <- tibble::tribble(
  ~subject, ~visit,      ~measure, ~value, ~units, ~date,
  "001",    "Screening", "SYSBP",  128,    "mmHg", "2023-01-05",
  "001",    "Screening", "DIABP",  82,     "mmHg", "2023-01-05",
  "001",    "Week 4",    "SYSBP",  122,    "mmHg", "2023-02-02",
  "001",    "Week 4",    "DIABP",  79,     "mmHg", "2023-02-02",
  "002",    "Screening", "SYSBP",  141,    "mmHg", "2023-01-06",
  "002",    "Screening", "DIABP",  90,     "mmHg", "2023-01-06",
  "003",    "Screening", "SYSBP",  118,    "mmHg", "2023-01-09",
  "003",    "Week 4",    "SYSBP",  115,    "mmHg", "2023-02-07"
)

raw_vitals
```

And a raw adverse-event log:

```{r}
raw_ae <- tibble::tribble(
  ~subject, ~event,        ~severity,  ~start,       ~end,
  "001",    "Headache",    "Mild",     "2023-01-20", "2023-01-22",
  "001",    "Nausea",      "Moderate", "2023-02-05", NA,
  "003",    "Fatigue",     "Mild",     "2023-01-15", "2023-01-30",
  "004",    "Dizziness",   "Severe",   "2023-01-25", "2023-01-26"
)

raw_ae
```

## Building the DM (Demographics) domain

DM is the anchor domain: exactly one row per subject. Everything else joins
back to it. Building it means renaming to SDTM variables, constructing the
`USUBJID` key, and deriving `AGE` from the birth date.

Here's a reusable function:

```{r}
make_dm <- function(raw, studyid) {
  raw |>
    mutate(
      STUDYID = studyid,
      DOMAIN = "DM",
      SITEID = site,
      SUBJID = subject,
      USUBJID = paste(studyid, site, subject, sep = "-"),
      RFSTDTC = birth_date,                       # reference start (placeholder)
      BRTHDTC = birth_date,
      AGE = as.integer(
        floor(as.numeric(Sys.Date() - as.Date(birth_date)) / 365.25)
      ),
      AGEU = "YEARS",
      SEX = dplyr::recode(sex, "Male" = "M", "Female" = "F"),
      RACE = toupper(race),
      ARM = arm,
      ACTARM = arm
    ) |>
    select(
      STUDYID, DOMAIN, USUBJID, SUBJID, SITEID,
      BRTHDTC, AGE, AGEU, SEX, RACE, ARM, ACTARM
    )
}

dm <- make_dm(raw_demographics, studyid = "DEMO01")
dm
```

A few SDTM conventions on display here:

- **`USUBJID` must be unique across the whole study**, so we build it from
  study + site + subject. Two sites could both have a subject "001"; the
  `USUBJID` keeps them distinct.
- **`SEX` is coded** to controlled terminology (`M`/`F`), not free text.
- **`AGE` is derived**, not collected. SDTM is full of derived variables like
  this.

<div class="callout callout-warning">
<div class="callout-title">⚠️ Controlled terminology</div>
Real SDTM uses CDISC **controlled terminology** — fixed, published code lists
for values like `SEX`, `RACE`, and units. You can't put "Male" in `SEX`; it
has to be `M`. Here we're doing a simplified `recode()`, but in a real study
you'd validate every coded column against the official code list.
</div>

## Building the VS (Vital Signs) domain

Vital signs are a **findings** domain: one row per measurement, using the
`--TESTCD` / `--ORRES` pattern. Our raw data is already long, which is the
right shape — we mostly need to rename and add the standard result columns.

```{r}
vs_test_labels <- c(
  SYSBP = "Systolic Blood Pressure",
  DIABP = "Diastolic Blood Pressure",
  PULSE = "Pulse Rate",
  TEMP = "Temperature",
  RESP = "Respiratory Rate"
)

make_vs <- function(raw, dm, studyid) {
  subject_key <- dm |>
    select(subject = SUBJID, USUBJID)
  raw |>
    left_join(subject_key, by = "subject") |>
    mutate(
      STUDYID = studyid,
      DOMAIN = "VS",
      VSTESTCD = measure,
      VSTEST = unname(vs_test_labels[measure]),
      VSORRES = as.character(value),
      VSORRESU = units,
      VSSTRESC = as.character(value),
      VSSTRESN = as.numeric(value),
      VSSTRESU = units,
      VISIT = visit,
      VSDTC = date
    ) |>
    arrange(USUBJID, VSDTC, VSTESTCD) |>
    group_by(USUBJID) |>
    mutate(VSSEQ = row_number()) |>
    ungroup() |>
    select(
      STUDYID, DOMAIN, USUBJID, VSSEQ, VSTESTCD, VSTEST,
      VSORRES, VSORRESU, VSSTRESC, VSSTRESN, VSSTRESU, VISIT, VSDTC
    )
}

vs <- make_vs(raw_vitals, dm, studyid = "DEMO01")
vs
```

The key moves:

- We **joined to `dm`** to attach the proper `USUBJID` — findings domains
  never re-derive the subject key, they inherit it.
- **`VSSEQ`** is generated per subject with `group_by()` + `row_number()`,
  giving each row a unique sequence number within the subject.
- We split the result into original (`VSORRES`, text) and standardized
  (`VSSTRESN`, numeric) forms. Downstream analysis uses the numeric one.

## Building the AE (Adverse Events) domain

Adverse events are an **events** domain: one row per event, with start/end
dates and a severity. AE also carries a couple of the most-scrutinized
variables in all of clinical data — seriousness and severity.

```{r}
make_ae <- function(raw, dm, studyid) {
  subject_key <- dm |>
    select(subject = SUBJID, USUBJID)
  raw |>
    left_join(subject_key, by = "subject") |>
    mutate(
      STUDYID = studyid,
      DOMAIN = "AE",
      AETERM = event,
      AESEV = toupper(severity),
      AESTDTC = start,
      AEENDTC = end,
      AEONGO = if_else(is.na(end), "Y", "N")   # ongoing at data cutoff?
    ) |>
    arrange(USUBJID, AESTDTC) |>
    group_by(USUBJID) |>
    mutate(AESEQ = row_number()) |>
    ungroup() |>
    select(
      STUDYID, DOMAIN, USUBJID, AESEQ,
      AETERM, AESEV, AESTDTC, AEENDTC, AEONGO
    )
}

ae <- make_ae(raw_ae, dm, studyid = "DEMO01")
ae
```

Note how a missing end date becomes a derived "ongoing" flag rather than
just an `NA` — deriving meaningful flags from raw gaps is a huge part of the
real work.

## Why the common structure pays off

Because every domain shares `STUDYID` and `USUBJID`, answering
cross-domain questions is just a join. For example: "for each subject, how
many adverse events did they have, and what arm were they in?"

```{r}
dm_arm <- dm |>
  select(USUBJID, ARM)

dm_arm |>
  left_join(
    ae |> count(USUBJID, name = "n_ae"),
    by = "USUBJID"
  ) |>
  mutate(n_ae = tidyr::replace_na(n_ae, 0)) |>
  arrange(desc(n_ae))
```

Or a quick clinical summary — mean systolic blood pressure at screening by
treatment arm:

```{r}
vs |>
  filter(VSTESTCD == "SYSBP", VISIT == "Screening") |>
  left_join(dm_arm, by = "USUBJID") |>
  group_by(ARM) |>
  summarise(
    n = n(),
    mean_sysbp = mean(VSSTRESN),
    .groups = "drop"
  )
```

This is the whole point of SDTM: once the data is in this shape, *any*
analysis is a predictable combination of filters, joins, and summaries.

## A tiny validation function

Regulators run automated conformance checks (Pinnacle 21 is the famous one)
against submitted SDTM. You can write lightweight versions of those checks
yourself. Here's one that flags the most common structural problems:

```{r}
check_domain <- function(df, domain, key_vars = c("STUDYID", "USUBJID")) {
  issues <- character(0)

  missing_keys <- setdiff(key_vars, names(df))
  if (length(missing_keys) > 0) {
    issues <- c(issues, paste("Missing key variable(s):",
                              paste(missing_keys, collapse = ", ")))
  }

  if ("DOMAIN" %in% names(df) && any(df$DOMAIN != domain)) {
    issues <- c(issues, "DOMAIN value does not match expected domain")
  }

  seq_var <- paste0(domain, "SEQ")
  if (seq_var %in% names(df)) {
    dup <- df |>
      count(USUBJID, .data[[seq_var]]) |>
      filter(n > 1)
    if (nrow(dup) > 0) {
      issues <- c(issues, paste0(seq_var, " is not unique within USUBJID"))
    }
  }

  if (length(issues) == 0) {
    cat("PASS:", domain, "looks structurally valid\n")
  } else {
    cat("ISSUES in", domain, ":\n")
    cat(paste0("  - ", issues, collapse = "\n"), "\n")
  }
  invisible(issues)
}

check_domain(dm, "DM")
check_domain(vs, "VS")
check_domain(ae, "AE")
```

Real conformance rules number in the hundreds, but the idea is the same:
programmatic, repeatable checks that either pass or point you at the exact
problem.

<div class="callout callout-tip">
<div class="callout-title">💡 The pharmaverse</div>
Everything here is hand-rolled to show the mechanics, but you rarely start
from zero in practice. The open-source **pharmaverse** — packages like
<a href="https://pharmaverse.github.io/admiral/">admiral</a> (for building
the analysis-ready ADaM datasets that come <em>after</em> SDTM),
<a href="https://pharmaverse.github.io/sdtm.oak/">sdtm.oak</a> (for building
SDTM itself), and <code>pharmaverseadam</code> / <code>pharmaversesdtm</code>
(realistic example data) — gives you tested, regulator-aware tooling. This
page is the "under the hood" version so those packages make sense when you
reach for them.
</div>

# The hard part: edge cases and persistent issues

The functions above work on clean, well-behaved data. Real trial data is
never clean. This is the part nobody warns you about — the edge cases that
quietly corrupt an analysis and the issues that resurface on every single
study. Each one below comes with a demonstration of *how it bites* and a
defensible fix.

## 1. Partial and impossible dates

This is the single most persistent issue in all of SDTM. Sites collect dates
that are incomplete ("the patient started the med sometime in March 2023,
they don't remember the day"), and the ISO 8601 `--DTC` format is built to
represent exactly that: `2023-03` (no day), `2023` (year only), or even
`--03-15` (no year). Naively parsing these with `as.Date()` silently turns
them into `NA`, and every downstream calculation that touches them breaks.

```{r}
messy_dates <- c(
  "2023-04-12",   # complete
  "2023-03",      # missing day
  "2023",         # year only
  "",             # entirely missing
  "2023-02-30"    # IMPOSSIBLE: February has no 30th
)

as.Date(messy_dates)   # naive parse: silent NAs, no signal about *why*
```

Notice the trap: the missing-day date, the year-only date, *and* the
impossible Feb 30th all collapse to `NA` indistinguishably. You can't tell a
legitimately-partial date from genuinely-bad data. The fix is to classify
the completeness first, then impute deterministically (the convention is to
impute to the *first* of the period for start dates and the *last* for end
dates), and flag what you did:

```{r}
classify_dtc <- function(dtc) {
  dplyr::case_when(
    is.na(dtc) | dtc == ""            ~ "missing",
    grepl("^\\d{4}-\\d{2}-\\d{2}$", dtc) ~ "complete",
    grepl("^\\d{4}-\\d{2}$", dtc)     ~ "no_day",
    grepl("^\\d{4}$", dtc)            ~ "year_only",
    TRUE                              ~ "other"
  )
}

impute_start_dtc <- function(dtc) {
  precision <- classify_dtc(dtc)
  imputed <- dplyr::case_when(
    precision == "complete"  ~ dtc,
    precision == "no_day"    ~ paste0(dtc, "-01"),
    precision == "year_only" ~ paste0(dtc, "-01-01"),
    TRUE                     ~ NA_character_
  )
  tibble::tibble(
    dtc_raw = dtc,
    precision = precision,
    dtc_imputed = imputed,
    # as.Date on a well-formed string still catches Feb 30th as NA:
    date = as.Date(imputed),
    imputed_flag = precision %in% c("no_day", "year_only")
  )
}

impute_start_dtc(messy_dates)
```

Now the year-only and missing-day dates are imputed *and flagged*
(`imputed_flag = TRUE`), the truly-missing one is honestly `NA`, and Feb 30th
is caught as bad data (`date` is `NA` even though `precision` said
"complete") — a signal you can act on rather than a silent hole.

<div class="callout callout-warning">
<div class="callout-title">⚠️ Imputation is a decision, not a default</div>
Every imputed date must carry a flag (in ADaM these become `--DTF` /
`--TMF`, the date/time imputation flags). If a statistician later computes
"days on treatment" from an imputed start date, they need to know it was
imputed. Silently guessing is how you end up with a finding in an FDA
audit.
</div>

## 2. Duplicate records that aren't obviously duplicates

Sites re-enter data. A subject's screening visit gets keyed twice, or an AE
gets logged by two coordinators. The `--SEQ` variable is supposed to be
unique within subject — but if you generate it with `row_number()` *after*
the duplicates are already in, you'll happily assign `1` and `2` to two rows
that are actually the same event, and the duplicate becomes invisible.

```{r}
raw_vitals_dupes <- dplyr::bind_rows(
  raw_vitals,
  raw_vitals |> dplyr::slice(1)   # subject 001's screening SYSBP, re-entered
)

# Wrong: seq assigned over the dupes, so they look like distinct records
raw_vitals_dupes |>
  dplyr::filter(subject == "001", measure == "SYSBP", visit == "Screening")
```

The fix is to define what makes a record *unique* (the "natural key" — here
subject + visit + measure + date) and de-duplicate on it *before* assigning
`--SEQ`:

```{r}
natural_key_vs <- c("subject", "visit", "measure", "date")

dedup_report <- raw_vitals_dupes |>
  dplyr::add_count(dplyr::across(dplyr::all_of(natural_key_vs)),
                   name = "n_with_key") |>
  dplyr::filter(n_with_key > 1) |>
  dplyr::distinct(dplyr::across(dplyr::all_of(natural_key_vs)))

dedup_report   # exactly which records collide

raw_vitals_clean <- raw_vitals_dupes |>
  dplyr::distinct(dplyr::across(dplyr::all_of(natural_key_vs)),
                  .keep_all = TRUE)

nrow(raw_vitals_dupes)
nrow(raw_vitals_clean)   # one fewer: the re-entered row is gone
```

Report the collisions (never delete silently — a data-management query goes
back to the site), then de-duplicate on the natural key, and *only then*
assign `--SEQ`.

## 3. Unit chaos in findings domains

Findings domains are where labs live, and labs are a unit nightmare. The same
test arrives from different sites in different units: glucose in `mg/dL` from
one lab and `mmol/L` from another. If you compare `--ORRES` numerically
without standardizing, you're comparing apples to a number that's ~18x
different. This is what `--STRESN` / `--STRESU` (the *standardized* result)
exists to solve, and getting the conversion factors right is a perpetual
source of bugs.

```{r}
raw_labs <- tibble::tribble(
  ~subject, ~test,      ~result, ~unit,
  "001",    "GLUC",     99,      "mg/dL",
  "002",    "GLUC",     5.5,     "mmol/L",   # same magnitude, different unit
  "003",    "GLUC",     110,     "mg/dL"
)

# Naive: treat the numbers as comparable -> subject 002 looks impossibly low
raw_labs |>
  dplyr::group_by(test) |>
  dplyr::summarise(mean_result = mean(result), .groups = "drop")
```

A mean of a `mg/dL` value and a `mmol/L` value is meaningless. Standardize to
a single unit per test with an explicit, auditable conversion table:

```{r}
# Conversion factors TO the standard unit for each test
lab_conversions <- tibble::tribble(
  ~test,  ~from_unit, ~std_unit, ~factor,
  "GLUC", "mg/dL",    "mg/dL",   1,
  "GLUC", "mmol/L",   "mg/dL",   18.0182   # 1 mmol/L glucose = 18.0182 mg/dL
)

lb <- raw_labs |>
  dplyr::left_join(lab_conversions, by = c("test", "unit" = "from_unit")) |>
  dplyr::mutate(
    LBORRES = as.character(result),
    LBORRESU = unit,
    LBSTRESN = result * factor,
    LBSTRESU = std_unit
  )

lb |>
  dplyr::select(subject, test, LBORRES, LBORRESU, LBSTRESN, LBSTRESU)

# Now the summary is meaningful because everything is in mg/dL
lb |>
  dplyr::group_by(test) |>
  dplyr::summarise(mean_std = mean(LBSTRESN), .groups = "drop")
```

<div class="callout callout-warning">
<div class="callout-title">⚠️ A missing conversion factor is worse than a wrong one</div>
If a unit shows up that isn't in your conversion table, the `left_join`
gives `factor = NA` and `LBSTRESN` silently becomes `NA`. Always assert that
every `--ORRES` produced a non-missing `--STRESN` (or a documented reason
why not), or an entire lab's worth of results can vanish without a warning.
</div>

## 4. Character results in a numeric world

The other half of the labs problem: not every result is a number.
Results come back as `"<0.5"` (below the limit of detection), `"POSITIVE"`,
`">1000"`, or `"NEGATIVE"`. `--ORRES` is character precisely so it can hold
these, but `--STRESN` is numeric — and `as.numeric("<0.5")` is `NA` with a
warning. You have to decide, per test, how each non-numeric value maps.

```{r}
raw_results <- c("42", "<0.5", ">1000", "POSITIVE", "12.7", "NEGATIVE")

suppressWarnings(as.numeric(raw_results))   # everything non-numeric -> NA
```

A defensible approach keeps the character result verbatim in `--STRESC`,
derives `--STRESN` only where a number is genuinely present, and records
*why* a value is non-numeric so it isn't mistaken for missing:

```{r}
standardize_result <- function(orres) {
  is_below <- grepl("^<", orres)
  is_above <- grepl("^>", orres)
  numeric_part <- suppressWarnings(as.numeric(gsub("[<>]", "", orres)))
  tibble::tibble(
    STRESC = orres,                       # always keep the original
    STRESN = numeric_part,                # NA for POSITIVE / NEGATIVE
    result_type = dplyr::case_when(
      is_below                 ~ "below_limit",
      is_above                 ~ "above_limit",
      !is.na(numeric_part)     ~ "numeric",
      TRUE                     ~ "qualitative"
    )
  )
}

standardize_result(raw_results)
```

Now `"<0.5"` is flagged `below_limit` (a stats analysis might substitute half
the detection limit), and `"POSITIVE"` is `qualitative` rather than a
mysterious `NA`. The information about *why* the number is absent survives.

## 5. Subjects who exist in one domain but not another

`DM` is supposed to be the master list of subjects, and every other domain
should only contain subjects who are in `DM`. In practice, an AE gets logged
against a subject ID that was mistyped, or a subject is in `DM` but has an AE
row that references a slightly different ID. These **orphan records** are a
top conformance finding and they silently drop out of any inner join.

```{r}
ae_with_orphan <- dplyr::bind_rows(
  ae,
  ae |> dplyr::slice(1) |> dplyr::mutate(USUBJID = "DEMO01-99-999")  # typo'd ID
)

# Orphans: AE subjects with no matching DM record
orphans <- dplyr::anti_join(ae_with_orphan, dm, by = "USUBJID")
orphans |> dplyr::select(USUBJID, AETERM)

# And the reverse: DM subjects with zero AEs (often fine, but worth knowing)
dm_no_ae <- dplyr::anti_join(dm, ae_with_orphan, by = "USUBJID")
dm_no_ae |> dplyr::select(USUBJID)
```

The `anti_join` in both directions is the workhorse referential-integrity
check. An orphan AE is never dropped silently — it's surfaced as a query.
The reverse (subjects with no AEs) is usually legitimate but is exactly the
kind of thing you confirm rather than assume.

## 6. The `left_join` vs `inner_join` trap

Following directly from the orphan problem: the choice between `left_join`
and `inner_join` when combining domains is one of the most consequential and
most-overlooked decisions in the whole pipeline. An `inner_join` silently
drops subjects, which can bias a result — and because nothing errors, it can
survive all the way to a submitted table.

```{r}
# inner_join drops the orphan AE and any DM subject without an AE
dm |>
  dplyr::inner_join(ae_with_orphan, by = "USUBJID") |>
  dplyr::distinct(USUBJID) |>
  nrow()

# left_join from DM keeps every enrolled subject (correct denominator)
dm |>
  dplyr::left_join(ae_with_orphan, by = "USUBJID") |>
  dplyr::distinct(USUBJID) |>
  nrow()
```

The safety habit: **decide the denominator explicitly.** For a "% of subjects
with an AE" table, the denominator is enrolled subjects (`DM`), so you
`left_join` *from* `DM` and count non-matches as "no event," never as
"absent." An `inner_join` here would quietly shrink your denominator and
inflate the rate.

## 7. Controlled terminology drift

Every coded variable (`SEX`, `RACE`, severity, units, ...) must match a
published CDISC code list *exactly* — including case and spelling. Sites and
labs supply free-ish text that drifts: `"Male"`, `"MALE"`, `"m"`, `"Man"`.
Recoding only the values you happened to see is a recipe for a value slipping
through unmapped on the next data cut.

```{r}
incoming_sex <- c("Male", "MALE", "m", "Female", "F", "Man", "")

# A closed allow-list, not an open recode: anything unmapped is caught
sex_map <- c(
  "male" = "M", "m" = "M", "man" = "M",
  "female" = "F", "f" = "F", "woman" = "F"
)

map_controlled <- function(x, map) {
  key <- tolower(trimws(x))
  mapped <- unname(map[key])
  tibble::tibble(
    raw = x,
    mapped = mapped,
    unmapped = is.na(mapped) & !(key %in% c("", NA))
  )
}

map_controlled(incoming_sex, sex_map)
```

`"Man"` maps fine (it's in the allow-list), the empty string is honestly
unmapped-but-not-flagged (genuinely missing), and if `"Intersex"` or a typo
like `"Femal"` showed up it would surface as `unmapped = TRUE` instead of
being silently dropped or, worse, passed through as invalid terminology. The
principle: **validate against a closed code list and flag the misses**,
rather than an open-ended `recode()` that only handles today's values.

## The through-line

Every one of these has the same shape: the naive operation *succeeds
silently* and produces plausible-looking garbage — an `NA` that means three
different things, a mean across incompatible units, a denominator that
quietly shrank. The discipline that makes SDTM trustworthy isn't clever code;
it's **refusing to let anything happen silently**: classify before you
convert, flag every imputation, assert referential integrity, and validate
coded values against closed lists. That mindset is most of what separates a
production clinical programmer from someone who merely knows the tidyverse.


## SDTM vs. ADaM (the next step)

One more thing worth knowing: SDTM is not the end of the pipeline. It's the
*tabulation* layer — a faithful, standardized record of what was collected.
For the actual statistical analysis, that data gets transformed again into
**ADaM** (Analysis Data Model) datasets, which are built to be
one-row-per-analysis-record and carry all the derived flags and baselines a
statistician needs.

The typical flow:

```
Raw EDC data  ->  SDTM (standardized tabulation)  ->  ADaM (analysis-ready)  ->  Tables, Figures, Listings
```

Each arrow is a set of documented, validated R (or SAS) programs. If you can
build the SDTM step, the ADaM step is the same skills applied one layer up.

## Recap

- **SDTM** is a CDISC standard that puts clinical trial data into a fixed set
  of domains (`DM`, `AE`, `VS`, `LB`, …) with standardized variable names.
- Every domain shares `STUDYID` and `USUBJID`, so cross-domain analysis is
  just a join.
- Building a domain in R is: **rename to SDTM variables, derive the keys and
  flags, add sequence numbers, and validate**.
- Findings domains (`VS`, `LB`) use the `--TESTCD` / `--ORRES` / `--STRESN`
  pattern; events domains (`AE`) use start/end dates; `DM` is one row per
  subject.
- In production you'd lean on the **pharmaverse**, but the transformations
  are ordinary tidyverse under the hood.
- The hard part isn't the happy path — it's **partial dates, duplicates,
  unit conversions, non-numeric results, orphan records, join semantics, and
  terminology drift**. The unifying discipline is refusing to let bad data
  pass silently: classify before converting, flag every imputation, and
  validate against closed code lists.

If you found this useful or spotted something to fix, email me at
**cavandonohoe@gmail.com** or open an issue on
[GitHub](https://github.com/cavandonohoe/cavandonohoe.github.io).
