---
title: "S&P 500 Perspective"
editor_options:
  chunk_output_type: console
output:
  html_document:
    includes:
      in_header: header/header.html
      after_body: include_footer.html
---

```{r og-meta, echo=FALSE, results='asis'}
cat('
<meta property="og:title" content="S&P 500 Perspective">
<meta property="og:description" content="Historical S&P 500 performance with record-breaking closes and intersection analysis.">
<meta property="og:url" content="https://cavandonohoe.github.io/sp500.html">
<meta property="og:type" content="article">
')
```

```{r pipe, include=FALSE}
`%>%` <- magrittr::`%>%`
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  results = "asis"
)
```

```{r data}
price_data <- readr::read_csv(here::here("data", "voo_stock_prices.csv"),
  show_col_types = FALSE)
```

```{r timestamp, results='asis'}
cat(sprintf(
  '<span class="data-timestamp">Data last updated: %s</span>\n',
  format(as.Date(system(paste("git log -1 --format=%ai --",
    here::here("data", "voo_stock_prices.csv")), intern = TRUE)), "%B %d, %Y")
))
```

```{r compute}
max_change <- price_data %>%
  dplyr::mutate(day_change = dplyr::lead(close) / close) %>%
  dplyr::filter(day_change == max(day_change, na.rm = TRUE))
max_date <- max_change %>%
  dplyr::slice(1) %>%
  dplyr::pull(date)
max_change_perc <- max_change %>%
  dplyr::slice(1) %>%
  dplyr::pull(day_change)
min_date <- price_data %>%
  dplyr::filter(date == min(date, na.rm = TRUE)) %>%
  dplyr::pull(date)

voo_full <- price_data %>%
  dplyr::select(symbol, date, close_official = close) %>%
  dplyr::filter(!is.na(close_official))

voo_full$rolling_max <- NA
for (i in seq_len(nrow(voo_full))) {
  voo_full$rolling_max[i] <- max(voo_full$close_official[1:i])
}

voo_maxes <- voo_full %>%
  dplyr::filter(close_official == rolling_max)

intersections_voo <- voo_full %>%
  dplyr::mutate(
    intersection = close_official >= close_official[date == max(date)],
    intersection_lead = dplyr::lead(intersection)
  ) %>%
  dplyr::filter(intersection != intersection_lead & !intersection)
```

## S&P 500 Historical Performance

Omg did the stock market just crash? 1% down in a day is a crash right? Well maybe not... Let's just take a look at the history of the S&P 500 for some perspective.

```{r sp500_full}
sp500_plot <- voo_full %>%
  ggplot2::ggplot(ggplot2::aes(x = date, y = close_official)) +
  ggplot2::geom_line() +
  ggplot2::geom_point(data = voo_maxes, ggplot2::aes(x = date, y = rolling_max),
    color = "red", size = .5) +
  ggplot2::ggtitle("S&P 500") +
  ggplot2::ylab("Index Level")

sp500_plot %>%
  plotly::ggplotly()
```

Btw, every red dot is a record breaking closing day. We are currently at `r voo_maxes %>% nrow()` record breaking closes since `r voo_full %>% dplyr::pull(date) %>% min()`.

## S&P 500 More Recently

Let's zoom in on just the last three months.

```{r sp500_intersections}
sp_intersections_plot <- voo_full %>%
  dplyr::filter(date >= min(intersections_voo$date) - 60) %>%
  ggplot2::ggplot(ggplot2::aes(x = date, y = close_official)) +
  ggplot2::geom_line() +
  ggplot2::ggtitle("S&P 500") +
  ggplot2::ylab("Index Level") +
  ggrepel::geom_text_repel(
    data = intersections_voo,
    ggplot2::aes(label = as.character(date), segment.color = "red"),
    box.padding = .8, nudge_y = 15, direction = "x", vjust = 0, hjust = 1,
    xlim = c(
      min(voo_full$date[voo_full$date >= min(intersections_voo$date) - 60]),
      max(voo_full$date)
    )
  ) +
  ggplot2::geom_hline(
    yintercept = voo_full$close_official[voo_full$date == max(voo_full$date)],
    linetype = 2, colour = "red"
  ) +
  ggplot2::labs(caption = paste0(
    "Note, the date captions note the price the day before\n",
    "the price closes just under today's price/close"
  ))

sp_intersections_plot
```

Guess what? If you invested in the S&P 500 on any of the labeled dates or prior to `r min(intersections_voo$date)`, then you're still net positive. So calm down.

Also, the largest single day increase from close to close since `r min_date` was on `r max_date` with a single day jump of `r scales::percent(max_change_perc - 1, accuracy = .1)`. So, keep investing and continue to buy the dip.
