---
title: "S&P 500 Perspective"
editor_options: 
  chunk_output_type: console
output:
  html_document:
    includes:
      in_header: header/header.html
---

```{r setup, include=FALSE}
# Set default chunk output
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  results = "asis"
)
```

```{r libraries}
library(tidyverse)
library(DT)
library(scales)
library(plotly)
library(tidyquant)
library(rvest)
```

```{r grab_stock_data}

price_data = tq_get(x = "VOO", from = "2010-01-01", to = as.character(Sys.Date()), get = "stock.prices")

# buy the dip
max_change = price_data %>% mutate(day_change = lead(close) / close) %>%
  filter(day_change == max(day_change, na.rm = TRUE))
max_date = max_change %>% slice(1) %>% pull(date)
max_change_perc = max_change %>% slice(1) %>% pull(day_change)
# percent(max_change_perc - 1, accuracy = .1)
min_date = price_data %>%
  filter(date == min(date, na.rm = TRUE)) %>% pull(date)


url = "https://finance.yahoo.com/quote/VOO/history?p=VOO"

voo_history_html = read_html(url)

voo_history = voo_history_html %>% 
  html_nodes("#Col1-1-HistoricalDataTable-Proxy > section") %>% 
  html_table()

voo_history_close = voo_history[[1]] %>%
  filter(!grepl(x = Open, pattern = "Dividend|Close")) %>%
  mutate(date = as.Date(Date, format = "%b %d, %Y"),
         `Close*` = as.numeric(`Close*`)) %>%
  select(date, close_yahoo = `Close*`) %>%
  mutate(symbol = "VOO")

voo_full = price_data %>% 
  select(symbol, date, close) %>% 
  full_join(voo_history_close) %>% 
  mutate(close_official = ifelse(is.na(close), close_yahoo, close)) %>% 
  select(symbol, date, close_official)

intersections_voo = voo_full %>%
  # arrange(desc(date)) %>% 
  mutate(intersection = close_official >= close_official[date == max(date)],
         intersection_lead = lead(intersection)) %>% 
  filter(intersection != intersection_lead)



```

## S&P 500 Over the Last 10 Years

Omg the stock market just crashed! 1% down in a day is a crash right? Well maybe not... Let's just take a look at the history of the S&P 500 Index Fund (VOO) for some perspective.

```{r sp500_last_2_years}
sp500_2018_pres_plot = voo_full %>% 
  ggplot(aes(x = date, y = close_official)) +
  geom_line() +
  ggtitle("S&P 500")+
  ylab("Closing Price")

sp500_2018_pres_plot %>% ggplotly()
```

## S&P More Recently

Let's zoom in on just the last two months. 

```{r sp500_intersections}
sp_intersections_plot = voo_full %>% 
  filter(date >= min(intersections_voo$date %m-% months(1))) %>% 
  ggplot(aes(x = date, y = close_official)) +
  geom_line() +
  ggtitle("S&P 500")+
  ylab("Closing Price") +
  ggrepel::geom_text_repel(data = intersections_voo, aes(label = as.character(date), segment.color = "red"),
                           box.padding = .8,
                           nudge_y = 20, 
                           direction = "x",
                           vjust = 0) +
  geom_hline(yintercept = voo_full$close_official[voo_full$date ==
                                                             max(voo_full$date)],
             linetype=2, colour="red")

sp_intersections_plot

```

Guess what? If you invested in the S&P 500 prior to `r min(intersections_voo$date)`, then you're still net positive. So calm down.

Also, the largest single day increase from close to close since `r min_date` was on `r max_date` with a single day jump of `r percent(max_change_perc - 1, accuracy = .1)`. So, keep investing and continue to buy the dip.





