---
title: "Lesson 5: Statistics with R"
author: "Cavan Donohoe"
description: "Descriptive statistics, t-tests, correlation, and linear regression — with diagnostics and predictions."
output:
  html_document:
    toc: true
    toc_float: true
    includes:
      in_header: header/header.html
      after_body: include_footer.html
lang: "en"
---

<link rel="stylesheet" href="/css/learn-r.css">

<div class="course-progress">
Lesson 5 of 7 · <a href="learn_r.html">Course overview</a>
</div>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.width = 6.5,
  fig.height = 4.2,
  dpi = 100
)
options(scipen = 999, digits = 4)
```

# Statistics with R

R was built by statisticians, so the basic statistics tools are right there in the base language. This lesson covers the four things you'll do most often:

1. Descriptive statistics — summarising one variable.
2. Comparing two groups — the t-test.
3. Measuring association — correlation.
4. Modeling — linear regression, with diagnostics and predictions.

We'll keep using `mtcars` so you can compare the numbers to the plots from Lesson 4.

## Setup

```{r}
library(dplyr)
library(ggplot2)

cars <- mtcars |>
  tibble::rownames_to_column("model") |>
  tibble::as_tibble()
```

## Descriptive statistics

The most common single-variable summaries:

```{r}
mean(cars$mpg)
median(cars$mpg)
sd(cars$mpg)              # standard deviation
var(cars$mpg)             # variance
min(cars$mpg)
max(cars$mpg)
range(cars$mpg)
quantile(cars$mpg, probs = c(0.25, 0.5, 0.75))
IQR(cars$mpg)
```

`summary()` prints a quick five-number summary plus the mean — handy for a first look at any variable or data frame:

```{r}
summary(cars$mpg)
summary(cars)
```

For grouped summaries, use `dplyr` from Lesson 3:

```{r}
cars |>
  group_by(cyl) |>
  summarise(
    n = n(),
    mean_mpg = mean(mpg),
    sd_mpg = sd(mpg),
    median_mpg = median(mpg),
    .groups = "drop"
  )
```

<div class="callout callout-tip">
<div class="callout-title">💡 Always look before you test</div>
Before running a t-test or fitting a model, make a plot. A boxplot or histogram will tell you in three seconds whether the assumptions you're about to make (normality, equal variance, no crazy outliers) are even close to true.
</div>

## Comparing two groups: the t-test

A t-test answers: "Are the means of two groups different by more than I'd expect from random sampling?"

Question: do manual cars get better gas mileage than automatic cars on average?

```{r fig.height=3.5}
ggplot(cars, aes(x = factor(am, labels = c("automatic", "manual")), y = mpg)) +
  geom_boxplot(fill = "lightsteelblue") +
  labs(x = "Transmission", y = "MPG", title = "MPG by transmission type") +
  theme_minimal()
```

The boxplot suggests yes. Let's quantify it:

```{r}
t_result <- t.test(mpg ~ am, data = cars)
t_result
```

How to read the output:

- **`t = -3.77`** — the test statistic. Bigger absolute value = stronger evidence the means differ.
- **`df`** — degrees of freedom (Welch's t-test uses fractional df, that's fine).
- **`p-value = 0.001374`** — probability of seeing a difference this large (or larger) if the true means were actually equal. By the conventional cutoff of 0.05, this is "significant."
- **`95 percent confidence interval: -11.28, -3.21`** — we're 95% confident the true difference (automatic minus manual) is between about -11 and -3 MPG.
- **`mean in group 0 = 17.15, mean in group 1 = 24.39`** — the actual group means.

So manual cars in this dataset get around **7 MPG more** than automatic cars on average, and the difference is statistically significant.

You can pull individual pieces out programmatically:

```{r}
t_result$p.value
t_result$conf.int
t_result$estimate
```

<div class="callout callout-warning">
<div class="callout-title">⚠️ "Significant" doesn't mean "important"</div>
A tiny effect with a huge sample will be "statistically significant." A huge effect with a small sample may not be. Always look at the confidence interval and the effect size, not just the p-value.

Also: this comparison ignores that manuals tend to be in different cars than automatics (lighter, fewer cylinders, etc.). The honest answer to "does transmission cause better MPG" requires a model that controls for those things — that's what regression is for.
</div>

## Measuring association: correlation

Correlation measures how strongly two numeric variables move together, on a scale of -1 to 1.

```{r}
cor(cars$wt, cars$mpg)
```

A correlation of about -0.87: strong negative association. As weight goes up, MPG goes down.

`cor.test()` adds a hypothesis test (is the correlation different from zero?) and a confidence interval:

```{r}
cor.test(cars$wt, cars$mpg)
```

For a whole correlation matrix:

```{r}
cars |>
  select(mpg, hp, wt, drat, qsec) |>
  cor() |>
  round(2)
```

## Linear regression

Regression is the workhorse of applied statistics. The basic question: "How does the response variable change as the predictors change?"

### A simple regression

Predict MPG from weight:

```{r}
fit1 <- lm(mpg ~ wt, data = cars)
fit1
```

The formula `mpg ~ wt` reads as "model `mpg` as a function of `wt`." `lm()` returns a fitted model object — print it for a quick look, `summary()` for the full output:

```{r}
summary(fit1)
```

What's important here:

- **Coefficients:** `wt = -5.34` means each additional 1000 lbs of weight is associated with a **5.34-MPG decrease**, on average.
- **`Pr(>|t|)` for `wt`:** the p-value for "is this coefficient different from 0?" Tiny — yes.
- **R-squared = 0.75:** 75% of the variation in MPG is explained by weight alone.
- **Residual standard error:** typical prediction error in MPG units.

### A multiple regression

Add more predictors with `+`:

```{r}
fit2 <- lm(mpg ~ wt + hp + factor(cyl), data = cars)
summary(fit2)
```

Now we're estimating the effect of weight **after controlling for** horsepower and cylinder count.

`factor(cyl)` tells R to treat `cyl` as a categorical variable rather than a number. The output gives you the difference between each level and a reference level (here, 4-cylinder).

### Comparing models

`anova()` tests whether the bigger model is significantly better than the smaller one:

```{r}
anova(fit1, fit2)
```

A small p-value means the extra predictors are pulling weight (so to speak).

`AIC()` and `BIC()` are alternative model-comparison metrics — lower is better:

```{r}
AIC(fit1, fit2)
```

### Diagnostics

Before trusting a regression, look at the diagnostic plots. R's built-in `plot()` on a model gives you four:

```{r fig.width=7, fig.height=5}
par(mfrow = c(2, 2))
plot(fit1)
par(mfrow = c(1, 1))
```

What you're looking for:

- **Residuals vs. Fitted** — should look like a flat blob with no pattern. A curve suggests you're missing nonlinearity.
- **Q-Q Residuals** — should hug the diagonal line. Big departures = non-normal residuals.
- **Scale-Location** — should be roughly flat. A trend means the variance changes with the fitted value (heteroscedasticity).
- **Residuals vs. Leverage** — points with high Cook's distance (top-right or bottom-right corners, past the dashed lines) are unusually influential. Investigate those rows.

### Predictions

Once you have a model, you can predict new values:

```{r}
new_cars <- tibble::tibble(wt = c(2.0, 3.5, 5.0))
predict(fit1, newdata = new_cars)
```

Add a confidence interval for the *mean* response:

```{r}
predict(fit1, newdata = new_cars, interval = "confidence")
```

Or a prediction interval for a *single new observation* (wider, because individual cars vary):

```{r}
predict(fit1, newdata = new_cars, interval = "prediction")
```

### Visualizing the fit

A simple regression line is one geom away in ggplot:

```{r fig.height=3.8}
ggplot(cars, aes(x = wt, y = mpg)) +
  geom_point(size = 2.5, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE, color = "steelblue") +
  labs(
    title = "MPG vs. Weight, with linear fit and 95% CI",
    x = "Weight (1000 lbs)", y = "MPG"
  ) +
  theme_minimal()
```

`geom_smooth(method = "lm")` is fitting the same model as `lm(mpg ~ wt)` under the hood and shading the confidence band.

## A few other tests worth knowing

- `wilcox.test()` — non-parametric alternative to the t-test (for non-normal data).
- `chisq.test()` — chi-square test for independence between two categorical variables.
- `prop.test()` — comparing proportions across groups.
- `aov()` — ANOVA, for comparing means across more than two groups.
- `glm(..., family = binomial)` — logistic regression, for binary outcomes.

The pattern is the same across all of them: a formula, a data argument, a summary or print of the result.

## Putting it together

A complete mini-analysis: is MPG associated with weight, controlling for horsepower? Report the model, check assumptions, and predict MPG for a hypothetical new car.

```{r}
fit <- lm(mpg ~ wt + hp, data = cars)
summary(fit)
```

Coefficient interpretations: holding horsepower constant, each additional 1000 lbs of weight is associated with a `r round(coef(fit)["wt"], 2)`-MPG decrease. Holding weight constant, each additional horsepower is associated with a `r round(coef(fit)["hp"], 4)`-MPG decrease.

Quick diagnostic look:

```{r fig.width=7, fig.height=3.5}
par(mfrow = c(1, 2))
plot(fit, which = c(1, 2))
par(mfrow = c(1, 1))
```

Predict MPG for a 3500-lb car with 130 hp:

```{r}
predict(fit, newdata = tibble::tibble(wt = 3.5, hp = 130), interval = "prediction")
```

That's a complete (if minimal) regression analysis: fit, interpret, diagnose, predict.

<div class="exercise">
<div class="exercise-title">✏️ Exercise 5.1 — Descriptive stats</div>

Compute the mean, median, standard deviation, and IQR of `cars$hp`. Are the mean and median similar? What does that tell you about the distribution?

<details>
<summary>Show solution</summary>

```{r}
mean(cars$hp)
median(cars$hp)
sd(cars$hp)
IQR(cars$hp)
```

The mean is noticeably higher than the median, which means the distribution is right-skewed — there are a few high-horsepower outliers pulling the mean up. A boxplot or histogram would confirm.

```{r fig.height=3}
ggplot(cars, aes(x = hp)) +
  geom_histogram(bins = 12, fill = "steelblue", color = "white") +
  theme_minimal()
```

</details>
</div>

<div class="exercise">
<div class="exercise-title">✏️ Exercise 5.2 — A t-test</div>

Test whether 4-cylinder cars have a different mean MPG than 8-cylinder cars. Report the p-value, the 95% CI, and a one-sentence interpretation.

<details>
<summary>Show solution</summary>

```{r}
sub <- cars |> filter(cyl %in% c(4, 8))
t.test(mpg ~ cyl, data = sub)
```

The mean MPG of 4-cylinder cars is roughly 11 MPG higher than 8-cylinder cars (95% CI ~9 to 13), and the p-value is tiny — the difference is far more than chance.

</details>
</div>

<div class="exercise">
<div class="exercise-title">✏️ Exercise 5.3 — A regression</div>

Fit a model that predicts MPG from weight, horsepower, and cylinder count (treat `cyl` as a factor). Which coefficients are statistically significant? What's the R-squared?

<details>
<summary>Show solution</summary>

```{r}
fit <- lm(mpg ~ wt + hp + factor(cyl), data = cars)
summary(fit)
```

`wt` has a clearly significant negative coefficient. `hp` and the cylinder-count contrasts get smaller and may no longer be significant once weight is in the model — they're correlated with weight (heavier cars tend to have more cylinders and more horsepower), so a lot of the "effect" of cylinders is really just weight. Adjusted R-squared should be around 0.83.

</details>
</div>

## What's next

You can now load data, manipulate it, plot it, and run real statistical analyses on it. The last two lessons stop adding new tools and start showing how to **package up an analysis** so someone else (or future-you) can run it again. Lesson 6 is R Markdown — the document format this whole course is written in.

<div class="lesson-nav">
  <a class="nav-prev" href="data_visualization_r.html">
    <span class="lesson-nav-label">← Previous</span>
    <span class="lesson-nav-title">Lesson 4: Data Visualization</span>
  </a>
  <a class="nav-next" href="rmarkdown_r.html">
    <span class="lesson-nav-label">Next →</span>
    <span class="lesson-nav-title">Lesson 6: Reproducible Reports</span>
  </a>
</div>
