---
title: "TOHS Class Reunion"
editor_options: 
  chunk_output_type: console
output:
  html_document:
    includes:
      in_header: header/header.html
      after_body: include_footer.html
lang: "en"
---

```{r pipe, include=FALSE}
`%>%` <- magrittr::`%>%`
```
```{r ci-skip-gs4, include=FALSE}
# if (Sys.getenv("CI") == "true") knitr::knit_exit()
```

```{r gs4-auth, include=FALSE}
# googlesheets4::gs4_deauth()
```


```{r setup, include=FALSE}
# Set default chunk output
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  results = "asis"
)
```

```{r}
snapshot_path <- "data/tohs_reunion.csv"
if (file.exists(snapshot_path)) {
  snapshot <- read.csv(snapshot_path, stringsAsFactors = FALSE)
  grad_list_clean <- dplyr::transmute(
    snapshot, `First Name` = first_name, `Last Name` = last_name,
    email_bool = email_bool
  )
  data_updated_at <- snapshot$updated_at[[1]]
} else {
  stop("Public reunion snapshot missing; run the reunion refresh workflow first")
}
form_url <- trimws(readLines("config/tohs_form_url.txt", warn = FALSE)[[1]])

percent_accounted_text <- sprintf(
  "%s (%d/%d)",
  scales::percent(mean(grad_list_clean$email_bool)),
  sum(grad_list_clean$email_bool), nrow(grad_list_clean)
)

missing_list <- grad_list_clean %>%
  dplyr::filter(!email_bool) %>%
  dplyr::select(`First Name`, `Last Name`) %>%
  dplyr::mutate(text = paste0(dplyr::row_number(), ". ", `First Name`, " ", `Last Name`)) %>%
  dplyr::pull(text) %>%
  paste(collapse = "\n")

```

Help us reach the TOHS Class of 2012 for our next reunion! We currently have emails for **`r percent_accounted_text`** of our class. If you see yourself on the list below, please [update your reunion contact details](`r form_url`). You can also tell us whether you're interested in attending or helping with planning. Your email, phone number, and responses are kept private to reunion organizers.

Also, here is a link to [**the lancer: SENIOR ISSUE 2012**](https://issuu.com/thelancer/docs/seniorissue2012) (our Lancer newspaper), if you want to walk down memory lane.

As of `r data_updated_at`, we have the following missing emails:

```{r}
cat(missing_list)
```
