---
title: "Did They Quit While They Were Ahead?"
editor_options:
  chunk_output_type: console
output:
  html_document:
    includes:
      in_header: header/header.html
      after_body: include_footer.html
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
lang: "en"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

suppressPackageStartupMessages({
  library(dplyr)
})

source(here::here("scripts", "shows_manifest.R"))
source(here::here("scripts", "tv_decline_metrics.R"))

shows <- tv_shows_manifest()
decline <- tv_decline_table(shows)

show_link <- function(title) {
  sprintf("<a href='imdb_rating_plot.html#%s'>%s</a>", anchor_id(title), title)
}

status_label <- function(ongoing) ifelse(ongoing, "still airing", "ended")
```

On [the season-by-season heatmaps](imdb_rating_plot.html) I said I wanted to find the shows
that figure out when to stop. That page draws the maps but never actually answers the
question, so this one does the arithmetic. Of the `r nrow(shows)` shows I track,
`r nrow(decline)` have more than one full season and can actually be ranked.

The idea is simple. For each show I take the average IMDb rating of every season, find its
best season, and compare that peak to where the show is now. A show that ended near its
peak went out on a high note. A show sitting two full points below its best season did not.

```{r headline-numbers}
worst <- decline %>% dplyr::slice_max(peak_to_latest_drop, n = 1, with_ties = FALSE)
on_peak <- decline %>% dplyr::filter(latest_is_peak, !likely_ongoing)
```

The single biggest fall from grace is **`r worst$title`**, which peaked in season
`r worst$peak_season` at `r sprintf("%.2f", worst$peak_avg)` and finished season
`r worst$latest_season` at `r sprintf("%.2f", worst$latest_avg)`, a drop of
`r sprintf("%.2f", worst$peak_to_latest_drop)` points. On the other side,
`r nrow(on_peak)` shows ended their run on their own best season.

## The fall-from-peak leaderboard

Bars are how far a show sits below its own best season. Shows still on the air are shaded
differently, because their story isn't finished. A dip might still turn into a comeback.

```{r drop-chart, fig.width=9, fig.height=8}
top_drops <- decline %>%
  dplyr::filter(peak_to_latest_drop > 0) %>%
  dplyr::slice_max(peak_to_latest_drop, n = 25) %>%
  dplyr::mutate(
    status = status_label(likely_ongoing),
    label = paste0(
      title, "<br>peak S", peak_season, ": ", sprintf("%.2f", peak_avg),
      "<br>latest S", latest_season, ": ", sprintf("%.2f", latest_avg),
      "<br>drop: ", sprintf("%.2f", peak_to_latest_drop)
    )
  )

plotly::plot_ly(
  top_drops,
  x = ~peak_to_latest_drop,
  y = ~reorder(title, peak_to_latest_drop),
  color = ~status,
  colors = c("ended" = "#b2182b", "still airing" = "#7fa8b8"),
  type = "bar",
  orientation = "h",
  text = ~label,
  hoverinfo = "text"
) %>%
  plotly::layout(
    xaxis = list(title = "Rating points below the show's best season"),
    yaxis = list(title = ""),
    legend = list(orientation = "h", x = 0, y = 1.05),
    margin = list(l = 180)
  )
```

## Shows that stuck the landing

These shows ended on their highest-rated season, as far as the episode data goes. This is
the list the whole page was built to find.

```{r stuck-landing}
on_peak %>%
  dplyr::arrange(dplyr::desc(latest_avg)) %>%
  dplyr::transmute(
    Show = show_link(title),
    Seasons = seasons,
    `Final season` = latest_season,
    `Final season avg` = round(latest_avg, 2),
    `Last episode` = last_year
  ) %>%
  DT::datatable(
    escape = FALSE,
    rownames = FALSE,
    options = list(pageLength = 15, autoWidth = TRUE, order = list(list(3, "desc"))),
    caption = "Shows whose last full season was also their best. Click a title for its heatmap."
  )
```

## The slow slide

The leaderboard above rewards one bad season. This looks at the whole trajectory instead:
the slope of a straight line through a show's season averages, in rating points per season.
I only include shows with at least four scored seasons, because a slope drawn through two
points is just a rumour.

```{r slope-table}
decline %>%
  dplyr::filter(seasons_scored >= 4) %>%
  dplyr::arrange(slope_per_season) %>%
  dplyr::transmute(
    Show = show_link(title),
    Status = status_label(likely_ongoing),
    `Seasons scored` = seasons_scored,
    `Points per season` = round(slope_per_season, 3),
    `Peak` = round(peak_avg, 2),
    `Latest` = round(latest_avg, 2)
  ) %>%
  DT::datatable(
    escape = FALSE,
    rownames = FALSE,
    filter = "top",
    options = list(pageLength = 15, autoWidth = TRUE, order = list(list(3, "asc"))),
    caption = "Negative means the show trends down season over season."
  )
```

## Everything, all at once

Every metric for every show, sortable and filterable.

```{r full-table}
decline %>%
  dplyr::arrange(dplyr::desc(peak_to_latest_drop)) %>%
  dplyr::transmute(
    Show = show_link(title),
    Status = status_label(likely_ongoing),
    Seasons = seasons,
    Episodes = episodes,
    `Peak season` = peak_season,
    `Peak avg` = round(peak_avg, 2),
    `Latest season` = latest_season,
    `Latest avg` = round(latest_avg, 2),
    Drop = round(peak_to_latest_drop, 2),
    `Pts/season` = round(slope_per_season, 3),
    `Episode SD` = round(episode_sd, 2),
    `Last episode` = last_year
  ) %>%
  DT::datatable(
    escape = FALSE,
    rownames = FALSE,
    filter = "top",
    options = list(pageLength = 20, autoWidth = TRUE, scrollX = TRUE, order = list(list(8, "desc"))),
    caption = "All tracked shows. Episode SD is how much individual episodes bounce around."
  )
```

## How this is calculated, and where it's unfair

A few things worth knowing before you take any of this personally:

- **Seasons with fewer than three episodes are ignored.** Otherwise a show two episodes into
  a new season would look like it had collapsed.
- **"Still airing" is a guess.** I flag a show as ongoing if its most recent episode landed
  within the last year or so. There's no cancellation data here, so a show on hiatus may be
  mislabelled either way.
- **Long-running comedies are punished by design.** The Simpsons, South Park, and Family Guy
  all peaked in the nineties or early 2000s and have been running ever since. Comparing
  season 37 to season 7 is a big drop, but it isn't really the same show, and "still on the
  air after 37 seasons" is its own kind of win.
- **Ratings drift with the audience, not just quality.** Early seasons get rated years later
  by people who already decided they love the show. Finales get rated the week they air by
  everyone who's angry. That asymmetry inflates exactly the drop this page measures.
- **A high peak makes falling easier.** A show that never went above 7.5 has less room to
  drop than one that hit 9.3.

The episode data refreshes automatically from IMDb, so these numbers move on their own. The
metric code lives in `scripts/tv_decline_metrics.R` and has tests, in case you want to argue
with the arithmetic rather than the conclusions.
