---
title: "US Rental Investment Markets"
editor_options:
  chunk_output_type: console
output:
  html_document:
    includes:
      in_header: header/header.html
      after_body: include_footer.html
---

```{r pipe, include=FALSE}
`%>%` <- magrittr::`%>%`
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.width = 10, fig.height = 8
)
```

```{r data}
raw <- readr::read_csv(here::here("data", "us_rental_markets.csv"), show_col_types = FALSE)

# Filter out noise: seasonal vacation rentals and extremely low-value markets
markets <- raw %>%
  dplyr::filter(
    zhvi >= 50000,
    zori <= 5000,
    zori >= 500,
    gross_yield_pct <= 20
  ) %>%
  dplyr::mutate(
    tier = factor(tier, levels = c("Strong Buy", "Moderate", "Lean Rent", "Rent Favored")),
    city_state = paste0(city, ", ", state)
  )

zhvi_date <- unique(raw$zhvi_date)[1]
zori_date <- unique(raw$zori_date)[1]

tier_colors <- c(
  "Strong Buy"   = "#27AE60",
  "Moderate"     = "#2E86C1",
  "Lean Rent"    = "#D4760A",
  "Rent Favored" = "#8B0000"
)
```

Where should you buy rental property in the US? This page ranks **`r nrow(markets)` cities** by investment metrics derived from Zillow's public home value and rent indices. The price-to-rent ratio is the standard benchmark:

- **Strong Buy** (P/R < 15): Cash flow positive from day one
- **Moderate** (P/R 15--20): Reasonable returns, some appreciation upside
- **Lean Rent** (P/R 20--25): Marginal; depends on appreciation
- **Rent Favored** (P/R > 25): Buy only if you're betting on price growth

### Data Sources

All data is traceable and reproducible. Run `Rscript scripts/us_rental_markets_data.R` to refresh.

| Source | What | Updated |
|--------|------|---------|
| [Zillow ZHVI](https://www.zillow.com/research/data/) | Typical home values (city level, SFR+condo, 35th--65th percentile) | `r zhvi_date` |
| [Zillow ZORI](https://www.zillow.com/research/data/) | Typical observed rents (city level, SFR+condo+MFR) | `r zori_date` |

Estimated cap rate assumes 35% expense ratio (taxes, insurance, vacancy, maintenance, management). Raw data: [us_rental_markets.csv](data/us_rental_markets.csv)

## Top 40 Cities by Cap Rate {.tabset}

```{r top_markets}
top40 <- markets %>%
  dplyr::slice_head(n = 40)
```

### Bar Chart

```{r cap_rate_bar, fig.height=12}
top40 %>%
  dplyr::mutate(city_state = forcats::fct_reorder(city_state, est_cap_rate_pct)) %>%
  ggplot2::ggplot(ggplot2::aes(x = city_state, y = est_cap_rate_pct, fill = tier)) +
  ggplot2::geom_col(width = 0.7) +
  ggplot2::geom_text(
    ggplot2::aes(label = paste0(est_cap_rate_pct, "%")),
    hjust = -0.1, size = 3, color = "grey30"
  ) +
  ggplot2::coord_flip(ylim = c(0, max(top40$est_cap_rate_pct) * 1.15)) +
  ggplot2::scale_fill_manual(values = tier_colors) +
  ggplot2::labs(
    title = "Estimated Cap Rate (Top 40 Cities)",
    subtitle = "Gross yield minus 35% expense ratio",
    x = NULL, y = "Estimated Cap Rate (%)", fill = "Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    legend.position = "top",
    panel.grid.major.y = ggplot2::element_blank()
  )
```

### Interactive Table

```{r top_table}
top40 %>%
  dplyr::transmute(
    `#` = dplyr::row_number(),
    City = city, State = state, Metro = metro,
    `Home Value` = scales::dollar(zhvi),
    `Rent/Mo` = scales::dollar(zori, accuracy = 1),
    `Gross Yield` = paste0(gross_yield_pct, "%"),
    `Est Cap Rate` = paste0(est_cap_rate_pct, "%"),
    `P/R Ratio` = price_to_rent,
    `ZHVI YoY` = paste0(zhvi_yoy_pct, "%"),
    `ZORI YoY` = dplyr::if_else(is.na(zori_yoy_pct), "---", paste0(zori_yoy_pct, "%")),
    Tier = tier
  ) %>%
  DT::datatable(
    filter = "top", extensions = "Buttons",
    options = list(
      paging = FALSE, dom = "Bfrtip",
      buttons = c("copy", "csv", "excel"),
      scrollY = "500px", scrollX = TRUE
    )
  )
```

## Gross Yield vs Home Value

```{r scatter_yield, fig.height=8}
top100 <- markets %>% dplyr::slice_head(n = 100)

ggplot2::ggplot(top100, ggplot2::aes(x = zhvi, y = gross_yield_pct, color = tier)) +
  ggplot2::geom_point(size = 3, alpha = 0.8) +
  ggrepel::geom_text_repel(
    ggplot2::aes(label = city_state),
    size = 2.5, max.overlaps = 20, segment.color = "grey70"
  ) +
  ggplot2::scale_x_continuous(labels = scales::label_dollar(scale_cut = scales::cut_short_scale())) +
  ggplot2::scale_color_manual(values = tier_colors) +
  ggplot2::labs(
    title = "Gross Yield vs Home Value (Top 100 Cities)",
    subtitle = "Higher yield + lower price = better cash flow market",
    x = "Zillow Home Value Index", y = "Gross Yield (%)", color = "Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(legend.position = "top")
```

## Full US Market Explorer

Search, filter, and sort all `r nrow(markets)` cities.

```{r full_table}
markets %>%
  dplyr::transmute(
    City = city, State = state, Metro = metro,
    `Home Value` = scales::dollar(zhvi),
    `ZHVI YoY` = paste0(zhvi_yoy_pct, "%"),
    `Rent/Mo` = scales::dollar(zori, accuracy = 1),
    `ZORI YoY` = dplyr::if_else(is.na(zori_yoy_pct), "---", paste0(zori_yoy_pct, "%")),
    `Gross Yield` = gross_yield_pct,
    `Est Cap Rate` = est_cap_rate_pct,
    `P/R Ratio` = price_to_rent,
    `Rent/Price %` = rent_to_price_pct,
    Tier = tier
  ) %>%
  DT::datatable(
    filter = "top", extensions = "Buttons",
    options = list(
      pageLength = 50, dom = "Bfrtip",
      buttons = c("copy", "csv", "excel"),
      scrollX = TRUE,
      order = list(list(8, "desc"))
    )
  )
```

## Tier Distribution

```{r tier_dist, fig.height=5, fig.width=8}
tier_counts <- markets %>%
  dplyr::count(tier) %>%
  dplyr::mutate(pct = round(100 * n / sum(n), 1))

ggplot2::ggplot(tier_counts, ggplot2::aes(x = tier, y = n, fill = tier)) +
  ggplot2::geom_col(width = 0.6) +
  ggplot2::geom_text(
    ggplot2::aes(label = paste0(n, " (", pct, "%)")),
    vjust = -0.3, size = 4
  ) +
  ggplot2::scale_fill_manual(values = tier_colors) +
  ggplot2::scale_y_continuous(expand = ggplot2::expansion(mult = c(0, 0.15))) +
  ggplot2::labs(
    title = "How Many US Cities Fall Into Each Investment Tier?",
    x = NULL, y = "Number of Cities"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(legend.position = "none")
```

## Metro-Level Comparison

Average metrics across cities within each metro area (metros with 5+ cities).

```{r metro_table}
metro_summary <- markets %>%
  dplyr::filter(!is.na(metro)) %>%
  dplyr::group_by(metro) %>%
  dplyr::summarise(
    n_cities = dplyr::n(),
    avg_zhvi = mean(zhvi, na.rm = TRUE),
    avg_zori = mean(zori, na.rm = TRUE),
    avg_gross_yield = mean(gross_yield_pct, na.rm = TRUE),
    avg_cap_rate = mean(est_cap_rate_pct, na.rm = TRUE),
    avg_ptr = mean(price_to_rent, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::filter(n_cities >= 5) %>%
  dplyr::arrange(dplyr::desc(avg_cap_rate))

metro_summary %>%
  dplyr::transmute(
    Metro = metro,
    Cities = n_cities,
    `Avg Home Value` = scales::dollar(avg_zhvi),
    `Avg Rent/Mo` = scales::dollar(avg_zori, accuracy = 1),
    `Avg Gross Yield` = paste0(round(avg_gross_yield, 2), "%"),
    `Avg Cap Rate` = paste0(round(avg_cap_rate, 2), "%"),
    `Avg P/R Ratio` = round(avg_ptr, 1)
  ) %>%
  DT::datatable(
    filter = "top", extensions = "Buttons",
    options = list(
      pageLength = 30, dom = "Bfrtip",
      buttons = c("copy", "csv", "excel"),
      scrollX = TRUE,
      order = list(list(5, "desc"))
    )
  )
```

---

**How to update this data:** Run `Rscript scripts/us_rental_markets_data.R` to download the latest Zillow ZHVI + ZORI and recompute all metrics. Then re-render this page.
