---
title: "US Rental Investment Markets"
editor_options:
  chunk_output_type: console
output:
  html_document:
    includes:
      in_header: header/header.html
      after_body: include_footer.html
---

```{r og-meta, echo=FALSE, results='asis'}
cat('
<meta property="og:title" content="US Rental Investment Markets">
<meta property="og:description" content="Cap rates, vacancy rates, and demand signals for 3,800+ US cities using Zillow and Census data.">
<meta property="og:url" content="https://cavandonohoe.github.io/us_rental_markets.html">
<meta property="og:type" content="article">
')
```

```{r pipe, include=FALSE}
`%>%` <- magrittr::`%>%`
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.width = 10, fig.height = 8
)
```

```{r data}
raw <- readr::read_csv(here::here("data", "us_rental_markets.csv"), show_col_types = FALSE)
```

```{r timestamp, results='asis'}
data_date <- max(raw$zhvi_date, raw$zori_date, na.rm = TRUE)
cat(sprintf(
  '<span class="data-timestamp">Zillow data as of: %s</span>\n',
  format(data_date, "%B %d, %Y")
))
```

```{r filter}
# Filter out noise: seasonal vacation rentals and extremely low-value markets
markets <- raw %>%
  dplyr::filter(
    zhvi >= 50000,
    zori <= 5000,
    zori >= 500,
    gross_yield_pct <= 20
  ) %>%
  dplyr::mutate(
    tier = factor(tier, levels = c("Strong Buy", "Moderate", "Lean Rent", "Rent Favored")),
    city_state = paste0(city, ", ", state)
  )

zhvi_date <- unique(raw$zhvi_date)[1]
zori_date <- unique(raw$zori_date)[1]

# Default leverage assumptions (match scripts/us_rental_demand_data.R)
DEFAULT_DOWN_PCT <- 25
DEFAULT_MORTGAGE_RATE <- 7.25
DEFAULT_TERM_YEARS <- 30
DEFAULT_EXPENSE_RATIO <- 35
DEFAULT_CLOSING_PCT <- 3

tier_colors <- c(
  "Strong Buy"   = "#27AE60",
  "Moderate"     = "#2E86C1",
  "Lean Rent"    = "#D4760A",
  "Rent Favored" = "#8B0000"
)
```

Where should you buy rental property in the US? This page ranks **`r nrow(markets)` cities** by investment metrics derived from Zillow's public home value and rent indices plus Census demand signals -- and a few unconventional ones like **Trader Joe's adjacency**, a classic real-estate proxy for walkable, amenity-rich neighborhoods with strong demographics.

Jump to: [Cash Flow Calculator](#interactive-cash-flow-calculator) · [Trader Joe's Adjacency](#the-trader-joes-adjacency-premium) · [Unicorn Leaderboard](#unicorn-leaderboard-cash-flow-trader-joes) · [TJ Map](#trader-joes-investment-map) · [Leveraged Returns](#leveraged-returns-at-25-down-7.25-30yr) · [Hidden Gems](#hidden-gems-small-towns-with-big-cash-flow) · [Investor Rules](#investor-rules-of-thumb) · [Composite Score](#composite-investment-score) · [State Explorer](#state-leaderboard) · [Price vs Rent Growth](#price-vs-rent-growth-yoy) · [City Lookup](#per-city-detail-lookup)

The price-to-rent ratio is the standard cash-flow benchmark:

- **Strong Buy** (P/R < 15): Cash flow positive from day one
- **Moderate** (P/R 15--20): Reasonable returns, some appreciation upside
- **Lean Rent** (P/R 20--25): Marginal; depends on appreciation
- **Rent Favored** (P/R > 25): Buy only if you're betting on price growth

### Data Sources

All data is traceable and reproducible. Run `Rscript scripts/us_rental_markets_data.R` to refresh.

```{r sources_table}
tibble::tribble(
  ~Source, ~Description, ~Updated,
  "Zillow ZHVI", "Typical home values (city level, SFR+condo, 35th\u201365th percentile)", as.character(zhvi_date),
  "Zillow ZORI", "Typical observed rents (city level, SFR+condo+MFR)", as.character(zori_date),
  "Census Pop Estimates", "City + county population (2020\u20132024), net migration", "V2024",
  "Census ACS 5-Year", "Vacancy rates, renter %, housing counts (county level)", "2019\u20132023",
  "Census Places Gazetteer", "Lat/lon for every incorporated place (for distance calcs)", "2024",
  "Trader Joe's Store Locator", paste0("All ", sum(!is.na(unique(raw$nearest_tj_city))), " US stores, scraped from locations.traderjoes.com"), format(Sys.Date(), "%Y")
) %>%
  gt::gt() %>%
  gt::tab_header(
    title = gt::md("**Data Sources**"),
    subtitle = "All public, no API key required"
  ) %>%
  gt::fmt_markdown(columns = Source) %>%
  gt::cols_width(
    Source ~ gt::px(180),
    Description ~ gt::px(400),
    Updated ~ gt::px(130)
  ) %>%
  gt::cols_align(align = "left") %>%
  gt::tab_source_note(
    gt::md("Cap rate = gross yield \u00d7 (1 \u2212 35% expense ratio). Raw data: [us_rental_markets.csv](data/us_rental_markets.csv)")
  ) %>%
  gt::tab_style(
    style = gt::cell_fill(color = "#f8f9fa"),
    locations = gt::cells_body(rows = seq(1, 6, by = 2))
  ) %>%
  gt::tab_options(
    table.font.size = gt::px(14),
    heading.title.font.size = gt::px(16),
    heading.subtitle.font.size = gt::px(13),
    column_labels.font.weight = "bold",
    column_labels.background.color = "#2C3E50",
    table.border.top.color = "#2C3E50",
    table.border.bottom.color = "#2C3E50",
    source_notes.font.size = gt::px(12)
  ) %>%
  gt::tab_style(
    style = gt::cell_text(color = "white"),
    locations = gt::cells_column_labels()
  )
```

## Top 40 Cities by Cap Rate {.tabset}

```{r top_markets}
top40 <- markets %>%
  dplyr::slice_head(n = 40)
```

### Bar Chart

```{r cap_rate_bar, fig.height=12}
top40 %>%
  dplyr::mutate(city_state = forcats::fct_reorder(city_state, est_cap_rate_pct)) %>%
  ggplot2::ggplot(ggplot2::aes(x = city_state, y = est_cap_rate_pct, fill = tier)) +
  ggplot2::geom_col(width = 0.7) +
  ggplot2::geom_text(
    ggplot2::aes(label = paste0(est_cap_rate_pct, "%")),
    hjust = -0.1, size = 3, color = "grey30"
  ) +
  ggplot2::coord_flip(ylim = c(0, max(top40$est_cap_rate_pct) * 1.15)) +
  ggplot2::scale_fill_manual(values = tier_colors) +
  ggplot2::labs(
    title = "Estimated Cap Rate (Top 40 Cities)",
    subtitle = "Gross yield minus 35% expense ratio",
    x = NULL, y = "Estimated Cap Rate (%)", fill = "Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    legend.position = "top",
    panel.grid.major.y = ggplot2::element_blank()
  )
```

### Interactive Table

<details class="table-accordion"><summary>Show table</summary>

```{r top_table}
top40 %>%
  dplyr::transmute(
    `#` = dplyr::row_number(),
    City = city, State = state, Metro = metro,
    `Home Value` = scales::dollar(zhvi),
    `Rent/Mo` = scales::dollar(zori, accuracy = 1),
    `Est Cap Rate` = paste0(est_cap_rate_pct, "%"),
    `P/R Ratio` = price_to_rent,
    Population = dplyr::if_else(is.na(pop_2024), "---", scales::comma(pop_2024)),
    `Pop Growth` = dplyr::if_else(is.na(pop_growth_pct), "---", paste0(pop_growth_pct, "%")),
    `Vacancy` = dplyr::if_else(is.na(vacancy_rate_pct), "---", paste0(vacancy_rate_pct, "%")),
    `Rental Vac` = dplyr::if_else(is.na(rental_vacancy_pct), "---", paste0(rental_vacancy_pct, "%")),
    `Net Mig/1K` = dplyr::if_else(is.na(net_mig_rate_per1k), "---", as.character(net_mig_rate_per1k)),
    `% Renter` = dplyr::if_else(is.na(renter_pct), "---", paste0(renter_pct, "%")),
    Tier = tier
  ) %>%
  DT::datatable(
    filter = "top", extensions = "Buttons",
    options = list(
      pageLength = 500, dom = "Bfrtip",
      buttons = c("copy", "csv", "excel"),
      scrollY = "400px", scrollX = TRUE
    )
  )
```

</details>

## Interactive Cash Flow Calculator

Plug in **your** financing assumptions and rerank every market by monthly cash flow, cash-on-cash return, and whether it actually clears the mortgage. Defaults are 25% down, 7.25% 30-yr fixed, 35% operating expenses, 3% closing costs.

<div class="calc-box">
<div class="calc-inputs">
  <label>Down payment (%) <input id="calc-down" type="number" value="25" min="0" max="100" step="1"></label>
  <label>Mortgage rate (%) <input id="calc-rate" type="number" value="7.25" min="0" max="20" step="0.125"></label>
  <label>Term (years) <input id="calc-term" type="number" value="30" min="5" max="40" step="1"></label>
  <label>Operating expense (%) <input id="calc-opex" type="number" value="35" min="0" max="80" step="1"></label>
  <label>Closing costs (%) <input id="calc-closing" type="number" value="3" min="0" max="10" step="0.5"></label>
  <label>Min population <input id="calc-minpop" type="number" value="10000" min="0" step="1000"></label>
  <label>Max home value ($) <input id="calc-maxprice" type="number" value="500000" min="50000" step="25000"></label>
  <label>Max miles to TJ's <input id="calc-tjmiles" type="number" value="9999" min="0" max="9999" step="1" title="Set to 9999 to disable the TJ proximity filter"></label>
  <label>Rows to show <input id="calc-rows" type="number" value="40" min="5" max="500" step="5"></label>
  <label>Sort by
    <select id="calc-sort">
      <option value="monthly_cash_flow">Monthly cash flow</option>
      <option value="cash_on_cash_pct">Cash-on-cash return</option>
      <option value="est_cap_rate_pct">Cap rate</option>
      <option value="investment_score">Composite score</option>
      <option value="tj_miles_sort">Distance to Trader Joe's</option>
    </select>
  </label>
</div>
<div id="calc-summary" class="calc-summary"></div>
<div class="calc-table-wrap"><table id="calc-table"><thead><tr>
  <th>#</th><th>City</th><th>State</th>
  <th class="num">Price</th><th class="num">Rent/mo</th>
  <th class="num">Mortgage</th><th class="num">Cash flow/mo</th>
  <th class="num">CoC %</th><th class="num">Cap rate</th>
  <th class="num">Pop</th><th class="num">TJ mi</th>
  <th class="num">Vacancy</th><th>Tier</th>
</tr></thead><tbody></tbody></table></div>
</div>

```{r calc_data, results='asis'}
# Embed a compact JSON blob so the calculator can rerun on every keystroke.
calc_rows <- markets %>%
  dplyr::filter(!is.na(zhvi), !is.na(zori)) %>%
  dplyr::transmute(
    city, state, metro,
    zhvi = round(zhvi, 0),
    zori = round(zori, 0),
    cap = est_cap_rate_pct,
    ptr = price_to_rent,
    pop = tidyr::replace_na(pop_2024, -1),
    pop_gr = tidyr::replace_na(pop_growth_pct, NA_real_),
    vac = tidyr::replace_na(vacancy_rate_pct, NA_real_),
    score = tidyr::replace_na(investment_score, NA_real_),
    tj_miles = tidyr::replace_na(tj_miles, NA_real_),
    nearest_tj = tidyr::replace_na(nearest_tj_city, "---"),
    tier = as.character(tier)
  )

calc_json <- jsonlite::toJSON(calc_rows, dataframe = "rows", na = "null", auto_unbox = TRUE)
cat(sprintf('<script id="calc-data" type="application/json">%s</script>\n', calc_json))
```

<style>
.calc-box { border: 1px solid var(--card-border); background: var(--card-bg);
  border-radius: 8px; padding: 16px; margin: 1em 0; }
.calc-inputs { display: grid; grid-template-columns: repeat(auto-fit, minmax(160px, 1fr));
  gap: 10px 16px; margin-bottom: 12px; }
.calc-inputs label { font-size: 0.82em; color: var(--muted-text); display: flex;
  flex-direction: column; gap: 3px; }
.calc-inputs input, .calc-inputs select {
  padding: 5px 8px; border: 1px solid var(--card-border);
  border-radius: 4px; background: var(--bg-color); color: var(--text-color);
  font-size: 0.95em;
}
.calc-summary { padding: 8px 10px; margin-bottom: 10px; font-size: 0.92em;
  background: var(--well-bg); border-radius: 4px; color: var(--text-color); }
.calc-summary strong { color: var(--link-color); }
.calc-table-wrap { overflow-x: auto; max-height: 520px; overflow-y: auto; }
#calc-table { width: 100%; border-collapse: collapse; font-size: 0.88em; }
#calc-table th, #calc-table td { padding: 6px 10px; border-bottom: 1px solid var(--card-border); }
#calc-table thead th { position: sticky; top: 0; background: var(--card-bg);
  text-align: left; font-weight: 600; }
#calc-table td.num, #calc-table th.num { text-align: right; font-variant-numeric: tabular-nums; }
#calc-table tr:hover td { background: var(--well-bg); }
.cf-pos { color: #27AE60; font-weight: 600; }
.cf-neg { color: #C0392B; font-weight: 600; }
.tier-pill { display: inline-block; padding: 1px 7px; border-radius: 10px;
  font-size: 0.78em; color: #fff; }
</style>

<script>
(function() {
  var data = JSON.parse(document.getElementById('calc-data').textContent);
  var tierColors = {
    "Strong Buy": "#27AE60",
    "Moderate": "#2E86C1",
    "Lean Rent": "#D4760A",
    "Rent Favored": "#8B0000"
  };

  function dollar(x) {
    if (x === null || isNaN(x)) return "---";
    var sign = x < 0 ? "-" : "";
    return sign + "$" + Math.abs(Math.round(x)).toLocaleString();
  }
  function pct(x, d) {
    if (x === null || isNaN(x)) return "---";
    return x.toFixed(d === undefined ? 2 : d) + "%";
  }
  function comma(x) {
    if (x === null || isNaN(x) || x < 0) return "---";
    return Math.round(x).toLocaleString();
  }
  function pmt(loan, annualRate, years) {
    var r = annualRate / 100 / 12;
    var n = years * 12;
    if (r === 0) return loan / n;
    return loan * (r * Math.pow(1 + r, n)) / (Math.pow(1 + r, n) - 1);
  }

  function recompute() {
    var dp = +document.getElementById('calc-down').value;
    var rate = +document.getElementById('calc-rate').value;
    var term = +document.getElementById('calc-term').value;
    var opex = +document.getElementById('calc-opex').value / 100;
    var closing = +document.getElementById('calc-closing').value / 100;
    var minPop = +document.getElementById('calc-minpop').value;
    var maxPrice = +document.getElementById('calc-maxprice').value;
    var maxTjMi = +document.getElementById('calc-tjmiles').value;
    var nRows = +document.getElementById('calc-rows').value;
    var sortKey = document.getElementById('calc-sort').value;

    var rows = data
      .filter(function(r) {
        if (r.zhvi > maxPrice) return false;
        if (minPop > 0 && (r.pop === null || r.pop < minPop)) return false;
        if (maxTjMi < 9999 && (r.tj_miles === null || r.tj_miles > maxTjMi)) return false;
        return true;
      })
      .map(function(r) {
        var loan = r.zhvi * (1 - dp / 100);
        var mort = pmt(loan, rate, term);
        var cashInv = r.zhvi * (dp / 100) + r.zhvi * closing;
        var cf = r.zori - mort - r.zori * opex;
        var coc = cashInv > 0 ? (cf * 12 / cashInv) * 100 : null;
        return Object.assign({}, r, {
          monthly_mortgage: mort,
          monthly_cash_flow: cf,
          cash_on_cash_pct: coc,
          est_cap_rate_pct: r.cap,
          investment_score: r.score,
          // For sorting: lower is better, so we invert (so bigger-is-better rule still works).
          tj_miles_sort: (r.tj_miles === null || isNaN(r.tj_miles)) ? null : -r.tj_miles
        });
      });

    rows.sort(function(a, b) {
      var av = a[sortKey], bv = b[sortKey];
      if (av === null || isNaN(av)) return 1;
      if (bv === null || isNaN(bv)) return -1;
      return bv - av;
    });

    var pos = rows.filter(function(r) { return r.monthly_cash_flow > 0; }).length;
    var summary = document.getElementById('calc-summary');
    summary.innerHTML =
      "<strong>" + rows.length.toLocaleString() + "</strong> cities match your filters. " +
      "<strong>" + pos.toLocaleString() + "</strong> (" +
      (rows.length ? (100 * pos / rows.length).toFixed(1) : "0") +
      "%) cash flow positive at " + dp + "% down / " + rate + "% rate.";

    var tbody = document.querySelector('#calc-table tbody');
    tbody.innerHTML = "";
    rows.slice(0, nRows).forEach(function(r, i) {
      var cfClass = r.monthly_cash_flow >= 0 ? "cf-pos" : "cf-neg";
      var tierHtml = r.tier
        ? '<span class="tier-pill" style="background:' + (tierColors[r.tier] || "#555") + '">' + r.tier + '</span>'
        : "";
      var tjDisp = (r.tj_miles === null || isNaN(r.tj_miles)) ? "---" : r.tj_miles.toFixed(1);
      var tjTitle = (r.nearest_tj && r.nearest_tj !== "---")
        ? "Nearest TJ: " + r.nearest_tj
        : "";
      var tr = document.createElement('tr');
      tr.innerHTML =
        "<td>" + (i + 1) + "</td>" +
        "<td>" + r.city + "</td>" +
        "<td>" + r.state + "</td>" +
        '<td class="num">' + dollar(r.zhvi) + "</td>" +
        '<td class="num">' + dollar(r.zori) + "</td>" +
        '<td class="num">' + dollar(r.monthly_mortgage) + "</td>" +
        '<td class="num ' + cfClass + '">' + dollar(r.monthly_cash_flow) + "</td>" +
        '<td class="num">' + pct(r.cash_on_cash_pct, 2) + "</td>" +
        '<td class="num">' + pct(r.cap, 2) + "</td>" +
        '<td class="num">' + comma(r.pop) + "</td>" +
        '<td class="num" title="' + tjTitle + '">' + tjDisp + "</td>" +
        '<td class="num">' + pct(r.vac, 1) + "</td>" +
        "<td>" + tierHtml + "</td>";
      tbody.appendChild(tr);
    });
  }

  var inputs = document.querySelectorAll('#calc-down, #calc-rate, #calc-term, #calc-opex, #calc-closing, #calc-minpop, #calc-maxprice, #calc-tjmiles, #calc-rows, #calc-sort');
  inputs.forEach(function(el) { el.addEventListener('input', recompute); });
  recompute();
})();
</script>

## The Trader Joe's Adjacency Premium

Real-estate folks have long used Trader Joe's as a proxy for walkable, amenity-rich, demographically-solid neighborhoods. Every one of the **`r length(unique(markets$nearest_tj_city[markets$tj_within_25mi]))` TJ store cities** on this page lives near one of **`r sum(!is.na(unique(markets$nearest_tj_city)))`** US Trader Joe's, scraped fresh from the official store locator.

Here's how TJ-adjacent cities (within 10 miles of a store) compare to everyone else on the metrics investors care about:

```{r tj_premium_overall, fig.height=5, fig.width=10}
tj_premium <- markets %>%
  dplyr::filter(!is.na(tj_miles)) %>%
  dplyr::mutate(
    tj_group = dplyr::case_when(
      tj_within_2mi ~ "< 2 miles",
      tj_within_5mi ~ "2-5 miles",
      tj_within_10mi ~ "5-10 miles",
      tj_within_25mi ~ "10-25 miles",
      TRUE ~ "> 25 miles"
    ),
    tj_group = factor(
      tj_group,
      levels = c("< 2 miles", "2-5 miles", "5-10 miles", "10-25 miles", "> 25 miles")
    )
  )

summary_tbl <- tj_premium %>%
  dplyr::group_by(tj_group) %>%
  dplyr::summarise(
    n_cities = dplyr::n(),
    median_price = median(zhvi, na.rm = TRUE),
    median_rent = median(zori, na.rm = TRUE),
    median_yield = median(gross_yield_pct, na.rm = TRUE),
    median_cap = median(est_cap_rate_pct, na.rm = TRUE),
    median_price_yoy = median(zhvi_yoy_pct, na.rm = TRUE),
    median_rent_yoy = median(zori_yoy_pct, na.rm = TRUE),
    median_pop_growth = median(pop_growth_pct, na.rm = TRUE),
    .groups = "drop"
  )

summary_long <- summary_tbl %>%
  dplyr::select(tj_group, `Median home value ($)` = median_price,
                `Median rent ($/mo)` = median_rent,
                `Home price YoY (%)` = median_price_yoy,
                `Rent YoY (%)` = median_rent_yoy) %>%
  tidyr::pivot_longer(-tj_group, names_to = "metric", values_to = "value")

ggplot2::ggplot(summary_long, ggplot2::aes(x = tj_group, y = value, fill = tj_group)) +
  ggplot2::geom_col(show.legend = FALSE) +
  ggplot2::geom_text(
    ggplot2::aes(label = ifelse(grepl("\\$", metric),
                                scales::dollar(round(value, 0)),
                                paste0(round(value, 1), "%"))),
    vjust = -0.3, size = 3
  ) +
  ggplot2::facet_wrap(~ metric, scales = "free_y", nrow = 1) +
  ggplot2::scale_fill_manual(
    values = c("< 2 miles" = "#D4760A", "2-5 miles" = "#E8894B", "5-10 miles" = "#F2B181",
               "10-25 miles" = "#A6A6A6", "> 25 miles" = "#555555")
  ) +
  ggplot2::scale_y_continuous(expand = ggplot2::expansion(mult = c(0, 0.15))) +
  ggplot2::labs(
    title = "Cities closer to a Trader Joe's have higher prices, higher rents, and faster growth",
    subtitle = "Median across all cities in each distance bucket",
    x = NULL, y = NULL
  ) +
  ggplot2::theme_minimal(base_size = 11) +
  ggplot2::theme(
    axis.text.x = ggplot2::element_text(angle = 25, hjust = 1),
    strip.text = ggplot2::element_text(face = "bold")
  )
```

The premium is real: cities within 2 miles of a TJ carry substantially higher prices and rents, with stronger YoY growth, than far-flung markets. That's not surprising -- **TJ picks its own locations carefully**, favoring areas with median income, education, and density profiles that are already strong.

The practical implication for an investor: **proximity to a TJ is a useful check on whether a cash-flow market is also demographically durable**. Cheap cap rates in dying small towns are a trap; cheap cap rates in neighborhoods near a Trader Joe's are usually worth a closer look.

```{r tj_premium_table}
summary_tbl %>%
  dplyr::mutate(
    `N cities` = scales::comma(n_cities),
    `Median price` = scales::dollar(round(median_price, 0)),
    `Median rent` = scales::dollar(round(median_rent, 0)),
    `Median cap rate` = paste0(round(median_cap, 2), "%"),
    `Price YoY` = paste0(round(median_price_yoy, 1), "%"),
    `Rent YoY` = paste0(round(median_rent_yoy, 1), "%"),
    `Pop growth` = paste0(round(median_pop_growth, 1), "%")
  ) %>%
  dplyr::select(`TJ distance` = tj_group, `N cities`, `Median price`, `Median rent`,
                `Median cap rate`, `Price YoY`, `Rent YoY`, `Pop growth`) %>%
  DT::datatable(
    rownames = FALSE,
    options = list(dom = "t", ordering = FALSE),
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold; font-size: 1.1em;",
      "Market metrics by distance to nearest Trader Joe's"
    )
  )
```

## Unicorn Leaderboard: Cash Flow + Trader Joe's

The holy grail: a rental market that's **cash-flow positive at 25% down and 7.25% rates AND within 10 miles of a Trader Joe's**. These are rare because TJ tends to pick areas where home prices have already outrun rents. But they exist.

```{r unicorn_table}
unicorns <- markets %>%
  dplyr::filter(tj_unicorn, !is.na(monthly_cash_flow)) %>%
  dplyr::arrange(dplyr::desc(investment_score)) %>%
  dplyr::transmute(
    City = city, State = state, Metro = metro,
    `Home Value` = zhvi,
    `Rent/mo` = zori,
    `Cash flow/mo` = monthly_cash_flow,
    `CoC %` = cash_on_cash_pct,
    `Cap Rate %` = est_cap_rate_pct,
    `TJ mi` = tj_miles,
    `Nearest TJ` = paste0(nearest_tj_city, ", ", nearest_tj_state),
    Score = investment_score,
    Tier = as.character(tier)
  )

cat("**", nrow(unicorns), "unicorn cities** found across ",
    dplyr::n_distinct(unicorns$State), " states, ",
    dplyr::n_distinct(unicorns$Metro), " metros.\n\n", sep = "")

DT::datatable(
  unicorns,
  rownames = FALSE,
  filter = "top",
  options = list(
    pageLength = 25,
    order = list(list(10, "desc")),
    dom = "Bfrtip"
  )
) %>%
  DT::formatCurrency(c("Home Value", "Rent/mo", "Cash flow/mo"), digits = 0) %>%
  DT::formatRound(c("CoC %", "Cap Rate %", "TJ mi"), digits = 2) %>%
  DT::formatStyle(
    "Cash flow/mo",
    color = DT::styleInterval(0, c("#C0392B", "#27AE60")),
    fontWeight = "bold"
  )
```

## Trader Joe's Investment Map

Every US Trader Joe's (purple dots) with the top rental-investment cities overlaid, colored by composite score. Click a marker for the full stats. The map loads live tiles, so it'll only render online.

```{r tj_map, fig.height=7}
tj_stores <- tryCatch(
  readr::read_csv(here::here("data/trader_joes_locations.csv"), show_col_types = FALSE),
  error = function(e) NULL
)

map_cities <- markets %>%
  dplyr::filter(!is.na(tj_miles), !is.na(investment_score),
                !is.na(city_lat), !is.na(city_lon)) %>%
  dplyr::arrange(dplyr::desc(investment_score)) %>%
  dplyr::slice_head(n = 400)

if (!is.null(tj_stores) && nrow(map_cities) > 0) {
  pal <- leaflet::colorNumeric(
    palette = c("#8B0000", "#D4760A", "#2E86C1", "#27AE60"),
    domain = map_cities$investment_score
  )

  city_popup <- sprintf(
    paste0(
      "<b>%s, %s</b><br>",
      "Score: <b>%.1f</b> | Cap rate: %.2f%% | CoC: %.2f%%<br>",
      "Price: $%s | Rent: $%s/mo<br>",
      "Cash flow: <b>$%s/mo</b><br>",
      "Nearest TJ: %s, %s (%.1f mi)"
    ),
    map_cities$city, map_cities$state,
    map_cities$investment_score, map_cities$est_cap_rate_pct,
    tidyr::replace_na(map_cities$cash_on_cash_pct, 0),
    scales::comma(map_cities$zhvi), scales::comma(map_cities$zori),
    scales::comma(round(tidyr::replace_na(map_cities$monthly_cash_flow, 0))),
    map_cities$nearest_tj_city, map_cities$nearest_tj_state, map_cities$tj_miles
  )

  tj_popup <- sprintf(
    "<b>Trader Joe's</b><br>%s<br>%s, %s %s",
    tj_stores$street, tj_stores$city, tj_stores$state, tj_stores$postal_code
  )

  leaflet::leaflet() %>%
    leaflet::addProviderTiles("CartoDB.Positron") %>%
    leaflet::setView(lng = -96, lat = 38.5, zoom = 4) %>%
    leaflet::addCircleMarkers(
      data = tj_stores,
      lng = ~lon, lat = ~lat,
      radius = 3,
      color = "#6A0DAD",
      fillOpacity = 0.5,
      stroke = FALSE,
      popup = tj_popup,
      group = "Trader Joe's stores"
    ) %>%
    leaflet::addCircleMarkers(
      data = map_cities,
      lng = ~city_lon, lat = ~city_lat,
      radius = 6,
      color = ~pal(investment_score),
      fillOpacity = 0.85,
      stroke = TRUE,
      weight = 1,
      popup = city_popup,
      group = "Top investment cities"
    ) %>%
    leaflet::addLegend(
      position = "bottomright",
      pal = pal,
      values = map_cities$investment_score,
      title = "Investment score",
      opacity = 0.9
    ) %>%
    leaflet::addLayersControl(
      overlayGroups = c("Trader Joe's stores", "Top investment cities"),
      options = leaflet::layersControlOptions(collapsed = FALSE)
    )
} else {
  cat("Map data unavailable. Run `Rscript scripts/trader_joes_data.R` to populate.")
}
```

## Gross Yield vs Home Value

```{r scatter_yield, fig.height=8}
top100 <- markets %>% dplyr::slice_head(n = 100)

ggplot2::ggplot(top100, ggplot2::aes(x = zhvi, y = gross_yield_pct, color = tier)) +
  ggplot2::geom_point(size = 3, alpha = 0.8) +
  ggrepel::geom_text_repel(
    ggplot2::aes(label = city_state),
    size = 2.5, max.overlaps = 20, segment.color = "grey70"
  ) +
  ggplot2::scale_x_continuous(labels = scales::label_dollar(scale_cut = scales::cut_short_scale())) +
  ggplot2::scale_color_manual(values = tier_colors) +
  ggplot2::labs(
    title = "Gross Yield vs Home Value (Top 100 Cities)",
    subtitle = "Higher yield + lower price = better cash flow market",
    x = "Zillow Home Value Index", y = "Gross Yield (%)", color = "Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(legend.position = "top")
```

## Leveraged Returns (at 25% Down, 7.25% 30yr)

Cap rate ignores financing. For the leveraged buyer the two numbers that matter are **monthly cash flow** (does the tenant cover the mortgage?) and **cash-on-cash return** (what does my out-of-pocket cash actually earn?). Both are computed at the default assumptions printed above -- use the calculator for your own numbers.

```{r leverage_bar, fig.height=11}
top_cf <- markets %>%
  dplyr::filter(!is.na(monthly_cash_flow)) %>%
  dplyr::arrange(dplyr::desc(monthly_cash_flow)) %>%
  dplyr::slice_head(n = 30)

top_cf %>%
  dplyr::mutate(city_state = forcats::fct_reorder(city_state, monthly_cash_flow)) %>%
  ggplot2::ggplot(ggplot2::aes(x = city_state, y = monthly_cash_flow, fill = tier)) +
  ggplot2::geom_col(width = 0.7) +
  ggplot2::geom_text(
    ggplot2::aes(label = scales::dollar(monthly_cash_flow)),
    hjust = -0.1, size = 3, color = "grey30"
  ) +
  ggplot2::coord_flip(ylim = c(0, max(top_cf$monthly_cash_flow) * 1.15)) +
  ggplot2::scale_fill_manual(values = tier_colors) +
  ggplot2::scale_y_continuous(labels = scales::label_dollar()) +
  ggplot2::labs(
    title = "Top 30 Cities by Monthly Cash Flow",
    subtitle = "25% down, 7.25% mortgage, 30yr, 35% opex. Rent minus PITI minus opex.",
    x = NULL, y = "Monthly Cash Flow ($)", fill = "Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    legend.position = "top",
    panel.grid.major.y = ggplot2::element_blank()
  )
```

```{r coc_bar, fig.height=11}
top_coc <- markets %>%
  dplyr::filter(!is.na(cash_on_cash_pct), cash_on_cash_pct <= 100) %>%
  dplyr::arrange(dplyr::desc(cash_on_cash_pct)) %>%
  dplyr::slice_head(n = 30)

top_coc %>%
  dplyr::mutate(city_state = forcats::fct_reorder(city_state, cash_on_cash_pct)) %>%
  ggplot2::ggplot(ggplot2::aes(x = city_state, y = cash_on_cash_pct, fill = tier)) +
  ggplot2::geom_col(width = 0.7) +
  ggplot2::geom_text(
    ggplot2::aes(label = paste0(cash_on_cash_pct, "%")),
    hjust = -0.1, size = 3, color = "grey30"
  ) +
  ggplot2::coord_flip(ylim = c(0, max(top_coc$cash_on_cash_pct) * 1.15)) +
  ggplot2::scale_fill_manual(values = tier_colors) +
  ggplot2::labs(
    title = "Top 30 Cities by Cash-on-Cash Return",
    subtitle = "Annual cash flow divided by initial out-of-pocket (down payment + closing).",
    x = NULL, y = "Cash-on-Cash Return (%)", fill = "Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    legend.position = "top",
    panel.grid.major.y = ggplot2::element_blank()
  )
```

## Demand Health Check: Cap Rate vs Population Growth

High cap rate means nothing if the town is emptying out. This chart cross-references cap rate with population growth -- the sweet spot is the upper-right quadrant: strong cash flow **and** growing population.

```{r demand_scatter, fig.height=9, fig.width=10}
demand_data <- markets %>%
  dplyr::filter(
    !is.na(pop_growth_pct), !is.na(est_cap_rate_pct), !is.na(pop_2024),
    pop_2024 >= 10000
  )

ggplot2::ggplot(
  demand_data,
  ggplot2::aes(x = pop_growth_pct, y = est_cap_rate_pct, color = tier, size = pop_2024)
) +
  ggplot2::geom_point(alpha = 0.6) +
  ggrepel::geom_text_repel(
    data = demand_data %>%
      dplyr::filter(est_cap_rate_pct >= 5 | pop_growth_pct >= 15 | pop_growth_pct <= -10 |
                    (est_cap_rate_pct >= 4 & pop_growth_pct >= 5)),
    ggplot2::aes(label = city_state),
    size = 2.5, max.overlaps = 25, segment.color = "grey70"
  ) +
  ggplot2::geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  ggplot2::geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  ggplot2::annotate("rect", xmin = 0, xmax = Inf, ymin = 3, ymax = Inf,
                    fill = "#27AE60", alpha = 0.05) +
  ggplot2::annotate("text", x = max(demand_data$pop_growth_pct) * 0.7, y = max(demand_data$est_cap_rate_pct) * 0.95,
                    label = "Sweet spot:\ngrowth + cash flow", size = 3.5, color = "#27AE60", fontface = "bold") +
  ggplot2::scale_color_manual(values = tier_colors) +
  ggplot2::scale_size_continuous(range = c(1, 8), labels = scales::label_comma(), guide = "none") +
  ggplot2::labs(
    title = "Cap Rate vs Population Growth (Cities 10K+ pop)",
    subtitle = "Bubble size = city population. Upper-right = cash flow + demand.",
    x = "Population Growth 2020\u20132024 (%)", y = "Estimated Cap Rate (%)", color = "Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(legend.position = "top")
```

## Hidden Gems: Small Towns with Big Cash Flow

A common investor strategy is to buy in smaller markets where price is low but rent holds up. The **Hidden Gem** flag filters for:

- Population 5K--50K
- Estimated cap rate ≥ 5%
- Non-negative population growth
- County vacancy rate ≤ 12%

```{r hidden_gems}
gems <- markets %>%
  dplyr::filter(hidden_gem) %>%
  dplyr::arrange(dplyr::desc(investment_score))

cat("**", nrow(gems), "cities** match the Hidden Gem criteria.\n")
```

<details class="table-accordion" open><summary>Show hidden gems</summary>

```{r hidden_gem_table}
gems %>%
  dplyr::transmute(
    City = city, State = state, Metro = metro,
    Score = investment_score,
    `Home Value` = scales::dollar(zhvi),
    `Rent/Mo` = scales::dollar(zori, accuracy = 1),
    `Cash Flow/Mo` = scales::dollar(monthly_cash_flow),
    `CoC %` = cash_on_cash_pct,
    `Cap Rate` = est_cap_rate_pct,
    Population = scales::comma(pop_2024),
    `Pop Gr %` = pop_growth_pct,
    `Vacancy %` = vacancy_rate_pct,
    Tier = tier
  ) %>%
  DT::datatable(
    filter = "top", extensions = "Buttons",
    options = list(
      pageLength = 500, dom = "Bfrtip",
      buttons = c("copy", "csv", "excel"),
      scrollY = "400px", scrollX = TRUE,
      order = list(list(4, "desc"))
    )
  )
```

</details>

## Investor Rules of Thumb

Two rent-to-price benchmarks investors use as a quick smell test:

- **1% rule**: Monthly rent ≥ 1% of purchase price. A pass usually means cash flow is plausible.
- **2% rule**: Monthly rent ≥ 2% of purchase price. Rare; typically only in distressed or low-priced markets.

```{r rules_summary, fig.height=4.5, fig.width=9}
rules_data <- markets %>%
  dplyr::summarise(
    `1% rule` = mean(one_pct_rule, na.rm = TRUE),
    `2% rule` = mean(two_pct_rule, na.rm = TRUE),
    `Positive cash flow` = mean(monthly_cash_flow > 0, na.rm = TRUE),
    `Cap rate >= 6%` = mean(est_cap_rate_pct >= 6, na.rm = TRUE)
  ) %>%
  tidyr::pivot_longer(dplyr::everything(), names_to = "rule", values_to = "pct") %>%
  dplyr::mutate(pct = 100 * pct)

ggplot2::ggplot(rules_data, ggplot2::aes(x = reorder(rule, pct), y = pct, fill = pct)) +
  ggplot2::geom_col(width = 0.7) +
  ggplot2::geom_text(
    ggplot2::aes(label = paste0(round(pct, 1), "%")),
    hjust = -0.15, size = 4.2, color = "grey25"
  ) +
  ggplot2::coord_flip(ylim = c(0, 100)) +
  ggplot2::scale_fill_gradient(low = "#D4760A", high = "#27AE60", guide = "none") +
  ggplot2::labs(
    title = "What % of US Cities Clear Each Bar?",
    subtitle = paste0("Out of ", nrow(markets), " cities with Zillow data."),
    x = NULL, y = "Cities meeting threshold (%)"
  ) +
  ggplot2::theme_minimal(base_size = 12)
```

## Vacancy Rate vs Cap Rate

Markets with high vacancy rates are risky -- your rental sits empty longer. Cities in the lower-right have strong cap rates **and** low vacancy.

```{r vacancy_scatter, fig.height=8}
vacancy_data <- markets %>%
  dplyr::filter(!is.na(vacancy_rate_pct), !is.na(est_cap_rate_pct), !is.na(pop_2024),
                pop_2024 >= 10000)

ggplot2::ggplot(
  vacancy_data,
  ggplot2::aes(x = est_cap_rate_pct, y = vacancy_rate_pct, color = tier, size = pop_2024)
) +
  ggplot2::geom_point(alpha = 0.6) +
  ggrepel::geom_text_repel(
    data = vacancy_data %>%
      dplyr::filter(est_cap_rate_pct >= 5 | vacancy_rate_pct >= 20 |
                    (est_cap_rate_pct >= 4 & vacancy_rate_pct <= 6)),
    ggplot2::aes(label = city_state),
    size = 2.5, max.overlaps = 20, segment.color = "grey70"
  ) +
  ggplot2::geom_hline(yintercept = 10, linetype = "dashed", color = "#E74C3C", linewidth = 0.5) +
  ggplot2::annotate("text", x = max(vacancy_data$est_cap_rate_pct) * 0.9,
                    y = 10.5, label = "10% vacancy threshold", size = 3, color = "#E74C3C") +
  ggplot2::scale_color_manual(values = tier_colors) +
  ggplot2::scale_size_continuous(range = c(1, 8), labels = scales::label_comma(), guide = "none") +
  ggplot2::labs(
    title = "Vacancy Rate vs Cap Rate (Cities 10K+ pop)",
    subtitle = "Lower-right = strong returns + low vacancy. Above the red line = risky.",
    x = "Estimated Cap Rate (%)", y = "County Vacancy Rate (%)", color = "Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(legend.position = "top")
```

## Full US Market Explorer

Search, filter, and sort all `r nrow(markets)` cities.

<details class="table-accordion"><summary>Show full table</summary>

```{r full_table}
markets %>%
  dplyr::transmute(
    City = city, State = state, Metro = metro,
    Score = investment_score,
    `Home Value` = scales::dollar(zhvi),
    `Rent/Mo` = scales::dollar(zori, accuracy = 1),
    `Cash Flow/Mo` = scales::dollar(monthly_cash_flow),
    `CoC %` = cash_on_cash_pct,
    `Cap Rate` = est_cap_rate_pct,
    `P/R Ratio` = price_to_rent,
    `Gross Yield` = gross_yield_pct,
    `Price YoY %` = zhvi_yoy_pct,
    `Rent YoY %` = zori_yoy_pct,
    Population = dplyr::if_else(is.na(pop_2024), NA_real_, pop_2024),
    `Pop Gr %` = pop_growth_pct,
    `County Pop Gr %` = county_pop_growth_pct,
    `Net Mig/1K` = net_mig_rate_per1k,
    `Vacancy %` = vacancy_rate_pct,
    `Rental Vac %` = rental_vacancy_pct,
    `Renter %` = renter_pct,
    Tier = tier
  ) %>%
  DT::datatable(
    filter = "top", extensions = "Buttons",
    options = list(
      pageLength = 500, dom = "Bfrtip",
      buttons = c("copy", "csv", "excel"),
      scrollY = "400px", scrollX = TRUE,
      order = list(list(4, "desc"))
    )
  )
```

</details>

## Composite Investment Score

A rank-based composite (0--100) that blends the signals investors actually care about: cap rate (35%), population growth (20%), net migration (15%), low vacancy (15%), and YoY price appreciation (15%). Each factor is converted to a percentile so outliers can't dominate.

```{r score_dist, fig.height=4.5}
ggplot2::ggplot(markets, ggplot2::aes(x = investment_score, fill = tier)) +
  ggplot2::geom_histogram(binwidth = 2, color = "white", linewidth = 0.2) +
  ggplot2::scale_fill_manual(values = tier_colors) +
  ggplot2::labs(
    title = "Distribution of Composite Scores",
    subtitle = "Higher is better. Most cities cluster in the middle.",
    x = "Investment Score", y = "Cities", fill = "Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(legend.position = "top")
```

<details class="table-accordion"><summary>Top 50 by composite score</summary>

```{r score_table}
markets %>%
  dplyr::arrange(dplyr::desc(investment_score)) %>%
  dplyr::slice_head(n = 50) %>%
  dplyr::transmute(
    `#` = dplyr::row_number(),
    City = city, State = state, Metro = metro,
    Score = investment_score,
    `Cap Rate` = paste0(est_cap_rate_pct, "%"),
    `Cash/Mo` = scales::dollar(monthly_cash_flow),
    `CoC %` = paste0(cash_on_cash_pct, "%"),
    `Home Value` = scales::dollar(zhvi),
    `Pop Gr %` = pop_growth_pct,
    `Price YoY` = paste0(zhvi_yoy_pct, "%"),
    `Vacancy %` = vacancy_rate_pct,
    Tier = tier
  ) %>%
  DT::datatable(
    filter = "top", extensions = "Buttons",
    options = list(
      pageLength = 50, dom = "Bfrtip",
      buttons = c("copy", "csv", "excel"),
      scrollY = "400px", scrollX = TRUE
    )
  )
```

</details>

## State Leaderboard

Best cash-flow city per state (highest monthly cash flow at default assumptions), plus state-level medians.

```{r state_leaderboard}
best_per_state <- markets %>%
  dplyr::filter(!is.na(monthly_cash_flow)) %>%
  dplyr::group_by(state) %>%
  dplyr::slice_max(monthly_cash_flow, n = 1, with_ties = FALSE) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(dplyr::desc(monthly_cash_flow))

state_summary <- markets %>%
  dplyr::group_by(state) %>%
  dplyr::summarise(
    n_cities = dplyr::n(),
    med_cap = round(median(est_cap_rate_pct, na.rm = TRUE), 2),
    med_cf = round(median(monthly_cash_flow, na.rm = TRUE), 0),
    med_coc = round(median(cash_on_cash_pct, na.rm = TRUE), 2),
    pct_positive = round(100 * mean(monthly_cash_flow > 0, na.rm = TRUE), 1),
    med_score = round(median(investment_score, na.rm = TRUE), 1),
    .groups = "drop"
  )

best_per_state %>%
  dplyr::left_join(state_summary, by = "state") %>%
  dplyr::transmute(
    State = state,
    `Cities in Data` = n_cities,
    `Best City` = city,
    `Best City Cash/Mo` = scales::dollar(monthly_cash_flow),
    `Best City CoC` = paste0(cash_on_cash_pct, "%"),
    `Best City Cap` = paste0(est_cap_rate_pct, "%"),
    `State Med Cap` = paste0(med_cap, "%"),
    `State Med Cash/Mo` = scales::dollar(med_cf),
    `% Cash Flow+` = paste0(pct_positive, "%"),
    `State Med Score` = med_score
  ) %>%
  DT::datatable(
    filter = "top", extensions = "Buttons",
    options = list(
      pageLength = 50, dom = "Bfrtip",
      buttons = c("copy", "csv", "excel"),
      scrollY = "400px", scrollX = TRUE,
      order = list(list(7, "desc"))
    )
  )
```

```{r state_boxplot, fig.height=9}
# Show states where rent covers debt service worst -> best
state_order <- markets %>%
  dplyr::filter(!is.na(monthly_cash_flow)) %>%
  dplyr::group_by(state) %>%
  dplyr::summarise(med_cf = median(monthly_cash_flow, na.rm = TRUE), n = dplyr::n()) %>%
  dplyr::filter(n >= 5) %>%
  dplyr::arrange(med_cf)

box_data <- markets %>%
  dplyr::filter(state %in% state_order$state, !is.na(monthly_cash_flow)) %>%
  dplyr::mutate(state = factor(state, levels = state_order$state))

ggplot2::ggplot(box_data, ggplot2::aes(x = state, y = monthly_cash_flow)) +
  ggplot2::geom_hline(yintercept = 0, linetype = "dashed", color = "#C0392B", linewidth = 0.6) +
  ggplot2::geom_boxplot(fill = "#2E86C1", color = "#1B4F72", alpha = 0.6, outlier.size = 0.6) +
  ggplot2::coord_flip() +
  ggplot2::scale_y_continuous(labels = scales::label_dollar(), limits = c(-3000, 3000)) +
  ggplot2::labs(
    title = "Monthly Cash Flow Distribution by State (states with 5+ cities)",
    subtitle = "Red line = $0 / mo (break-even). States below = appreciation plays only.",
    x = NULL, y = "Monthly Cash Flow ($)"
  ) +
  ggplot2::theme_minimal(base_size = 11)
```

## Yield Minus Mortgage Rate Spread

Gross rental yield minus the current mortgage rate (7.25%). Positive spread means the rent covers debt service with room to spare; negative means you're subsidizing the mortgage and relying on appreciation.

```{r spread_chart, fig.height=6}
spread_data <- markets %>%
  dplyr::filter(!is.na(yield_spread_pct), !is.na(pop_2024), pop_2024 >= 20000)

ggplot2::ggplot(spread_data, ggplot2::aes(x = yield_spread_pct, fill = tier)) +
  ggplot2::geom_histogram(binwidth = 0.5, color = "white", linewidth = 0.2) +
  ggplot2::geom_vline(xintercept = 0, linetype = "dashed", color = "grey30") +
  ggplot2::scale_fill_manual(values = tier_colors) +
  ggplot2::labs(
    title = "Gross Yield Minus Mortgage Rate",
    subtitle = paste0(
      "Cities with 20K+ population. ",
      round(100 * mean(spread_data$yield_spread_pct > 0, na.rm = TRUE), 1),
      "% have positive spread vs ",
      DEFAULT_MORTGAGE_RATE,
      "% mortgage."
    ),
    x = "Yield - Mortgage Rate (percentage points)", y = "Cities", fill = "Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(legend.position = "top")
```

## Price vs Rent Growth (YoY)

Home value growth and rent growth tell different stories. When home prices outrun rents, cap rates compress (bad for investors); when rents outrun prices, cap rates expand.

```{r yoy_scatter, fig.height=7}
yoy_data <- markets %>%
  dplyr::filter(!is.na(zhvi_yoy_pct), !is.na(zori_yoy_pct), !is.na(pop_2024),
                pop_2024 >= 20000,
                abs(zori_yoy_pct) <= 30, abs(zhvi_yoy_pct) <= 30)

ggplot2::ggplot(
  yoy_data,
  ggplot2::aes(x = zhvi_yoy_pct, y = zori_yoy_pct, color = tier, size = pop_2024)
) +
  ggplot2::geom_point(alpha = 0.6) +
  ggplot2::geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey40") +
  ggrepel::geom_text_repel(
    data = yoy_data %>%
      dplyr::filter(abs(zori_yoy_pct) >= 8 | abs(zhvi_yoy_pct) >= 12 |
                    (zori_yoy_pct >= 5 & zhvi_yoy_pct <= 0)),
    ggplot2::aes(label = city_state),
    size = 2.6, max.overlaps = 25, segment.color = "grey70"
  ) +
  ggplot2::scale_color_manual(values = tier_colors) +
  ggplot2::scale_size_continuous(range = c(1, 8), labels = scales::label_comma(), guide = "none") +
  ggplot2::labs(
    title = "Home Price Growth vs Rent Growth (YoY)",
    subtitle = "Above diagonal = rent growing faster than price (cap rate expanding). Cities 20K+ pop.",
    x = "ZHVI YoY (%)", y = "ZORI YoY (%)", color = "Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(legend.position = "top")
```

## Per-City Detail Lookup

Search for any city to see the full investor profile in one place.

<details class="table-accordion"><summary>Show full detail table</summary>

```{r detail_table}
markets %>%
  dplyr::arrange(dplyr::desc(investment_score)) %>%
  dplyr::transmute(
    City = city, State = state, Metro = metro, County = county,
    Score = investment_score,
    Tier = tier,
    `Home Value` = scales::dollar(zhvi),
    `Price YoY %` = zhvi_yoy_pct,
    `Rent/Mo` = scales::dollar(zori, accuracy = 1),
    `Rent YoY %` = zori_yoy_pct,
    `Cap Rate %` = est_cap_rate_pct,
    `P/R Ratio` = price_to_rent,
    `Mortgage/Mo` = scales::dollar(monthly_mortgage, accuracy = 1),
    `Cash Flow/Mo` = scales::dollar(monthly_cash_flow),
    `Cash-on-Cash %` = cash_on_cash_pct,
    `Break-even Rent` = scales::dollar(break_even_rent, accuracy = 1),
    `Yield Spread %` = yield_spread_pct,
    `1% Rule` = dplyr::if_else(one_pct_rule, "Yes", "No"),
    `2% Rule` = dplyr::if_else(two_pct_rule, "Yes", "No"),
    `Hidden Gem` = dplyr::if_else(hidden_gem, "Yes", "No"),
    Population = dplyr::if_else(is.na(pop_2024), NA_real_, as.numeric(pop_2024)),
    `Pop Gr %` = pop_growth_pct,
    `County Pop Gr %` = county_pop_growth_pct,
    `Net Mig/1K` = net_mig_rate_per1k,
    `Vacancy %` = vacancy_rate_pct,
    `Rental Vac %` = rental_vacancy_pct,
    `% Renter` = renter_pct
  ) %>%
  DT::datatable(
    filter = "top", extensions = "Buttons",
    options = list(
      pageLength = 25, dom = "Bfrtip",
      buttons = c("copy", "csv", "excel"),
      scrollY = "500px", scrollX = TRUE
    )
  )
```

</details>

## Tier Distribution

```{r tier_dist, fig.height=5, fig.width=8}
tier_counts <- markets %>%
  dplyr::count(tier) %>%
  dplyr::mutate(pct = round(100 * n / sum(n), 1))

ggplot2::ggplot(tier_counts, ggplot2::aes(x = tier, y = n, fill = tier)) +
  ggplot2::geom_col(width = 0.6) +
  ggplot2::geom_text(
    ggplot2::aes(label = paste0(n, " (", pct, "%)")),
    vjust = -0.3, size = 4
  ) +
  ggplot2::scale_fill_manual(values = tier_colors) +
  ggplot2::scale_y_continuous(expand = ggplot2::expansion(mult = c(0, 0.15))) +
  ggplot2::labs(
    title = "How Many US Cities Fall Into Each Investment Tier?",
    x = NULL, y = "Number of Cities"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(legend.position = "none")
```

## Metro-Level Comparison

Average metrics across cities within each metro area (metros with 5+ cities).

<details class="table-accordion"><summary>Show metro table</summary>

```{r metro_table}
metro_summary <- markets %>%
  dplyr::filter(!is.na(metro)) %>%
  dplyr::group_by(metro) %>%
  dplyr::summarise(
    n_cities = dplyr::n(),
    avg_zhvi = mean(zhvi, na.rm = TRUE),
    avg_zori = mean(zori, na.rm = TRUE),
    avg_cap_rate = mean(est_cap_rate_pct, na.rm = TRUE),
    avg_ptr = mean(price_to_rent, na.rm = TRUE),
    avg_pop_growth = mean(county_pop_growth_pct, na.rm = TRUE),
    avg_vacancy = mean(vacancy_rate_pct, na.rm = TRUE),
    avg_net_mig = mean(net_mig_rate_per1k, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::filter(n_cities >= 5) %>%
  dplyr::arrange(dplyr::desc(avg_cap_rate))

metro_summary %>%
  dplyr::transmute(
    Metro = metro,
    Cities = n_cities,
    `Avg Home Value` = scales::dollar(avg_zhvi),
    `Avg Rent/Mo` = scales::dollar(avg_zori, accuracy = 1),
    `Avg Cap Rate` = paste0(round(avg_cap_rate, 2), "%"),
    `Avg P/R` = round(avg_ptr, 1),
    `County Pop Gr %` = round(avg_pop_growth, 2),
    `Avg Vacancy %` = round(avg_vacancy, 2),
    `Avg Net Mig/1K` = round(avg_net_mig, 2)
  ) %>%
  DT::datatable(
    filter = "top", extensions = "Buttons",
    options = list(
      pageLength = 500, dom = "Bfrtip",
      buttons = c("copy", "csv", "excel"),
      scrollY = "400px", scrollX = TRUE,
      order = list(list(4, "desc"))
    )
  )
```

</details>

---

**How to update this data:** Run `Rscript scripts/us_rental_markets_data.R` to download Zillow ZHVI + ZORI. Then run `Rscript scripts/us_rental_demand_data.R` to merge Census population / vacancy / migration and compute leverage metrics (mortgage payment, cash flow, cash-on-cash, composite score). Then re-render this page.
