---
title: "US Rental Investment Markets"
editor_options:
  chunk_output_type: console
output:
  html_document:
    includes:
      in_header: header/header.html
      after_body: include_footer.html
---

```{r pipe, include=FALSE}
`%>%` <- magrittr::`%>%`
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.width = 10, fig.height = 8
)
```

```{r data}
raw <- readr::read_csv(here::here("data", "us_rental_markets.csv"), show_col_types = FALSE)

# Filter out noise: seasonal vacation rentals and extremely low-value markets
markets <- raw %>%
  dplyr::filter(
    zhvi >= 50000,
    zori <= 5000,
    zori >= 500,
    gross_yield_pct <= 20
  ) %>%
  dplyr::mutate(
    tier = factor(tier, levels = c("Strong Buy", "Moderate", "Lean Rent", "Rent Favored")),
    city_state = paste0(city, ", ", state)
  )

zhvi_date <- unique(raw$zhvi_date)[1]
zori_date <- unique(raw$zori_date)[1]

tier_colors <- c(
  "Strong Buy"   = "#27AE60",
  "Moderate"     = "#2E86C1",
  "Lean Rent"    = "#D4760A",
  "Rent Favored" = "#8B0000"
)
```

Where should you buy rental property in the US? This page ranks **`r nrow(markets)` cities** by investment metrics derived from Zillow's public home value and rent indices. The price-to-rent ratio is the standard benchmark:

- **Strong Buy** (P/R < 15): Cash flow positive from day one
- **Moderate** (P/R 15--20): Reasonable returns, some appreciation upside
- **Lean Rent** (P/R 20--25): Marginal; depends on appreciation
- **Rent Favored** (P/R > 25): Buy only if you're betting on price growth

### Data Sources

All data is traceable and reproducible. Run `Rscript scripts/us_rental_markets_data.R` to refresh.

```{r sources_table}
tibble::tribble(
  ~Source, ~Description, ~Updated,
  "Zillow ZHVI", "Typical home values (city level, SFR+condo, 35th\u201365th percentile)", as.character(zhvi_date),
  "Zillow ZORI", "Typical observed rents (city level, SFR+condo+MFR)", as.character(zori_date),
  "Census Pop Estimates", "City + county population (2020\u20132024), net migration", "V2024",
  "Census ACS 5-Year", "Vacancy rates, renter %, housing counts (county level)", "2019\u20132023"
) %>%
  gt::gt() %>%
  gt::tab_header(
    title = gt::md("**Data Sources**"),
    subtitle = "All public, no API key required"
  ) %>%
  gt::fmt_markdown(columns = Source) %>%
  gt::cols_width(
    Source ~ gt::px(180),
    Description ~ gt::px(400),
    Updated ~ gt::px(130)
  ) %>%
  gt::cols_align(align = "left") %>%
  gt::tab_source_note(
    gt::md("Cap rate = gross yield \u00d7 (1 \u2212 35% expense ratio). Raw data: [us_rental_markets.csv](data/us_rental_markets.csv)")
  ) %>%
  gt::tab_style(
    style = gt::cell_fill(color = "#f8f9fa"),
    locations = gt::cells_body(rows = seq(1, 4, by = 2))
  ) %>%
  gt::tab_options(
    table.font.size = gt::px(14),
    heading.title.font.size = gt::px(16),
    heading.subtitle.font.size = gt::px(13),
    column_labels.font.weight = "bold",
    column_labels.background.color = "#2C3E50",
    table.border.top.color = "#2C3E50",
    table.border.bottom.color = "#2C3E50",
    source_notes.font.size = gt::px(12)
  ) %>%
  gt::tab_style(
    style = gt::cell_text(color = "white"),
    locations = gt::cells_column_labels()
  )
```

## Top 40 Cities by Cap Rate {.tabset}

```{r top_markets}
top40 <- markets %>%
  dplyr::slice_head(n = 40)
```

### Bar Chart

```{r cap_rate_bar, fig.height=12}
top40 %>%
  dplyr::mutate(city_state = forcats::fct_reorder(city_state, est_cap_rate_pct)) %>%
  ggplot2::ggplot(ggplot2::aes(x = city_state, y = est_cap_rate_pct, fill = tier)) +
  ggplot2::geom_col(width = 0.7) +
  ggplot2::geom_text(
    ggplot2::aes(label = paste0(est_cap_rate_pct, "%")),
    hjust = -0.1, size = 3, color = "grey30"
  ) +
  ggplot2::coord_flip(ylim = c(0, max(top40$est_cap_rate_pct) * 1.15)) +
  ggplot2::scale_fill_manual(values = tier_colors) +
  ggplot2::labs(
    title = "Estimated Cap Rate (Top 40 Cities)",
    subtitle = "Gross yield minus 35% expense ratio",
    x = NULL, y = "Estimated Cap Rate (%)", fill = "Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    legend.position = "top",
    panel.grid.major.y = ggplot2::element_blank()
  )
```

### Interactive Table

<details><summary>Show table</summary>

```{r top_table}
top40 %>%
  dplyr::transmute(
    `#` = dplyr::row_number(),
    City = city, State = state, Metro = metro,
    `Home Value` = scales::dollar(zhvi),
    `Rent/Mo` = scales::dollar(zori, accuracy = 1),
    `Est Cap Rate` = paste0(est_cap_rate_pct, "%"),
    `P/R Ratio` = price_to_rent,
    Population = dplyr::if_else(is.na(pop_2024), "---", scales::comma(pop_2024)),
    `Pop Growth` = dplyr::if_else(is.na(pop_growth_pct), "---", paste0(pop_growth_pct, "%")),
    `Vacancy` = dplyr::if_else(is.na(vacancy_rate_pct), "---", paste0(vacancy_rate_pct, "%")),
    `Rental Vac` = dplyr::if_else(is.na(rental_vacancy_pct), "---", paste0(rental_vacancy_pct, "%")),
    `Net Mig/1K` = dplyr::if_else(is.na(net_mig_rate_per1k), "---", as.character(net_mig_rate_per1k)),
    `% Renter` = dplyr::if_else(is.na(renter_pct), "---", paste0(renter_pct, "%")),
    Tier = tier
  ) %>%
  DT::datatable(
    filter = "top", extensions = "Buttons",
    options = list(
      pageLength = 500, dom = "Bfrtip",
      buttons = c("copy", "csv", "excel"),
      scrollX = TRUE
    )
  )
```

</details>

## Gross Yield vs Home Value

```{r scatter_yield, fig.height=8}
top100 <- markets %>% dplyr::slice_head(n = 100)

ggplot2::ggplot(top100, ggplot2::aes(x = zhvi, y = gross_yield_pct, color = tier)) +
  ggplot2::geom_point(size = 3, alpha = 0.8) +
  ggrepel::geom_text_repel(
    ggplot2::aes(label = city_state),
    size = 2.5, max.overlaps = 20, segment.color = "grey70"
  ) +
  ggplot2::scale_x_continuous(labels = scales::label_dollar(scale_cut = scales::cut_short_scale())) +
  ggplot2::scale_color_manual(values = tier_colors) +
  ggplot2::labs(
    title = "Gross Yield vs Home Value (Top 100 Cities)",
    subtitle = "Higher yield + lower price = better cash flow market",
    x = "Zillow Home Value Index", y = "Gross Yield (%)", color = "Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(legend.position = "top")
```

## Demand Health Check: Cap Rate vs Population Growth

High cap rate means nothing if the town is emptying out. This chart cross-references cap rate with population growth -- the sweet spot is the upper-right quadrant: strong cash flow **and** growing population.

```{r demand_scatter, fig.height=9, fig.width=10}
demand_data <- markets %>%
  dplyr::filter(
    !is.na(pop_growth_pct), !is.na(est_cap_rate_pct), !is.na(pop_2024),
    pop_2024 >= 10000
  )

ggplot2::ggplot(
  demand_data,
  ggplot2::aes(x = pop_growth_pct, y = est_cap_rate_pct, color = tier, size = pop_2024)
) +
  ggplot2::geom_point(alpha = 0.6) +
  ggrepel::geom_text_repel(
    data = demand_data %>%
      dplyr::filter(est_cap_rate_pct >= 5 | pop_growth_pct >= 15 | pop_growth_pct <= -10 |
                    (est_cap_rate_pct >= 4 & pop_growth_pct >= 5)),
    ggplot2::aes(label = city_state),
    size = 2.5, max.overlaps = 25, segment.color = "grey70"
  ) +
  ggplot2::geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  ggplot2::geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  ggplot2::annotate("rect", xmin = 0, xmax = Inf, ymin = 3, ymax = Inf,
                    fill = "#27AE60", alpha = 0.05) +
  ggplot2::annotate("text", x = max(demand_data$pop_growth_pct) * 0.7, y = max(demand_data$est_cap_rate_pct) * 0.95,
                    label = "Sweet spot:\ngrowth + cash flow", size = 3.5, color = "#27AE60", fontface = "bold") +
  ggplot2::scale_color_manual(values = tier_colors) +
  ggplot2::scale_size_continuous(range = c(1, 8), labels = scales::label_comma(), guide = "none") +
  ggplot2::labs(
    title = "Cap Rate vs Population Growth (Cities 10K+ pop)",
    subtitle = "Bubble size = city population. Upper-right = cash flow + demand.",
    x = "Population Growth 2020\u20132024 (%)", y = "Estimated Cap Rate (%)", color = "Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(legend.position = "top")
```

## Vacancy Rate vs Cap Rate

Markets with high vacancy rates are risky -- your rental sits empty longer. Cities in the lower-right have strong cap rates **and** low vacancy.

```{r vacancy_scatter, fig.height=8}
vacancy_data <- markets %>%
  dplyr::filter(!is.na(vacancy_rate_pct), !is.na(est_cap_rate_pct), !is.na(pop_2024),
                pop_2024 >= 10000)

ggplot2::ggplot(
  vacancy_data,
  ggplot2::aes(x = est_cap_rate_pct, y = vacancy_rate_pct, color = tier, size = pop_2024)
) +
  ggplot2::geom_point(alpha = 0.6) +
  ggrepel::geom_text_repel(
    data = vacancy_data %>%
      dplyr::filter(est_cap_rate_pct >= 5 | vacancy_rate_pct >= 20 |
                    (est_cap_rate_pct >= 4 & vacancy_rate_pct <= 6)),
    ggplot2::aes(label = city_state),
    size = 2.5, max.overlaps = 20, segment.color = "grey70"
  ) +
  ggplot2::geom_hline(yintercept = 10, linetype = "dashed", color = "#E74C3C", linewidth = 0.5) +
  ggplot2::annotate("text", x = max(vacancy_data$est_cap_rate_pct) * 0.9,
                    y = 10.5, label = "10% vacancy threshold", size = 3, color = "#E74C3C") +
  ggplot2::scale_color_manual(values = tier_colors) +
  ggplot2::scale_size_continuous(range = c(1, 8), labels = scales::label_comma(), guide = "none") +
  ggplot2::labs(
    title = "Vacancy Rate vs Cap Rate (Cities 10K+ pop)",
    subtitle = "Lower-right = strong returns + low vacancy. Above the red line = risky.",
    x = "Estimated Cap Rate (%)", y = "County Vacancy Rate (%)", color = "Tier"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(legend.position = "top")
```

## Full US Market Explorer

Search, filter, and sort all `r nrow(markets)` cities.

<details><summary>Show full table</summary>

```{r full_table}
markets %>%
  dplyr::transmute(
    City = city, State = state, Metro = metro,
    `Home Value` = scales::dollar(zhvi),
    `Rent/Mo` = scales::dollar(zori, accuracy = 1),
    `Cap Rate` = est_cap_rate_pct,
    `P/R Ratio` = price_to_rent,
    `Gross Yield` = gross_yield_pct,
    Population = dplyr::if_else(is.na(pop_2024), NA_real_, pop_2024),
    `Pop Gr %` = pop_growth_pct,
    `County Pop Gr %` = county_pop_growth_pct,
    `Net Mig/1K` = net_mig_rate_per1k,
    `Vacancy %` = vacancy_rate_pct,
    `Rental Vac %` = rental_vacancy_pct,
    `Renter %` = renter_pct,
    Tier = tier
  ) %>%
  DT::datatable(
    filter = "top", extensions = "Buttons",
    options = list(
      pageLength = 500, dom = "Bfrtip",
      buttons = c("copy", "csv", "excel"),
      scrollX = TRUE,
      scrollX = TRUE,
      order = list(list(5, "desc"))
    )
  )
```

</details>

## Tier Distribution

```{r tier_dist, fig.height=5, fig.width=8}
tier_counts <- markets %>%
  dplyr::count(tier) %>%
  dplyr::mutate(pct = round(100 * n / sum(n), 1))

ggplot2::ggplot(tier_counts, ggplot2::aes(x = tier, y = n, fill = tier)) +
  ggplot2::geom_col(width = 0.6) +
  ggplot2::geom_text(
    ggplot2::aes(label = paste0(n, " (", pct, "%)")),
    vjust = -0.3, size = 4
  ) +
  ggplot2::scale_fill_manual(values = tier_colors) +
  ggplot2::scale_y_continuous(expand = ggplot2::expansion(mult = c(0, 0.15))) +
  ggplot2::labs(
    title = "How Many US Cities Fall Into Each Investment Tier?",
    x = NULL, y = "Number of Cities"
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(legend.position = "none")
```

## Metro-Level Comparison

Average metrics across cities within each metro area (metros with 5+ cities).

<details><summary>Show metro table</summary>

```{r metro_table}
metro_summary <- markets %>%
  dplyr::filter(!is.na(metro)) %>%
  dplyr::group_by(metro) %>%
  dplyr::summarise(
    n_cities = dplyr::n(),
    avg_zhvi = mean(zhvi, na.rm = TRUE),
    avg_zori = mean(zori, na.rm = TRUE),
    avg_cap_rate = mean(est_cap_rate_pct, na.rm = TRUE),
    avg_ptr = mean(price_to_rent, na.rm = TRUE),
    avg_pop_growth = mean(county_pop_growth_pct, na.rm = TRUE),
    avg_vacancy = mean(vacancy_rate_pct, na.rm = TRUE),
    avg_net_mig = mean(net_mig_rate_per1k, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::filter(n_cities >= 5) %>%
  dplyr::arrange(dplyr::desc(avg_cap_rate))

metro_summary %>%
  dplyr::transmute(
    Metro = metro,
    Cities = n_cities,
    `Avg Home Value` = scales::dollar(avg_zhvi),
    `Avg Rent/Mo` = scales::dollar(avg_zori, accuracy = 1),
    `Avg Cap Rate` = paste0(round(avg_cap_rate, 2), "%"),
    `Avg P/R` = round(avg_ptr, 1),
    `County Pop Gr %` = round(avg_pop_growth, 2),
    `Avg Vacancy %` = round(avg_vacancy, 2),
    `Avg Net Mig/1K` = round(avg_net_mig, 2)
  ) %>%
  DT::datatable(
    filter = "top", extensions = "Buttons",
    options = list(
      pageLength = 500, dom = "Bfrtip",
      buttons = c("copy", "csv", "excel"),
      scrollX = TRUE,
      scrollX = TRUE,
      order = list(list(4, "desc"))
    )
  )
```

</details>

---

**How to update this data:** Run `Rscript scripts/us_rental_markets_data.R` to download Zillow ZHVI + ZORI. Then run `Rscript scripts/us_rental_demand_data.R` to merge Census population, vacancy, and migration data. Then re-render this page.
