---
title: "When Were Confederate Statues Created?"
editor_options: 
  chunk_output_type: console
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```

# When were Confederate Statues made?


R Project for web scraping Wikipedia looking for dates when confederate statues were erected


```{r prep, include=FALSE}
# Set default chunk output
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  tidy = FALSE, results = "asis"
)

library(tidyverse)
library(plotly)
```

This is so wild. Let this sink in.

This was motivated by my recent trip to South Carolina and seeing my first confederate statue. I am not sure why we glorify these people who fought for people to be enslaved.

Check these out btw:

* https://www.youtube.com/watch?v=dOkFXPblLpU
  - how the United Daughters of the Confederacy brainwashed the youth by pretending slavery wasn't the cause of the Civil War
* https://www.youtube.com/watch?v=WClgR6Q0aPE
  - the reason why each peek of confederate statues were created
* https://www.theguardian.com/us-news/2017/aug/16/confederate-monuments-civil-war-history-trump
  - how each peek and why these statues were created


```{r}
all_dates = readr::read_csv("https://raw.githubusercontent.com/cavandonohoe/confederate_statues/main/data/confederate_statue_dates.csv")

# find mode
max_statues = all_dates %>% group_by(value3) %>%
  count %>%
  ungroup() %>%
  arrange(desc(n)) %>%
  filter(n == max(n))

mode = max_statues %>% 
  pull(value3)

with_max = max_statues %>% 
  pull(n)
```

Also, we witness the most statues being built in a single year in `r mode` with `r with_max` statues.

These bins are of width = 5 years.

```{r graph1, fig.width=10, fig.height=7.5}
all_dates_clean <- all_dates %>% group_by(value3) %>%
  count %>%
  ungroup() %>%
  mutate(date_bins = cut(value3, breaks = seq(min(value3), max(value3)+5, by = 5), right = FALSE),
         text = paste0(value3, ": ", n)) %>%
  group_by(date_bins) %>%
  summarise(n = sum(n),
            text = paste(text, collapse = "\n")) %>%
  ungroup() %>%
  mutate(date_min = as.numeric(gsub("\\[(\\d+),.*", "\\1", date_bins))) %>%
  rename(date = date_min)

confederate_statue_hist = all_dates_clean %>%
  ggplot(aes(x = date, text = text, y = n)) +
  geom_col(fill = "red") +
  scale_x_continuous(breaks = seq(min(all_dates_clean$date), max(all_dates_clean$date), by = 5)) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size=12, angle = 90),
        axis.text.y = element_text(size=15),
        title = element_text(size=15)) +
  ylab("# of Statues") +
  ggtitle("Number of Confederate Statues Over Time")

confederate_statue_hist %>% ggplotly()
```

It's just something to think about...

Also web scraping is fun



sources:  
https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_Alabama
https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_Georgia
https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_Mississippi
https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_North_Carolina
https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials_in_South_Carolina
https://en.wikipedia.org/wiki/List_of_Confederate_monuments_and_memorials



To view web scraping methodology, please view: https://github.com/cavandonohoe/confederate_statues



