---
title: "Covid Cases: Then & Now"
date: '`r format(Sys.time(), "%Y-%m-%d %X")`'
editor_options: 
  chunk_output_type: console
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
    includes:
      in_header: header/header.html
---

```{r pipe, include=FALSE}
` %>%
  ` <- magrittr::` %>%
  `
```

```{r setup, include=FALSE}
# Set default chunk output
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  results = "asis"
)
```

```{r libraries}

```



```{r grab_data}
covid.cases <- readr::read_csv(paste0(
  "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/",
  "csse_covid_19_data/csse_covid_19_time_series/",
  "time_series_covid19_confirmed_global.csv"
))
```

## United States

```{r united_states_data}
covid.cases_v2 <- covid.cases %>%
  
  tidyr::gather(key="Date", value="confirmed", -c(`Country/Region`, `Province/State`, Lat,
    Long)) %>%
    
  dplyr::mutate(Date = as.Date(Date, format="%m/%d/%y"))


covid.cases_v3 <- covid.cases_v2 %>%
  dplyr::filter(grepl(x=`Country/Region`, pattern= "US")) %>%
  
  dplyr::mutate(confirmed.previous.day = dplyr::lag(confirmed)) %>%
    
  dplyr::mutate(New.Cases = confirmed - confirmed.previous.day)

covid.cases_v4 <- covid.cases_v3 %>%
  
  dplyr::mutate(Seven.Day.Average = zoo::rollmean(New.Cases, 7, na.pad = TRUE, align = "right"))


covid.cases_v5 <- covid.cases_v4 %>%
  dplyr::filter(!is.na(Seven.Day.Average)) %>%
  
  dplyr::mutate(Seven.Day.Average_lag = dplyr::lag(Seven.Day.Average)) %>%
    
  dplyr::mutate(Change = Seven.Day.Average - Seven.Day.Average_lag) %>%
    
  dplyr::mutate(Change = ifelse(Change >= 0, paste0("+", round(Change, 0)),
                         paste0(round(Change, 0))))
```

Please note I only have data until `r max(covid.cases_v2$Date)`. The rolling average is the mean of that day's cases as well as the 6 days prior. 

Also for the US total aggregate cases, my data is coming from [https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv](https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv)

Same source as Johns Hopkins, so I hope it's legit.

[Git Repo](https://github.com/CSSEGISandData/COVID-19)

```{r us_plots}
us_cases_plot <- covid.cases_v5 %>%
  ggplot2::ggplot(ggplot2::aes(x=Date, y=Seven.Day.Average)) +
  ggplot2::geom_line() + ggplot2::ggtitle("Seven Day Average Across the US") +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::scale_x_date(date_labels = "%b %y", breaks = scales::pretty_breaks(n=7))+
  ggplot2::geom_hline(yintercept = covid.cases_v5$Seven.Day.Average[covid.cases_v5$Date ==
                                                             max(covid.cases_v5$Date)],
             linetype=2, colour="red")+
  ggplot2::ylab("Seven Day Average") +
  ggplot2::theme(axis.title.x=ggplot2::element_blank())

us_cases_plot %>%
  plotly::ggplotly()

```

The red dashed line helps us visualize where our cases used to be. The following table provides the actual days and seven day average cases of those intersections.

```{r intersection_points_us}
covid.cases_v5 %>%
  
  dplyr::mutate(Intersection = Seven.Day.Average >=
           covid.cases_v5$Seven.Day.Average[covid.cases_v5$Date ==
                                              max(covid.cases_v5$Date, na.rm = TRUE)]) %>%
                                                
  dplyr::mutate(Intersection_lead = dplyr::lead(Intersection)) %>%
    
  dplyr::filter(Intersection_lead != Intersection | Date == max(Date)) %>%
    
  dplyr::select(-Lat, -Long) %>%
    
  dplyr::select(Date, seven_day_average = Seven.Day.Average) %>%
    
  dplyr::mutate(seven_day_average = round(seven_day_average)) %>%
    
  DT::datatable(filter = "top", extensions = "Buttons",
            options = list(paging = FALSE, dom = "Bfrtip", buttons = c("copy", "csv", "excel"),
                           scrollY = "400px", scrollX = TRUE))

```

## California

I like to check on my state as well.

When parsing for states, my data is coming from here: [https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv](https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv)

Same as NY Times.

[Git Repo](https://github.com/nytimes/covid-19-data)

```{r  grab_data_california}
covid.cases_state <- readr::read_csv(paste0(
  "https://raw.githubusercontent.com/nytimes/covid-19-data/master/",
  "us-states.csv"
))

```

```{r clean_ca_data}
covid.cases_state_v2 <- covid.cases_state %>%
  dplyr::select(date, state, cases) %>%
  dplyr::arrange(date) %>%
  
  dplyr::rename(Date = date, confirmed = cases, NAME = state) %>%
    
  dplyr::left_join(tibble::tibble(NAME = state.name, state = state.abb))


covid.cases_state_v3 <- covid.cases_state_v2 %>%
  dplyr::filter(state == "CA") %>%
  
  dplyr::mutate(confirmed = tidyr::replace_na(confirmed, 0)) %>%
    
  dplyr::mutate(confirmed.previous.day = dplyr::lag(confirmed)) %>%
    
  dplyr::mutate(New.Cases = confirmed - confirmed.previous.day)

covid.cases_state_v4 <- covid.cases_state_v3 %>%
  
  dplyr::mutate(Seven.Day.Average = zoo::rollmean(New.Cases, 7, na.pad = TRUE, align = "right"))


covid.cases_state_v5 <- covid.cases_state_v4 %>%
  dplyr::filter(!is.na(Seven.Day.Average)) %>%
  
  dplyr::mutate(Seven.Day.Average_lag = dplyr::lag(Seven.Day.Average)) %>%
    
  dplyr::mutate(Change = Seven.Day.Average - Seven.Day.Average_lag) %>%
    
  dplyr::mutate(Change_format = ifelse(Change >= 0, paste0("+", round(Change, 0)),
                         paste0(round(Change, 0))))
```


```{r ca_plot}
ca_plot <- covid.cases_state_v5 %>%
  ggplot2::ggplot(ggplot2::aes(x=Date, y=Seven.Day.Average, label=Change_format)) +
    ggplot2::geom_line() +
  ggplot2::ggtitle(paste("New", unique(covid.cases_state_v5$NAME), "Covid Cases")) +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::scale_x_date(breaks = scales::pretty_breaks(7), date_labels = "%b %y")+
  ggplot2::geom_hline(
    yintercept = covid.cases_state_v5$Seven.Day.Average[
      covid.cases_state_v5$Date == max(covid.cases_state_v5$Date)
    ],
    linetype = 2,
    colour = "red"
  ) +
  ggplot2::ylab("Seven Day Average")

ca_plot %>%
  plotly::ggplotly()
```

Just like the US chart, the red dashed line can help us visualize the last times we were at today's COVID cases. Below is a table of the exact dates of intersection between today's cases vs previous cases. 

```{r ca_intersections}
covid.cases_state_v5 %>%
  
  dplyr::mutate(Intersection = Seven.Day.Average >=
           covid.cases_state_v5$Seven.Day.Average[covid.cases_state_v5$Date ==
                                              max(covid.cases_state_v5$Date, na.rm = TRUE)]) %>%
                                                
  dplyr::mutate(Intersection_lead = dplyr::lead(Intersection)) %>%
    
  dplyr::filter(Intersection_lead != Intersection | Date == max(Date)) %>%
    
  dplyr::select(Date, seven_day_average = Seven.Day.Average) %>%
    
  dplyr::mutate(seven_day_average = round(seven_day_average)) %>%
    
  DT::datatable(filter = "top", extensions = "Buttons",
            options = list(paging = FALSE, dom = "Bfrtip", buttons = c("copy", "csv", "excel"),
                           scrollY = "400px", scrollX = TRUE))

```

Also, the way the intersection table works is if today's 7 Day Average Cases = 10000, and we had 9,982 cases on August 2nd, 2021 and 10,190 cases on August 3rd, 2021, then August 2nd, 2021 would be listed in the table.

## State vs State

State populations are coming from here: [http://www2.census.gov/programs-surveys/popest/datasets/2010-2019/national/totals/nst-est2019-alldata.csv](http://www2.census.gov/programs-surveys/popest/datasets/2010-2019/national/totals/nst-est2019-alldata.csv)

```{r grab_statewide_data}
# luckily we already have this in the script
# covid.cases_state <- readr::read_csv(paste0(
#   "https://raw.githubusercontent.com/nytimes/covid-19-data/master/",
#   "us-states.csv"
# ))

# need to normalize everything to the same standard
state_pop <- readr::read_csv(paste0(
  "http://www2.census.gov/programs-surveys/popest/datasets/2010-2019/",
  "national/totals/nst-est2019-alldata.csv"
))

```

```{r clean_statewide_data}
state_pop_v2 <- state_pop %>%
  dplyr::select(NAME, POPESTIMATE2019) %>%
  
  dplyr::left_join(tibble::tibble(NAME = state.name, state = state.abb)) %>%
    
  dplyr::rename(population = POPESTIMATE2019)

covid.cases_state_v2 <- covid.cases_state %>%
  
  dplyr::select(date, state, cases, deaths) %>%
    dplyr::arrange(date) %>%
    
  dplyr::rename(Date = date, confirmed = cases, NAME = state) %>%
    dplyr::left_join(state_pop_v2) %>%
    
  dplyr::select(Date, state, confirmed, population, deaths) %>%
    
  dplyr::arrange(Date) %>%
    
  dplyr::mutate(confirmed = confirmed / population,
         death_per_cap = deaths / population) %>%
           
  dplyr::filter(state %in% state.abb)


covid.cases_state_v3 <- covid.cases_state_v2 %>%
  dplyr::group_by(state) %>%
  
  dplyr::mutate(confirmed = tidyr::replace_na(confirmed, 0),
         death_per_cap = tidyr::replace_na(death_per_cap, 0)) %>%
           
  dplyr::mutate(confirmed.previous.day = dplyr::lag(confirmed),
         death_per_cap_previous_day = dplyr::lag(death_per_cap)) %>%
           
  dplyr::mutate(New.Cases = confirmed - confirmed.previous.day,
         new_deaths = death_per_cap - death_per_cap_previous_day)

covid.cases_state_v4 <- covid.cases_state_v3 %>%
  
  dplyr::mutate(Seven.Day.Average = zoo::rollmean(New.Cases, 7, na.pad = TRUE, align = "right"),
         seven_day_average_deaths = zoo::rollmean(new_deaths, 7, na.pad = TRUE,
           align = "right")) %>%
           
  dplyr::ungroup()


covid.cases_state_v5 <- covid.cases_state_v4 %>%
  
  dplyr::filter(!is.na(Seven.Day.Average))


```


```{r state_by_state_plot}
covid.cases_state_plot <- covid.cases_state_v5 %>%
  ggplot2::ggplot(ggplot2::aes(x=Date, y=Seven.Day.Average, colour = state)) +
    ggplot2::geom_line() +
  ggplot2::ggtitle("New Covid Cases per Capita") +
  ggplot2::scale_y_continuous(labels = scales::comma) +
    ggplot2::scale_x_date(breaks = scales::pretty_breaks(7), date_labels = "%b %y") +
  ggplot2::ylab("Seven Day Average per Capita")

covid.cases_state_plot %>%
  plotly::ggplotly()

```

Yes, this is noisy, but you have the option to double click any number of states on the right to isolate. 

## Vaccines

Now, I'm no public health expert, but it does seem like the higher vaccine rate a state has, the lower the COVID cases are. Let's just let the data talk.

Vaccination data from here: [https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/us_state_vaccinations.csv](https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/us_state_vaccinations.csv)

```{r vaccine_data}
vaccines <- readr::read_csv(paste0(
  "https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/",
  "vaccinations/us_state_vaccinations.csv"
))

state_tib <- tibble::tibble(location = state.name, state = state.abb)

vaccines_current_rate <- vaccines %>%
  
  dplyr::mutate(location = ifelse(location == "New York State", "New York", location)) %>%
    
  dplyr::filter(date == max(date)) %>%
    
  dplyr::filter(location %in% state.name) %>%
    
  dplyr::select(location, people_fully_vaccinated_per_hundred) %>%
    
  dplyr::arrange(dplyr::desc(people_fully_vaccinated_per_hundred))

covid_current <- covid.cases_state_v5 %>%
  dplyr::ungroup() %>%
  dplyr::filter(Date == max(Date, na.rm = TRUE)) %>%
  
  dplyr::arrange(Seven.Day.Average) %>%
    
  dplyr::mutate(new_cases_actual = New.Cases * population) %>%
    
  dplyr::select(state, Seven.Day.Average) %>%
    
  dplyr::left_join(state_tib)

covid_cases_w_vaccines <- covid_current %>%
  dplyr::full_join(vaccines_current_rate)

vaccines_historic_rate <- vaccines %>%
  
  dplyr::mutate(location = ifelse(location == "New York State", "New York", location)) %>%
    
  dplyr::filter(location %in% state.name) %>%
    
  dplyr::select(date, location, people_fully_vaccinated_per_hundred, people_fully_vaccinated) %>%
    
  dplyr::arrange(dplyr::desc(people_fully_vaccinated_per_hundred))

```



```{r vaccines_by_state}

vaccines_historic_rate_plot <- vaccines_historic_rate %>%
  
  ggplot2::ggplot(ggplot2::aes(x=date, y= people_fully_vaccinated_per_hundred, colour = location)) +
  ggplot2::geom_point() +
  ggplot2::scale_x_date(date_labels = "%h %d", breaks = scales::pretty_breaks(7)) +
  ggplot2::ylab("% fully vaxxed") +
  ggplot2::ggtitle("vax rate per state")

vaccines_historic_rate_plot %>%
  plotly::ggplotly()

```

Now if that is a little too noisy, I'll break it down between red and blue.

```{r vaccine_party_plot}
usa_pop <- state_pop_v2 %>%
  dplyr::filter(NAME == "United States") %>%
  dplyr::pull(population)

vaccines_usa <- vaccines %>%
  
  dplyr::mutate(location = ifelse(location == "New York State", "New York", location)) %>%
    
  dplyr::filter(location %in% state.name) %>%
    
  dplyr::group_by(date) %>%
    
  dplyr::summarise(people_fully_vaccinated = sum(people_fully_vaccinated, na.rm = TRUE)) %>%
    
  dplyr::ungroup() %>%
    dplyr::mutate(per_fully_vaxxed = people_fully_vaccinated / usa_pop)

# voting preference
voting_pref <- readr::read_csv(here::here("data", "president_1976_2020.csv"))

voting_pref2 <- voting_pref %>%
  dplyr::filter(year == 2020 & party_detailed %in% c("DEMOCRAT", "REPUBLICAN")) %>%
  
  dplyr::mutate(per_votes = candidatevotes / totalvotes) %>%
    
  dplyr::group_by(state) %>%
    
  dplyr::filter(candidatevotes == max(candidatevotes)) %>%
    
  dplyr::ungroup() %>%
    
  dplyr::select(state, state_po, party_detailed, per_votes) %>%
    
  dplyr::mutate(party_simp = ifelse(party_detailed == "REPUBLICAN", "R", "D"),
         party_color = ifelse(party_simp == "R", "red", "blue"),
         STATE = state,
         location = snakecase::to_any_case(string = tolower(STATE), case = "title"),
         per_votes_sign = ifelse(party_simp == "D", per_votes, -per_votes))

vaccine_party <- vaccines_historic_rate %>%
  dplyr::left_join(voting_pref2)


vaccine_party_plot <- vaccine_party %>%
  
  ggplot2::ggplot(ggplot2::aes(x=date, y=people_fully_vaccinated_per_hundred,
    colour = per_votes_sign,
             label = location)) +
  ggplot2::geom_point() +
  ggplot2::scale_colour_gradient(low = ("red"),
                        high = ("blue")) +
  ggplot2::scale_x_date(date_labels = "%h %d", breaks = scales::pretty_breaks(7)) +
  ggplot2::ylab("% fully vaxxed") +
  ggplot2::ggtitle("vax rate per state per party")

vaccine_party_plot %>%
  plotly::ggplotly()

```

## Deaths since July 2021

```{r deaths}
deaths_since_july <- covid.cases_state %>%
  
  dplyr::group_by(date) %>%
    dplyr::summarise(total_deaths = sum(deaths)) %>%
    
  dplyr::ungroup() %>%
    
  dplyr::mutate(death_change = total_deaths - dplyr::lag(total_deaths)) %>%
    
  dplyr::mutate(seven_day_average_death = zoo::rollmean(death_change, 7, na.pad = TRUE)) %>%
    
  dplyr::filter(date >= as.Date("2021-07-01")) %>%
    
  ggplot2::ggplot(ggplot2::aes(x=date, y=seven_day_average_death)) +
  ggplot2::geom_path()+
  ggplot2::scale_y_continuous(breaks = scales::pretty_breaks(n = 20))+
  ggplot2::ggtitle("Covid Deaths per Day after 7/1/2021")+
  ggplot2::ylab("Seven Day Average Deaths")

deaths_since_july %>%
  plotly::ggplotly()

july2021_deaths <- covid.cases_state %>%
  
  dplyr::group_by(date) %>%
    dplyr::summarise(total_deaths = sum(deaths)) %>%
    
  dplyr::ungroup() %>%
    
  dplyr::mutate(death_change = total_deaths - dplyr::lag(total_deaths)) %>%
    
  dplyr::mutate(seven_day_average_death = zoo::rollmean(death_change, 7, na.pad = TRUE)) %>%
    
  dplyr::filter(date >= as.Date("2021-07-01")) %>%
    
  dplyr::filter(!is.na(seven_day_average_death))

perc_change <-
  july2021_deaths$seven_day_average_death[july2021_deaths$date == max(july2021_deaths$date)] /
  min(july2021_deaths$seven_day_average_death, na.rm = TRUE)


```

This is `r round(perc_change, digits = 1)` times the amount we had at the lowest 7 day average death count in July.






