---
title: "S&P 500 Perspective"
date: '`r format(Sys.time(), "%Y-%m-%d %X")`'
editor_options: 
  chunk_output_type: console
output:
  html_document:
    includes:
      in_header: header/header.html
---

```{r pipe, include=FALSE}
`%>%` <- magrittr::`%>%`
```

```{r setup, include=FALSE}
# Set default chunk output
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  results = "asis"
)
```

```{r libraries}
```

```{r grab_stock_data}

price_data = tidyquant::tq_get(x = "VOO", from = "2010-01-01", to = as.character(Sys.Date()), get = "stock.prices")

# buy the dip
max_change = price_data %>% dplyr::mutate(day_change = dplyr::lead(close) / close) %>%
  dplyr::filter(day_change == max(day_change, na.rm = TRUE))
max_date = max_change %>% dplyr::slice(1) %>% dplyr::pull(date)
max_change_perc = max_change %>% dplyr::slice(1) %>% dplyr::pull(day_change)
# scales::percent(max_change_perc - 1, accuracy = .1)
min_date = price_data %>%
  dplyr::filter(date == min(date, na.rm = TRUE)) %>% dplyr::pull(date)


url = "https://finance.yahoo.com/quote/VOO/history?p=VOO"

voo_history_html = rvest::read_html(url)
# //*[@id="Col1-1-HistoricalDataTable-Proxy"]/section/div[2]/table
voo_history = voo_history_html %>%
  # rvest::html_nodes("#Col1-1-HistoricalDataTable-Proxy > section") %>% 
  rvest::html_nodes(xpath = "//*[@id='Col1-1-HistoricalDataTable-Proxy']/section/div[2]/table") %>% 
  # html_name()
  rvest::html_table()

voo_history_close = voo_history[[1]] %>%
  dplyr::filter(!grepl(x = Open, pattern = "Dividend|Close")) %>%
  dplyr::mutate(date = as.Date(Date, format = "%b %d, %Y"),
         `Close*` = as.numeric(`Close*`)) %>%
  dplyr::select(date, close_yahoo = `Close*`) %>%
  dplyr::mutate(symbol = "VOO")

voo_full = price_data %>% 
  dplyr::select(symbol, date, close) %>% 
  dplyr::full_join(voo_history_close) %>% 
  dplyr::mutate(close_official = ifelse(is.na(close), close_yahoo, close)) %>% 
  dplyr::select(symbol, date, close_official) %>%
  dplyr::filter(!is.na(close_official))


voo_full$rolling_max = NA
for (i in 1:nrow(voo_full)) {
  voo_full$rolling_max[i] = rollmax(x = voo_full$close_official[1:i], k = i)
}

voo_maxes = voo_full %>% 
  dplyr::filter(close_official == rolling_max)


intersections_voo = voo_full %>%
  # dplyr::arrange(dplyr::desc(date)) %>% 
  dplyr::mutate(intersection = close_official >= close_official[date == max(date)],
         intersection_lead = dplyr::lead(intersection)) %>% 
  dplyr::filter(intersection != intersection_lead & !intersection)



```

## S&P 500 Over the Last 11 Years

Omg did the stock market just crash? 1% down in a day is a crash right? Well maybe not... Let's just take a look at the history of the S&P 500 Index Fund (VOO) for some perspective.

```{r sp500_last_2_years}
sp500_2018_pres_plot = voo_full %>% 
  ggplot2::ggplot(ggplot2::aes(x = date, y = close_official)) +
  ggplot2::geom_line() +
  ggplot2::geom_point(data = voo_maxes, ggplot2::aes(x = date, y = rolling_max), color = "red", size = .5) +
  ggplot2::ggtitle("S&P 500")+
  ggplot2::ylab("Closing Price")

sp500_2018_pres_plot %>% plotly::ggplotly()
```

Btw, every red dot is a record breaking closing day. We are currently at `r voo_maxes %>% nrow` record breaking closes since `r voo_full %>% dplyr::pull(date) %>% min`.

## S&P 500 More Recently

Let's zoom in on just the last three months. 

```{r sp500_intersections}
sp_intersections_plot = voo_full %>% 
  dplyr::filter(date >= min(intersections_voo$date %m-% months(2))) %>% 
  ggplot2::ggplot(ggplot2::aes(x = date, y = close_official)) +
  ggplot2::geom_line() +
  ggplot2::ggtitle("S&P 500")+
  ggplot2::ylab("Closing Price") +
  ggrepel::geom_text_repel(data = intersections_voo, ggplot2::aes(label = as.character(date), segment.color = "red"),
                           box.padding = .8,
                           nudge_y = 15,
                           direction = "x",
                           vjust = 0,
                           hjust = 1,
                           xlim =
                             c(voo_full %>% 
                                 dplyr::filter(date >= min(intersections_voo$date %m-% months(2))) %>%
                                 dplyr::pull(date) %>% min,
                               voo_full %>%
                                 dplyr::pull(date) %>% max)) +
  ggplot2::geom_hline(yintercept = voo_full$close_official[voo_full$date ==
                                                             max(voo_full$date)],
             linetype=2, colour="red") +
  ggplot2::labs(caption = "Note, the date captions note the price the day before\nthe price closes just under today's price/close")

sp_intersections_plot

```

Guess what? If you invested in the S&P 500 on any of the labeled dates or prior to `r min(intersections_voo$date)`, then you're still net positive. So calm down.

Also, the largest single day increase from close to close since `r min_date` was on `r max_date` with a single day jump of `r scales::percent(max_change_perc - 1, accuracy = .1)`. So, keep investing and continue to buy the dip.





